### 2016-02-09 podc16/paper2.tex
Added the draft of the brief annoucement.

\documentclass[11]{article}
\usepackage[english]{babel}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{comment}
\hypersetup{pdftex,colorlinks=true,allcolors=blue}
%\usepackage{hypcap}
\usepackage[url=true,doi=false,isbn=false,backend=biber]{biblatex}
\addbibresource{managed.bib}
%\setcounter{tocdepth}{2}
%\usepackage{pdfpages}
%\usepackage{wrapfig}
%\makeatletter
%\newcommand{\setword}[2]{%
%    \phantomsection%
%    #1\def\@currentlabel{\unexpanded{#1}}\label{#2}%
%}
%\makeatother

\title{Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms}

\author{}

\date{}

\begin{document}

\maketitle

\section{Introduction}

Many agreement algorithms inspired by Paxos and MultiPaxos~\cite{lamport2001paxos} rely on the existence of a unique leader replica which enforces an ordering on commands during fault-free periods.
In practice, the unique leader can easily become a performance bottleneck for the system.
Therefore, recently proposed agreement algorithms like Mencius~\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine}, EPaxos~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments}, or Alvin~\cite{TurcuETAL14BeGeneralDontGiveUpConsistency} strive to increase performance by avoiding the use of a unique leader.
Mencius rotates the leader among the replicas but its structure remains close to MultiPaxos. 
However, EPaxos and Alvin are based on a novel algorithmic idea first introduced in EPaxos.

In this paper, we show that the core algorithmic idea underlying EPaxos can be used to build a family of leaderless algorithms parameterized by (1) a per-command dependency-set computation algorithm, a concept that we define in the paper, and (2) an algorithm ensuring agreement on a mapping from commands to dependency sets. %(2) an algorithm ensuring a per-command agreement on a dependency set and which is oblivious of any ordering between commands. 
An algorithm in the family is obtained by instantiating the two parameters to a concrete agreement algorithm and a concrete dependency-set computation algorithm, which are then used as sub-routines. 
Note that any generalized consensus algorithm can be used to instantiate parameter 2, but that only a weaker problem than generalized consensus needs to be solved because no ordering constraints need to be enforced.

We prove that all the algorithms in the family are safe and live, and we have formalized our ideas in TLA+ and thoroughly model-check our claims with the TLC model-checker.

Our work sheds light on the common structure of algorithms like EPaxos and Alvin and allows to tune the performance of leaderless algorithms for specific scenarios by changing the underlying generalized-consensus and dependency-set algorithm. 
We expect that these two tunable knobs will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by this new family of algorithms.

\printbibliography%

\end{document}

### 2016-02-10 podc16/paper2.tex
Work on the ba.

@@
-\documentclass[11]{article}
-\usepackage[english]{babel}
-\usepackage{hyperref}
-\usepackage{cleveref}
-\usepackage{comment}
-\hypersetup{pdftex,colorlinks=true,allcolors=blue}
-%\usepackage{hypcap}
-\usepackage[url=true,doi=false,isbn=false,backend=biber]{biblatex}
-\addbibresource{managed.bib}
-%\setcounter{tocdepth}{2}
-%\usepackage{pdfpages}
-%\usepackage{wrapfig}
-%\makeatletter
-%\newcommand{\setword}[2]{%
-%    \phantomsection%
-%    #1\def\@currentlabel{\unexpanded{#1}}\label{#2}%
-%}
-%\makeatother
+\input{header}
 
@@
 
+\section{Leaderless Algorithms}
+
+We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network.
+We also consider a deterministic state-machine $M$ whose set of commands is $C$ and equipped with a dependency relation $R \subseteq C\times C$ such that for any command $c$, $\left( c,c \right)\notin R$; for any commands $c_1$ and $c_2$, if $\left( c_1,c_2 \right)\in R$ then $\left( c_2,c_1 \right)\in R$; if $\left( c_1,c_2 \right)\in R$ and $\aseq{s_1,c_1,s_2,c_2,\cdots,s_n}$
+
+\subsection{Computing Dependency-Sets}
+
+A dependency-set computation algorithm allows processes to announce a command $c$ and to learn about a suggested dependency set for a previously announced command.
+
+A dependency-set computation algorithm exposes the input actions ${announce\left( c \right)}^p$, where $c$ is a command, and the output actions ${suggest\left( c, deps \right)}^p$, where $c$ is a command and $deps$ is a set of commands.
+Moreover, a dependency-set computation algorithm must ensure that:
+\begin{itemize}[noitemsep,nolistsep]
+    \item For any suggestion ${suggest\left( c, deps \right)}$, the command $c$ must have been announced before.
+    \item for any two suggestions $suggest\left( c_1, deps_1 \right)$ and $suggest\left( c_2, deps_2 \right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_2$ or $c_2 \in deps_1$.
+\end{itemize}
+
+\subsection{Agreeing on Dependency Sets}
+
+A map-agreement algorithm exposes
+
 \printbibliography%

### 2016-02-10 podc16/paper2.tex
Work on paper2.tex

@@
 We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network.
-We also consider a deterministic state-machine $M$ whose set of commands is $C$ and equipped with a dependency relation $R \subseteq C\times C$ such that for any command $c$, $\left( c,c \right)\notin R$; for any commands $c_1$ and $c_2$, if $\left( c_1,c_2 \right)\in R$ then $\left( c_2,c_1 \right)\in R$; if $\left( c_1,c_2 \right)\in R$ and $\aseq{s_1,c_1,s_2,c_2,\cdots,s_n}$
+We also consider a deterministic state-machine $M$ whose set of commands is $C$ and equipped with a dependency relation $R \subseteq C\times C$ such that for any command $c$, $\left( c,c \right)\notin R$; for any commands $c_1$ and $c_2$, if $\left( c_1,c_2 \right)\in R$ then $\left( c_2,c_1 \right)\in R$; if $\left( c_a,c_b \right)\in R$ then $\aseq{s_1,c_1,\cdots,s_i,c_a,s_{i+1},c_b,\cdots,s_n}$ is an execution of the state-machine $M$ if and only if
+$\aseq{s_1,c_1,\cdots,s_i,c_b,s_{i+1},c_a,\cdots,s_n}$.
 
@@
 
-A map-agreement algorithm exposes
+A map-agreement algorithm exposes the input action $propose\left( c, deps \right)$
 
-\printbibliography%
+%\printbibliography%
+\bibliography{managed}{}
+\bibliographystyle{plain}
 

### 2016-02-10 podc16/paper2.tex
work on paper2.tex

--- 2016-02-10 podc16/paper2.tex
@@
 We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network.
-We also consider a deterministic state-machine $M$ whose set of commands is $C$ and equipped with a dependency relation $R \subseteq C\times C$ such that for any command $c$, $\left( c,c \right)\notin R$; for any commands $c_1$ and $c_2$, if $\left( c_1,c_2 \right)\in R$ then $\left( c_2,c_1 \right)\in R$; if $\left( c_1,c_2 \right)\in R$ and $\aseq{s_1,c_1,s_2,c_2,\cdots,s_n}$
+We also consider a deterministic state-machine $M$ whose set of commands is $C$ and equipped with a dependency relation symmetric relation $R \subseteq C\times C$ such that for any command $c$, $\left( c,c \right)\notin R$, and if $\left( c_a,c_b \right)\in R$ then $\aseq{s_1,c_1,\cdots,s_i,c_a,s_{i+1},c_b,\cdots,s_n}$ is an execution of the state-machine $M$ if and only if
+$\aseq{s_1,c_1,\cdots,s_i,c_b,s_{i+1},c_a,\cdots,s_n}$.
 
@@
 
-A dependency-set computation algorithm exposes the input actions ${announce\left( c \right)}^p$, where $c$ is a command, and the output actions ${suggest\left( c, deps \right)}^p$, where $c$ is a command and $deps$ is a set of commands.
+A dependency-set computation algorithm exposes the input actions ${\operatorname{announce}\left( c \right)}^p$, where $c$ is a command, and the output actions ${suggest\left( c, deps \right)}^p$, where $c$ is a command and $deps$ is a set of commands.
 Moreover, a dependency-set computation algorithm must ensure that:
 \begin{itemize}[noitemsep,nolistsep]
-    \item For any suggestion ${suggest\left( c, deps \right)}$, the command $c$ must have been announced before.
+    \item For any suggestion ${suggest\left( c, deps \right)}$, every command in $c\cup deps$ must have been announced before.
     \item for any two suggestions $suggest\left( c_1, deps_1 \right)$ and $suggest\left( c_2, deps_2 \right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_2$ or $c_2 \in deps_1$.
@@
 
-A map-agreement algorithm exposes
+A map-agreement algorithm exposes the input action $propose\left( c, deps \right)$
 

### 2016-02-10 podc16/paper2.tex
work on paper2.tex

@@
     \item For any suggestion ${suggest\left( c, deps \right)}$, every command in $c\cup deps$ must have been announced before.
-    \item for any two suggestions $suggest\left( c_1, deps_1 \right)$ and $suggest\left( c_2, deps_2 \right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_2$ or $c_2 \in deps_1$.
+    \item for any two suggestions $suggest\left( c_1, deps_1 \right)$ and $suggest\left( c_2, deps_2 \right)$,  if $c_1$ and $c_2$ do not commute ($\left( c_1,c_2 \right) \in R$) then either $c_1\in deps_2$ or $c_2 \in deps_1$.
 \end{itemize}

### 2016-02-10 podc16/paper2.tex
commute

--- 2016-02-10 podc16/paper2.tex
@@
 We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network.
-We also consider a deterministic state-machine $M$ whose set of commands is $C$ and equipped with a dependency relation symmetric relation $R \subseteq C\times C$ such that for any command $c$, $\left( c,c \right)\notin R$, and if $\left( c_a,c_b \right)\in R$ then $\aseq{s_1,c_1,\cdots,s_i,c_a,s_{i+1},c_b,\cdots,s_n}$ is an execution of the state-machine $M$ if and only if
-$\aseq{s_1,c_1,\cdots,s_i,c_b,s_{i+1},c_a,\cdots,s_n}$.
+We also consider a deterministic state-machine $M$ whose set of commands is $C$.
+We say that two commands $c_a$, $c_b$ \textit{commute} if executing $c_a$ before $c_b$ has the same effect as executing $c_b$ before $c_a$.
+We also consider a symmetric relation $R \subseteq C\times C$, called \textit{dependency relation}, such that: for any command $c$, $\left( c,c \right)\notin R$;  if $\left( c_a,c_b \right)\notin R$ $c_a$, $c_b$ \textit{commute}.
+then $\aseq{s_1,c_1,\cdots,s_i,c_a,s_{i+1},c_b,s_{i+2}\cdots,s_n}$ is an execution of the state-machine $M$ if and only if
+%$\aseq{s_1,c_1,\cdots,s_i,c_b,s_{i+1},c_a,\cdots,s_n}$.
+%$\aseq{s_1,c_1,\cdots,s_i,c_b,s_{i+2},c_a,s_{i+1},\cdots,s_n}$
 
@@
 \begin{itemize}[noitemsep,nolistsep]
-    \item For any suggestion ${suggest\left( c, deps \right)}$, every command in $c\cup deps$ must have been announced before.
-    \item for any two suggestions $suggest\left( c_1, deps_1 \right)$ and $suggest\left( c_2, deps_2 \right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_2$ or $c_2 \in deps_1$.
+    \item For any suggestion ${suggest\left( c, deps \right)}^p$, every command in $\{c\}\cup deps$ must have been announced before.
+    \item For any two suggestions $suggest\left( c_1, deps_1 \right)^p$ and $suggest\left( c_2, deps_2 \right)^p$,  if $c_1$ and $c_2$ do not commute (i.e., $\left( c_1,c_2 \right) \in R$) then either $c_1\in deps_2$ or $c_2 \in deps_1$.
 \end{itemize}

### 2016-02-10 podc16/paper2.tex
Work on map-agreement.

--- 2016-02-10 podc16/paper2.tex
@@
 
-A map-agreement algorithm exposes the input action $propose\left( c, deps \right)$
+A map-agreement algorithm exposes the input actions $propose\left( c, deps \right)^p$ and the output actions $decide\left( c, deps \right)^p$, for a command $c$, a set of commands $deps$, and a process $p$.
+Moreover, a map-agreement algorithm must ensure that 
+\begin{itemize}[noitemsep,nolistsep]
+    \item test
+\end{itemize}
 

### 2016-02-10 podc16/paper2.tex
Added comments on the basic model and commutativity.

@@
 We also consider a deterministic state-machine $M$ whose set of commands is $C$. 
-We say that two commands $c_a$, $c_b$ \textit{commute} if executing $c_a$ before $c_b$ has the same effect as executing $c_b$ before $c_a$.
-We also consider a symmetric relation $R \subseteq C\times C$, called \textit{dependency relation}, such that: for any command $c$, $\left( c,c \right)\notin R$;  if $\left( c_a,c_b \right)\notin R$ $c_a$, $c_b$ \textit{commute}.
-then $\aseq{s_1,c_1,\cdots,s_i,c_a,s_{i+1},c_b,s_{i+2}\cdots,s_n}$ is an execution of the state-machine $M$ if and only if
-%$\aseq{s_1,c_1,\cdots,s_i,c_b,s_{i+1},c_a,\cdots,s_n}$.
-%$\aseq{s_1,c_1,\cdots,s_i,c_b,s_{i+2},c_a,s_{i+1},\cdots,s_n}$
+We say that two commands $c_a$, $c_b$ \textit{commute} if executing $c_a$ before $c_b$ has the same effect on the state machine as executing $c_b$ before $c_a$.
+We assume that a command does not commute with itself.
+We also consider a symmetric relation $R \subseteq C\times C$, called \textit{dependency relation}, such that: for any command $c$, $\left( c,c \right)\notin R$, and if $\left( c_a,c_b \right)\notin R$ then $c_a$ and $c_b$ commute.
+
+\todo{Why do we need $R$ if we already define what commute and non-commutative means?}
+
+\todo{About the uniqueness of commands: why not leave that our of our scope?}
+
+\todo{Why not base everything on C-Structs? We can say that two commands commute when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $ for every cstruct $\sigma$.}
 

### 2016-02-10 podc16/paper2.tex
Work on "agreeing on dependency sets"

@@
 
-A dependency-set computation algorithm exposes the input actions ${\operatorname{announce}\left( c \right)}^p$, where $c$ is a command, and the output actions ${suggest\left( c, deps \right)}^p$, where $c$ is a command and $deps$ is a set of commands.
+A dependency-set computation algorithm exposes the input actions ${\operatorname{announce}\left( c \right)}^p$, where $c$ is a command, and the output actions ${suggest\left( c, d \right)}^p$, where $c$ is a command and $d$ is a set of commands.
 Moreover, a dependency-set computation algorithm must ensure that:
-\begin{itemize}[noitemsep,nolistsep]
-    \item For any suggestion ${suggest\left( c, deps \right)}^p$, every command in $\{c\}\cup deps$ must have been announced before.
-    \item For any two suggestions $suggest\left( c_1, deps_1 \right)^p$ and $suggest\left( c_2, deps_2 \right)^p$,  if $c_1$ and $c_2$ do not commute (i.e., $\left( c_1,c_2 \right) \in R$) then either $c_1\in deps_2$ or $c_2 \in deps_1$.
-\end{itemize}
+\begin{enumerate}[noitemsep,nolistsep]
+    \item For any suggestion ${suggest\left( c, d \right)}^p$, every command in $\{c\}\cup d$ must have been announced before.
+    \item For any two suggestions ${suggest\left( c_1, d_1 \right)}^p$ and ${suggest\left( c_2, d_2 \right)}^p$,  if $c_1$ and $c_2$ do not commute (i.e., $\left( c_1,c_2 \right) \in R$) then either $c_1\in d_2$ or $c_2 \in d_1$.
+\end{enumerate}
+
+\todo{Plain english version below. Which is better?}
+
+A dependency-set computation algorithm allows processes to announce a command $c$ and to learn about a suggested dependency set for a previously announced command.
+Moreover, a dependency-set computation algorithm must ensure that:
+\begin{enumerate}[noitemsep,nolistsep]
+    \item If the dependency set $d$ has been suggested for $c$, then every command in $\aset{c}\cup d$ must have been announced before.
+    \item If the dependency set $d_1$ has been suggested for the command $c_1$, the dependency set $d_2$ has been suggested for the command $c_2$, and $c_1$ and $c_2$ do not commute, then either $c_1\in d_2$ or $c_2 \in d_1$.
+\end{enumerate}
 
@@
 
-A map-agreement algorithm exposes the input actions $propose\left( c, deps \right)^p$ and the output actions $decide\left( c, deps \right)^p$, for a command $c$, a set of commands $deps$, and a process $p$.
-Moreover, a map-agreement algorithm must ensure that 
-\begin{itemize}[noitemsep,nolistsep]
-    \item test
-\end{itemize}
+A map-agreement algorithm exposes the input actions ${propose\left( c, d \right)}^p$ and the output actions ${decide\left( c, d \right)}^p$, for a command $c$, a set of commands $d$, and a process $p$.
+Moreover, a map-agreement algorithm must ensure that: 
+\begin{enumerate}[noitemsep,nolistsep]
+    \item If the set of dependencies $d$ is decided for a command $c$, then $d$ has been proposed for $c$ at an earlier time.
+    \item If $d_1$ is decided for a command $c$ and $d_2$ is decided for the same command $c$, then $d_1$ is equal to $d_2$.
+\end{enumerate}
+Note that any SMR algorithm, like MultiPaxos, can be used to implementing map-agreement by uniquely associating commands to MultiPaxos instance numbers.
 

### 2016-02-10 podc16/paper2.tex
started explaining the algo, and other comments.

@@
 
+\todo{Let's base everything on C-Structs; it will be easier. We can say that two commands commute when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $ for every cstruct $\sigma$.}
+
 We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network.
@@
 
-\todo{Why do we need $R$ if we already define what commute and non-commutative means?}
-
-\todo{About the uniqueness of commands: why not leave that our of our scope?}
-
-\todo{Why not base everything on C-Structs? We can say that two commands commute when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $ for every cstruct $\sigma$.}
-
 \subsection{Computing Dependency-Sets}
-
-A dependency-set computation algorithm allows processes to announce a command $c$ and to learn about a suggested dependency set for a previously announced command.
 
@@
 
+\subsection{Implementing Leaderless Generalized Consensus}
+
+Genearalized consensus~\cite{Lamport05GeneralizeConsensus} is a specification for state-machine replication algorithms. 
+A generalized-consensus algorithm exposes the input actions $propose\left( c \right)$, for a command $c$, and $learn\left( \sigma \right)$, for a c-struct $\sigma$.
+Moreover, a generalized-consensus algorithm must ensure that TODO.
+
+We now propose an algorithm that uses the dependency-set computation abstraction and the map-agreement abstraction to implement generalized consensus.
+We therefore obtain a family of generalized consensus algorithms where each member of the family is obtained by using concrete dependency-set and map-agreement algorithms.
+
+\todo{Explain the algorithm with pseudocode?}
+
+\todo{Explain the graph processing algo independently? Its key property is that when $G1 \subseteq G2$ then $lin\left( c_1,G1 \right)$ is a prefix of $lin\left( c_2,G2 \right)$ or vice-versa.}
+
 \printbibliography%

### 2016-02-10 podc16/paper2.tex
Work on explaining the algorithm.

@@
 
+\todo{``Suggest'' can also be used to trigger recovery}.
+
 \subsection{Agreeing on Dependency Sets}
@@
 \todo{Explain the algorithm with pseudocode?}
+Local variables: committed, a map from commands to dependency sets, initially the empty map.
+\begin{itemize}[noitemsep,nolistsep]
+    \item Upon ${propose\left( c \right)}^p$: call ${announce\left( c \right)}^p$.
+    \item Upon ${suggest\left( c,d \right)}^p$: call ${propose\left( c,d \right)}^p$.
+    \item Upon ${decide\left( c,d \right)}^p$: insert the mapping $c \rightarrow d$ in the map ``committed''.
+    \item Periodically: process the committed commands, as described below.
+\end{itemize}
+
+The variable ``committed'' of the algorithm above can be seen as describing a directed graph $g$ whose vertices are commands and such that there is an edge from $c_1$ to $c_2$ if and only if ``committed'' maps $v_1$ to a set $deps$ where $v_2\in deps$, i.e.
+$v_1$ depends on $v_2$.
 

### 2016-02-10 podc16/paper2.tex
c-struct and notation

--- 2016-02-10 podc16/paper2.tex
@@
 We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network.
-We also consider a deterministic state-machine $M$ whose set of commands is $C$. 
-We say that two commands $c_a$, $c_b$ \textit{commute} if executing $c_a$ before $c_b$ has the same effect on the state machine as executing $c_b$ before $c_a$.
+Processes in $P$ must solve the Generalized Consensus problem~\cite{lamport-generalized}.
+Each process receives proposals for commands of the form \texttt{Propose}$(c)$, and it produces responses of the form \texttt{Decide}$(\sigma)$, where $\sigma$ is a c-struct.
+A c-struct set is a set, containing the element $\bot$, with an append operator $\bullet$ such that $\sigma \bullet c$ is a c-struct, for any c-struct $\sigma$ and command $c$.
+We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
+Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands (see~\cite{lamport-generalized} for a complete discussion on c-struct).
+In the rest of the paper, unless otherwise specified, the symbol $c$, possibly sub-scripted, ranges over commands, and the symbol $\mathcal{D}$, possibly sub-scripted, ranges over sets of commands.
+
+In the next sections we present our specification of leaderless generalized consensus algorithms, which is parameterized over a \textit{dependency-set} algorithm (\Cref{dep-algo}) and a \textit{map-agreement} algorithm (\Cref{map-algo}).
+%Processes proposals of commands to the \textit{Generalized Consensus} layer
+%We also consider a deterministic state-machine $M$ whose set of commands is $C$. 
+%We say that two commands $c_a$, $c_b$ \textit{commute} if executing $c_a$ before $c_b$ has the same effect on the state machine as executing $c_b$ before $c_a$.
 We assume that a command does not commute with itself.
-We also consider a symmetric relation $R \subseteq C\times C$, called \textit{dependency relation}, such that: for any command $c$, $\left( c,c \right)\notin R$, and if $\left( c_a,c_b \right)\notin R$ then $c_a$ and $c_b$ commute.
+%We also consider a symmetric relation $R \subseteq C\times C$, called \textit{dependency relation}, such that: for any command $c$, $\left( c,c \right)\notin R$, and if $\left( c_a,c_b \right)\notin R$ then $c_a$ and $c_b$ commute.
 
 \subsection{Computing Dependency-Sets}
+\label{dep-algo}
 
-A dependency-set computation algorithm exposes the input actions ${\operatorname{announce}\left( c \right)}^p$, where $c$ is a command, and the output actions ${suggest\left( c, d \right)}^p$, where $c$ is a command and $d$ is a set of commands.
+A dependency-set algorithm exposes the input interface $\texttt{announce}_p\left( c \right)$, where $c$ is a command, and the output interface $\texttt{suggest}_p \left( c,\mathcal{D}\right)$, where $p\in P$ is the process executing the call.
 Moreover, a dependency-set computation algorithm must ensure that:
@@
 \subsection{Agreeing on Dependency Sets}
+\label{map-algo}
 

### 2016-02-10 podc16/paper2.tex
dependency-set algorithm specification

@@
 
-\todo{Let's base everything on C-Structs; it will be easier. We can say that two commands commute when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $ for every cstruct $\sigma$.}
+%\todo{Let's base everything on C-Structs; it will be easier. We can say that two commands commute when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $ for every cstruct $\sigma$.}
 
@@
 A dependency-set algorithm exposes the input interface $\texttt{announce}_p\left( c \right)$, where $c$ is a command, and the output interface $\texttt{suggest}_p \left( c,\mathcal{D}\right)$, where $p\in P$ is the process executing the call.
-Moreover, a dependency-set computation algorithm must ensure that:
+Moreover, a dependency-set algorithm must ensure that:
 \begin{enumerate}[noitemsep,nolistsep]
-    \item For any suggestion ${suggest\left( c, d \right)}^p$, every command in $\{c\}\cup d$ must have been announced before.
-    \item For any two suggestions ${suggest\left( c_1, d_1 \right)}^p$ and ${suggest\left( c_2, d_2 \right)}^p$,  if $c_1$ and $c_2$ do not commute (i.e., $\left( c_1,c_2 \right) \in R$) then either $c_1\in d_2$ or $c_2 \in d_1$.
+    \item For any suggestion $\texttt{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before.
+    \item For any two suggestions $\texttt{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\texttt{suggest}_q \left( c_2,\mathcal{D}_2\right)$ (with $p$ possibly equal to $q$),  if $c_1$ and $c_2$ do not commute then either $c_1\in d_2$ or $c_2 \in d_1$.
 \end{enumerate}
 
-\todo{Plain english version below. Which is better?}
+%\todo{Plain english version below. Which is better?}
 
-A dependency-set computation algorithm allows processes to announce a command $c$ and to learn about a suggested dependency set for a previously announced command.
-Moreover, a dependency-set computation algorithm must ensure that:
-\begin{enumerate}[noitemsep,nolistsep]
-    \item If the dependency set $d$ has been suggested for $c$, then every command in $\aset{c}\cup d$ must have been announced before.
-    \item If the dependency set $d_1$ has been suggested for the command $c_1$, the dependency set $d_2$ has been suggested for the command $c_2$, and $c_1$ and $c_2$ do not commute, then either $c_1\in d_2$ or $c_2 \in d_1$.
-\end{enumerate}
+%A dependency-set computation algorithm allows processes to announce a command $c$ and to learn about a suggested dependency set for a previously announced command.
+%Moreover, a dependency-set computation algorithm must ensure that:
+%\begin{enumerate}[noitemsep,nolistsep]
+  %  \item If the dependency set $d$ has been suggested for $c$, then every command in $\aset{c}\cup d$ must have been announced before.
+  %  \item If the dependency set $d_1$ has been suggested for the command $c_1$, the dependency set $d_2$ has been suggested for the command $c_2$, and $c_1$ and $c_2$ do not commute, then either $c_1\in d_2$ or $c_2 \in d_1$.
+%\end{enumerate}
 

### 2016-02-10 podc16/paper2.tex
cover page and margins
correct typo in dependency-sets section

@@
 
-\title{Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms}
+%\title{Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms}
 
-\author{}
+%\author{}
 
-\date{}
+%\date{}
 
@@
 
-\maketitle
+%\maketitle
+
+\input{cover}
 
@@
     \item For any suggestion $\texttt{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before.
-    \item For any two suggestions $\texttt{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\texttt{suggest}_q \left( c_2,\mathcal{D}_2\right)$ (with $p$ possibly equal to $q$),  if $c_1$ and $c_2$ do not commute then either $c_1\in d_2$ or $c_2 \in d_1$.
+    \item For any two suggestions $\texttt{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\texttt{suggest}_q \left( c_2,\mathcal{D}_2\right)$ (with $p$ possibly equal to $q$),  if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
 \end{enumerate}

### 2016-02-10 podc16/paper2.tex
map-agreement section

@@
 
-\section{Leaderless Algorithms}
+\section{Leaderless Generalized Consensus Algorithms}
 
@@
 
-A dependency-set algorithm exposes the input interface $\texttt{announce}_p\left( c \right)$, where $c$ is a command, and the output interface $\texttt{suggest}_p \left( c,\mathcal{D}\right)$, where $p\in P$ is the process executing the call.
+A dependency-set algorithm exposes the input interface $\texttt{announce}_p\left( c \right)$, to announce a command $c$, and the output interface $\texttt{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
 Moreover, a dependency-set algorithm must ensure that:
@@
 
-A map-agreement algorithm exposes the input actions ${propose\left( c, d \right)}^p$ and the output actions ${decide\left( c, d \right)}^p$, for a command $c$, a set of commands $d$, and a process $p$.
+A map-agreement algorithm exposes the input interface $\texttt{propose}_p\left( c, \mathcal{D} \right)$, to propose a set of dependencies $\mathcal{D}$ for a command $c$, and the output interface $\texttt{decide}_p\left( c, \mathcal{D} \right)$, to decide $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
 Moreover, a map-agreement algorithm must ensure that: 
 \begin{enumerate}[noitemsep,nolistsep]
-    \item If the set of dependencies $d$ is decided for a command $c$, then $d$ has been proposed for $c$ at an earlier time.
-    \item If $d_1$ is decided for a command $c$ and $d_2$ is decided for the same command $c$, then $d_1$ is equal to $d_2$.
+    \item If the set of dependencies $\mathcal{D}$ is decided for a command $c$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
+    \item If $\mathcal{D}_1$ is decided for a command $c$ and $\mathcal{D}_2$ is decided for the same command $c$, then $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
  \end{enumerate}
+
+\todo{OR MORE FORMALLY}
+
+\begin{enumerate}[noitemsep,nolistsep]
+    \item For any decision $\texttt{decide}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
+    \item For any two decisions $\texttt{decide}_p\left( c, \mathcal{D}_1 \right)$, $\texttt{decide}_q\left( c, \mathcal{D}_2 \right)$ (with $p$ possibly equal to $q$), then $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
+
+\end{enumerate}
+
 Note that any SMR algorithm, like MultiPaxos, can be used to implementing map-agreement by uniquely associating commands to MultiPaxos instance numbers.

### 2016-02-10 podc16/paper2.tex
trivial changes

@@
 We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network.
-Processes in $P$ must solve the Generalized Consensus problem~\cite{lamport-generalized}.
-Each process receives proposals for commands of the form \texttt{Propose}$(c)$, and it produces responses of the form \texttt{Decide}$(\sigma)$, where $\sigma$ is a c-struct.
-A c-struct set is a set, containing the element $\bot$, with an append operator $\bullet$ such that $\sigma \bullet c$ is a c-struct, for any c-struct $\sigma$ and command $c$.
+Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}.
+Each process receives proposals for commands of the form $\operatorname{Propose}(c)$, and it produces responses of the form $\operatorname{Decide}(\sigma)$, where $\sigma$ is a c-struct.
+A c-struct set is a set, containing the element $\bot$, with an append operator
+$\bullet$ such that $\sigma \bullet c$ is a c-struct, for any c-struct $\sigma$ and command $c$.
 We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
-Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands (see~\cite{lamport-generalized} for a complete discussion on c-struct).
+Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands (see~\cite{Lamport05GeneralizeConsensus} for a complete discussion on c-struct).
 In the rest of the paper, unless otherwise specified, the symbol $c$, possibly sub-scripted, ranges over commands, and the symbol $\mathcal{D}$, possibly sub-scripted, ranges over sets of commands.
@@
 
-A dependency-set algorithm exposes the input interface $\texttt{announce}_p\left( c \right)$, to announce a command $c$, and the output interface $\texttt{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
+A dependency-set algorithm exposes the input interface $\operatorname{announce}_p\left( c \right)$, to announce a command $c$, and the output interface $\operatorname{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
 Moreover, a dependency-set algorithm must ensure that:
 \begin{enumerate}[noitemsep,nolistsep]
-    \item For any suggestion $\texttt{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before.
-    \item For any two suggestions $\texttt{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\texttt{suggest}_q \left( c_2,\mathcal{D}_2\right)$ (with $p$ possibly equal to $q$),  if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
+    \item For any suggestion $\operatorname{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before.
+    \item For any two suggestions $\operatorname{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\operatorname{suggest}_q \left( c_2,\mathcal{D}_2\right)$ (with $p$ possibly equal to $q$),  if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
 \end{enumerate}
@@
 
-A map-agreement algorithm exposes the input interface $\texttt{propose}_p\left( c, \mathcal{D} \right)$, to propose a set of dependencies $\mathcal{D}$ for a command $c$, and the output interface $\texttt{decide}_p\left( c, \mathcal{D} \right)$, to decide $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
+A map-agreement algorithm exposes the input interface $\operatorname{propose}_p\left( c, \mathcal{D} \right)$, to propose a set of dependencies $\mathcal{D}$ for a command $c$, and the output interface $\operatorname{decide}_p\left( c, \mathcal{D} \right)$, to decide $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
 Moreover, a map-agreement algorithm must ensure that: 
@@
 \begin{enumerate}[noitemsep,nolistsep]
-    \item For any decision $\texttt{decide}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
-    \item For any two decisions $\texttt{decide}_p\left( c, \mathcal{D}_1 \right)$, $\texttt{decide}_q\left( c, \mathcal{D}_2 \right)$ (with $p$ possibly equal to $q$), then $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
+    \item For any decision $\operatorname{decide}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
+    \item For any two decisions $\operatorname{decide}_p\left( c, \mathcal{D}_1 \right)$, $\operatorname{decide}_q\left( c, \mathcal{D}_2 \right)$ (with $p$ possibly equal to $q$), then $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
 

### 2016-02-10 podc16/paper2.tex
minor changes

@@
 \todo{Explain the algorithm with pseudocode?}
-Local variables: committed, a map from commands to dependency sets, initially the empty map.
+
+Local variables: $\operatorname{committed}_p$, a map from commands to dependency sets, initially the empty map.
 \begin{itemize}[noitemsep,nolistsep]
-    \item Upon ${propose\left( c \right)}^p$: call ${announce\left( c \right)}^p$.
-    \item Upon ${suggest\left( c,d \right)}^p$: call ${propose\left( c,d \right)}^p$.
-    \item Upon ${decide\left( c,d \right)}^p$: insert the mapping $c \rightarrow d$ in the map ``committed''.
-    \item Periodically: process the committed commands, as described below.
+    \item Upon ${\operatorname{propose}\left( c \right)}^p$: call ${\operatorname{announce}\left( c \right)}^p$.
+    \item Upon ${\operatorname{suggest}\left( c,d \right)}^p$: call ${\operatorname{propose}\left( c,d \right)}^p$.
+    \item Upon ${\operatorname{decide}\left( c,d \right)}^p$: insert the mapping $c \rightarrow d$ in the map $\operatorname{committed}_p$.
+    \item Periodically,  every process $p$ processes $\operatorname{committed}_p$, as described below.
 \end{itemize}
 
-The variable ``committed'' of the algorithm above can be seen as describing a directed graph $g$ whose vertices are commands and such that there is an edge from $c_1$ to $c_2$ if and only if ``committed'' maps $v_1$ to a set $deps$ where $v_2\in deps$, i.e.
-$v_1$ depends on $v_2$.
+The local variable $committed_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $\operatorname{commited}_p$ and such that there is an edge from $c_1$ to $c_2$ if and only both are in $V\left( g_p \right)$ and $v_2 \in \operatorname{commited}_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$).
+%Let us assume that
 

### 2016-02-11 podc16/paper2.tex
description of linearization functions

--- 2016-02-10 podc16/paper2.tex
@@
 Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}.
-Each process receives proposals for commands of the form $\operatorname{Propose}(c)$, and it produces responses of the form $\operatorname{Decide}(\sigma)$, where $\sigma$ is a c-struct.
+Each process receives proposals for commands of the form $\opn{Propose}(c)$, and it produces responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct.
 A c-struct set is a set, containing the element $\bot$, with an append operator
@@
 
-A dependency-set algorithm exposes the input interface $\operatorname{announce}_p\left( c \right)$, to announce a command $c$, and the output interface $\operatorname{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
+A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, and the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
 Moreover, a dependency-set algorithm must ensure that:
 \begin{enumerate}[noitemsep,nolistsep]
-    \item For any suggestion $\operatorname{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before.
-    \item For any two suggestions $\operatorname{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\operatorname{suggest}_q \left( c_2,\mathcal{D}_2\right)$ (with $p$ possibly equal to $q$),  if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
+    \item For any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before.
+    \item For any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$ (with $p$ possibly equal to $q$),  if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
 \end{enumerate}
@@
 
-A map-agreement algorithm exposes the input interface $\operatorname{propose}_p\left( c, \mathcal{D} \right)$, to propose a set of dependencies $\mathcal{D}$ for a command $c$, and the output interface $\operatorname{decide}_p\left( c, \mathcal{D} \right)$, to decide $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
+A map-agreement algorithm exposes the input interface $\opn{propose}_p\left( c, \mathcal{D} \right)$, to propose a set of dependencies $\mathcal{D}$ for a command $c$, and the output interface $\opn{decide}_p\left( c, \mathcal{D} \right)$, to decide $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
 Moreover, a map-agreement algorithm must ensure that: 
@@
 \begin{enumerate}[noitemsep,nolistsep]
-    \item For any decision $\operatorname{decide}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
-    \item For any two decisions $\operatorname{decide}_p\left( c, \mathcal{D}_1 \right)$, $\operatorname{decide}_q\left( c, \mathcal{D}_2 \right)$ (with $p$ possibly equal to $q$), then $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
+    \item For any decision $\opn{decide}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
+    \item For any two decisions $\opn{decide}_p\left( c, \mathcal{D}_1 \right)$, $\opn{decide}_q\left( c, \mathcal{D}_2 \right)$ (with $p$ possibly equal to $q$), then $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
 
@@
 \todo{Explain the algorithm with pseudocode?}
-Local variables: committed, a map from commands to dependency sets, initially the empty map.
+
+Local variables: $\opn{cs}_p$, a map from commands to dependency sets, initially the empty map.
 \begin{itemize}[noitemsep,nolistsep]
-    \item Upon ${propose\left( c \right)}^p$: call ${announce\left( c \right)}^p$.
-    \item Upon ${suggest\left( c,d \right)}^p$: call ${propose\left( c,d \right)}^p$.
-    \item Upon ${decide\left( c,d \right)}^p$: insert the mapping $c \rightarrow d$ in the map ``committed''.
-    \item Periodically: process the committed commands, as described below.
+    \item Upon ${\opn{propose}\left( c \right)}^p$: call ${\opn{announce}\left( c \right)}^p$.
+    \item Upon ${\opn{suggest}\left( c,d \right)}^p$: call ${\opn{propose}\left( c,d \right)}^p$.
+    \item Upon ${\opn{decide}\left( c,d \right)}^p$: insert the mapping $c \rightarrow d$ in the map $\opn{cs}_p$.
+    \item Periodically,  every process $p$ processes $\opn{cs}_p$, as described below.
 \end{itemize}
 
-The variable ``committed'' of the algorithm above can be seen as describing a directed graph $g$ whose vertices are commands and such that there is an edge from $c_1$ to $c_2$ if and only if ``committed'' maps $v_1$ to a set $deps$ where $v_2\in deps$, i.e.
-$v_1$ depends on $v_2$.
+The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $\opn{cs}_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $v_2 \in \opn{cs}_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$).
+For example, if $cs_p = \left[ c_1 \mapsto \aset{c_2,c_3}, c_2 \mapsto \aset{c_1} \right]$ then $V\left(
+g_p \right) = \aset{c_1,c_2}$ and $E\left( g_p \right) = \aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right)}$.
+Given a graph $g$, let $\opn{sccs}\left( g \right)$ be the graph whose vertices are the strongly connected components of $g$ and where there is an edge from $scc_1$ to $scc_2$ if and only if there are two vertices $v_1$ and $v_2$ of $g$ such that $v_1 \in V\left( scc_1 \right)$, $v_2 \in V\left( scc_2 \right)$, and $\left( v_1,v_2 \right)\in E\left( g \right)$.
+Note that $\opn{sccs}\left( g \right)$ is necessarily a forest.
+
+Define the set of linearizations $\opn{Lins}\left( g \right)$ of a graph $g$ a the set of sequences whose length is the cardinality of $V\left( g \right)$, which contain each element of $V\left( g \right)$ exactly once, and ...
+\todo{or first define the partial order induced by a graph and then its compatible total orders.}
+
+Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ cs \right]$ contains each command in $cs$ exactly once.
+A sequence of commands $cseq$ is an $f$-linearization of the graph $g$ if and only there is a sequence whose element of strongly connected
 
 \todo{Explain the graph processing algo independently? Its key property is that when $G1 \subseteq G2$ then $lin\left( c_1,G1 \right)$ is a prefix of $lin\left( c_2,G2 \right)$ or vice-versa.}
+
+\subsection{Correctenss Proof}
+
+\subsection{Obtaining EPaxos}
 

### 2016-02-11 podc16/paper2.tex
work on describing the linearization algorithm + wrote an introductory paragraph to describe the main idea.

@@
 
-%\todo{Let's base everything on C-Structs; it will be easier. We can say that two commands commute when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $ for every cstruct $\sigma$.}
-
 We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network.
@@
 Each process receives proposals for commands of the form $\opn{Propose}(c)$, and it produces responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct.
-A c-struct set is a set, containing the element $\bot$, with an append operator
-$\bullet$ such that $\sigma \bullet c$ is a c-struct, for any c-struct $\sigma$ and command $c$.
+A c-struct set is a set, containing the element $\bot$, with an append operator $\bullet$ such that $\sigma \bullet c$ is a c-struct, for any c-struct $\sigma$ and command $c$.
+We also write the $\sigma\bullet\aseq{c_1,\cdots,c_n}$ for the c-struct $\sigma\bullet c_1\bullet\cdots\bullet c_n$. 
+C-structs are partially ordered by the relation $\preceq$ defined such that $\sigma_1\preceq\sigma_2$ iff there exists a sequence of commands $cs$ such that $\sigma_2=\sigma_1\bullet cs$.
 We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
 Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands (see~\cite{Lamport05GeneralizeConsensus} for a complete discussion on c-struct).
-In the rest of the paper, unless otherwise specified, the symbol $c$, possibly sub-scripted, ranges over commands, and the symbol $\mathcal{D}$, possibly sub-scripted, ranges over sets of commands.
+In the rest of the paper the symbol $c$, possibly sub-scripted, ranges over commands, and the symbol $\mathcal{D}$, possibly sub-scripted, ranges over sets of commands.
+\todo{Present generalized consensus here?}
 
-In the next sections we present our specification of leaderless generalized consensus algorithms, which is parameterized over a \textit{dependency-set} algorithm (\Cref{dep-algo}) and a \textit{map-agreement} algorithm (\Cref{map-algo}).
-%Processes proposals of commands to the \textit{Generalized Consensus} layer
-%We also consider a deterministic state-machine $M$ whose set of commands is $C$. 
-%We say that two commands $c_a$, $c_b$ \textit{commute} if executing $c_a$ before $c_b$ has the same effect on the state machine as executing $c_b$ before $c_a$.
-We assume that a command does not commute with itself.
-%We also consider a symmetric relation $R \subseteq C\times C$, called \textit{dependency relation}, such that: for any command $c$, $\left( c,c \right)\notin R$, and if $\left( c_a,c_b \right)\notin R$ then $c_a$ and $c_b$ commute.
+We now present our abstract leaderless generalized consensus algorithm.
+A process $p$ receiving a command $c$ proceeds in three phases.
+First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $cs$ that the command $c$ may depend on.
+Second, the process $p$ proposes the mapping  $c\mapsto cs$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto cs'$ for $c$, in which case $cs'$ becomes the set of dependencies of $c$, and we say that $c$ is executable when its dependency set is empty or when all its dependencies are executable.
+Finally, when $c$ is executable, the process $p$ may execute $c$ by deciding the c-struct obtained by running the \emph{graph-processing} algorithm described in \Cref{graph-proc}.
+Any other process can run the three phases independently of $p$ and eventually execute the command.
+However, we show that a leaderless  generalized consensus algorithm is always safe and is live when at least a majority of the processes are correct.
+
+%In the next sections we present our specification of leaderless generalized consensus algorithms, which is parameterized over a \textit{dependency-set} algorithm (\Cref{dep-algo}) and a \textit{map-agreement} algorithm (\Cref{map-algo}).
 
@@
 
+\subsection{Graph Processing}
+\label{graph-proc}
+
 The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $\opn{cs}_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $v_2 \in \opn{cs}_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$).
@@
 g_p \right) = \aset{c_1,c_2}$ and $E\left( g_p \right) = \aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right)}$.
-Given a graph $g$, let $\opn{sccs}\left( g \right)$ be the graph whose vertices are the strongly connected components of $g$ and where there is an edge from $scc_1$ to $scc_2$ if and only if there are two vertices $v_1$ and $v_2$ of $g$ such that $v_1 \in V\left( scc_1 \right)$, $v_2 \in V\left( scc_2 \right)$, and $\left( v_1,v_2 \right)\in E\left( g \right)$.
-Note that $\opn{sccs}\left( g \right)$ is necessarily a forest.
+
+Given a graph $g$, let $\hat{g}$ be the graph whose vertices are the strongly connected components of $g$ and where there is an edge from $g_1$ to $g_2$ if and only if there are two vertices $v_1$ and $v_2$ of $g$ such that $v_1 \in V\left( g_1 \right)$, $v_2 \in V\left( g_2 \right)$, and $\left( v_1,v_2 \right)\in E\left( g \right)$.
+Note that the reflexive and transitive closure of $V\left(  \hat{g} \right)$ is necessarily a partial order.
+For example, if $V\left( g
+\right)=\aset{c_1,c_2,c_3,c_4}$
+and $E\left( g \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$ then $V\left(\hat{g}\right) = \aset{\aset{c_1,c_2},\aset{c_3},\aset{c_4}}$ and $E\left( \hat{g} \right)=\aset{\left( \aset{c_1,c_2},\aset{c_3} \right)}$.
+
+A directed graph $g$ induces a partial order $\preceq_g$ on its vertices defined such that $v_1 \preceq_g v_2$ iff there is a path from $v_1$ to $v_2$ and there is no path from $v_2$ to $v_1$.
+Let $\opn{TOs}\left( g \right)$ be the set of total orders $\leq$ such that $\preceq_g \subseteq \leq$
+
+Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ cs \right]$ is a total order on the set $cs$.
+
+When $\hat{o}$ is a total order in $\opn{TOs}\left( \hat{g} \right)$, define the total order $\leq_{\hat{o}}^g$ on $V\left( g \right)$ such that $v_1 \leq_{\hat{o}}^g v_2$ iff (1) there exists $V_1,V_2\in V\left(\hat{g}\right)$ such that $v_1\in V_1$, $v_2\in V_2$, and $\left( V_1,V_2 \right)\in \hat{o}$, or (2) there exists $V\in V\left(\hat{g}\right)$ such that $v_1,v_2\in V\left( V \right)$ and $\left( v_1,v_2 \right)\in f\left[ V\left( V \right) \right]$.
+We say that $\leq_{\hat{o}}^g$ is a \emph{linearization} of $g$.
+
+Property: if $V\left( g_1 \right)\subseteq V\left( g_2 \right)$, $E\left( g_1 \right)\subseteq E\left( g_2 \right)$, $l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$, then $\bot\bullet l1 \preceq \bot\bullet l2$.
+
+\begin{comment}
+Define $\opn{Lin}_g\left( \leq \right) = f\left[ V_1 \right] \cdot F\left[ V_2 \right] \cdot F\left[ V_n \right]$, where $V_i\in V\left( \hat{g} \right)$
 
@@
 
+\end{comment}
+
 \subsection{Correctenss Proof}

### 2016-02-11 podc16/paper2.tex
dependency-set algo - first part

@@
 
+An example of implementation of such an algorithm is depicted by the pseudocode of \Cref{fig:deps-set-algo}.
+The idea is to let every process broadcast $c$ to all the processes in the system, in order to announce $c$.
+Then whenever a process $p$ receives $c$, it replies to the sender with the set $\mathcal{D}$ of commands that $p$ has observed so far, and such that $c$ and any command in $\mathcal{D}$ are \textit{non-commutative}.
+Afterwards, the process announcing $c$ collects the sets $\mathcal{D}$ for $c$ from a quorum of $\mathcal{Q}$ processes, and it locally suggests the union of those sets as final dependency-set for $c$.
+
+The provided implementation satisfies the two properties of a dependency-set algorithm for the following reasons.
+Since any command in a dependency-set that is suggested by a process must be in the set $cmds_q$ of the commands received by some process $q$, and a process receives a command if and only if there exists some process that previously announced the command, then it follows that any command in any suggested dependency-set has been previously announced.
+
+\begin{figure*}
+\centering
+\subfigure[$\texttt{init}_p$ and $\texttt{announce}_p$ implementations.]{
+\noindent\fbox{%
+	\begin{varwidth}{\dimexpr\linewidth-2\fboxsep-2\fboxrule\relax}
+
+		\begin{algorithmic}[1]
+			{\small
+			\START[\texttt{init$_p$}]{{\bf upon}}{$()$} \label{alg:init}
+			\State $cmds_p\gets\emptyset$ \label{alg:init-cmds}
+			\State $\forall c,$ $deps_p[c]\gets \emptyset$ \label{alg:init-deps}
+			\State $\forall c,$ $heard_p[c]\gets \emptyset$ \label{alg:init-heard}
+			\State $\forall c,$ $suggested_p[c]\gets false$ \label{alg:init-suggested}
+			\END
+			\Statex
+			\START[\texttt{announce$_p$}]{{\bf upon}}{$(c)$} \label{alg:announce}
+			\State $cmds_p\gets cmds_p\cup\{c\}$ \label{alg:add-command}
+			\State $\forall q\in P$ {\bf send} $\langle \texttt{cmd}$ $|$ $c \rangle$ {\bf to} $q$\label{alg:send-new-cmd}
+			\END
+			}
+		\end{algorithmic}
+	\end{varwidth}%
+}
+\label{fig:deps-set-algo1}
+}
+\hspace*{0.3in}\subfigure[receive of \texttt{cmd} and \texttt{deps} messages, and suggestion of dependencies for a command on process $p$.]{
+\noindent\fbox{%
+	\begin{varwidth}{\dimexpr\linewidth-2\fboxsep-2\fboxrule\relax}
+
+		\begin{algorithmic}[1]
+			{\small
+			\START[$\langle \texttt{cmd}$ $|$ $c \rangle$]{{\bf upon receive}}{ {\bf from} $q$} \label{alg:receive-new-cmd}
+			\State $\mathcal{D} \gets \{ d : d \in cmds_p \land d\neq c \land d \asymp c \}$
+			\State $cmds_p\gets cmds_p\cup\{c\}$
+			\State {\bf send} $\langle \texttt{deps}$ $|$ $c,\mathcal{D} \rangle$ {\bf to} $q$\label{alg:send-deps}
+			\END
+			\Statex
+			\START[$\langle \texttt{deps}$ $|$ $c,\mathcal{D} \rangle$]{{\bf upon receive}}{ {\bf from} $q$} \label{alg:receive-deps}
+			\State $deps_p[c]\gets deps_p[c]\cup \mathcal{D}$
+			\State $heard_p[c]\gets heard_p[c]\cup \{q\}$
+			\END
+			\Statex
+			\START[]{{\bf upon}}{$\exists c$ $:$ $ | heard_p[c] | \geq \mathcal{Q}$ $\land$ $suggested_p[c]=false$} \label{alg:receive-deps}
+			\State $suggested_p[c]\gets true$
+			\State {\bf call} $\texttt{suggest}_p(c,deps_p[c])$
+			\END
+			}
+		\end{algorithmic}
+	\end{varwidth}%
+}
+\label{fig:deps-set-algo2}
+}
+\caption{Example of dependency-set algorithm implementation.}
+\label{fig:deps-set-algo}
+\end{figure*}
+
 %\todo{Plain english version below. Which is better?}
@@
 \todo{``Suggest'' can also be used to trigger recovery}.
+
+\todo{Write a suggestion on about GC of dependencies. Remove any command c such that a committed command depends on c. Also us the GC of the map-agreement component.}
 
@@
 Note that any SMR algorithm, like MultiPaxos, can be used to implementing map-agreement by uniquely associating commands to MultiPaxos instance numbers.
+
+\todo{Cite Generalized Paxos as an example for bounding the data structures.}
 

### 2016-02-11 podc16/paper2.tex
work on graph linearization

--- 2016-02-11 podc16/paper2.tex
@@
 
-The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $\opn{cs}_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $v_2 \in \opn{cs}_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$).
+\textbf{Local dependency graph computed by a process.} The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $\opn{cs}_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $v_2 \in \opn{cs}_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$).
 For example, if $cs_p = \left[ c_1 \mapsto \aset{c_2,c_3}, c_2 \mapsto \aset{c_1} \right]$ then $V\left(
@@
 
-Given a graph $g$, let $\hat{g}$ be the graph whose vertices are the strongly connected components of $g$ and where there is an edge from $g_1$ to $g_2$ if and only if there are two vertices $v_1$ and $v_2$ of $g$ such that $v_1 \in V\left( g_1 \right)$, $v_2 \in V\left( g_2 \right)$, and $\left( v_1,v_2 \right)\in E\left( g \right)$.
+\textbf{Graph of strongly connected components.} Given a graph $g$, let $\hat{g}$ be the graph whose vertices are the strongly connected components of $g$ (represented as sets of vertices) and where there is an edge from $g_1$ to $g_2$ if and only if there are two vertices $v_1$ and $v_2$ of $g$ such that $v_1 \in V\left( g_1 \right)$, $v_2 \in V\left( g_2 \right)$, and $\left( v_1,v_2 \right)\in E\left( g \right)$.
 Note that the reflexive and transitive closure of $V\left(  \hat{g} \right)$ is necessarily a partial order.
@@
 
-A directed graph $g$ induces a partial order $\preceq_g$ on its vertices defined such that $v_1 \preceq_g v_2$ iff there is a path from $v_1$ to $v_2$ and there is no path from $v_2$ to $v_1$.
-Let $\opn{TOs}\left( g \right)$ be the set of total orders $\leq$ such that $\preceq_g \subseteq \leq$
+\textbf{Orders induced by a graph.} A directed graph $g$ induces a partial order $\preceq_g$ on its vertices defined such that $v_1 \preceq_g v_2$ iff there is a path from $v_1$ to $v_2$ and there is no path from $v_2$ to $v_1$.
+Let $\opn{TOs}\left( g \right)$ be the set of total orders $o$ such that $\preceq_g \subseteq o$.
 
-Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ cs \right]$ is a total order on the set $cs$.
+\textbf{Linearization function.} Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ cs \right]$ is a total order on the set $cs$.
 
-When $\hat{o}$ is a total order in $\opn{TOs}\left( \hat{g} \right)$, define the total order $\leq_{\hat{o}}^g$ on $V\left( g \right)$ such that $v_1 \leq_{\hat{o}}^g v_2$ iff (1) there exists $V_1,V_2\in V\left(\hat{g}\right)$ such that $v_1\in V_1$, $v_2\in V_2$, and $\left( V_1,V_2 \right)\in \hat{o}$, or (2) there exists $V\in V\left(\hat{g}\right)$ such that $v_1,v_2\in V\left( V \right)$ and $\left( v_1,v_2 \right)\in f\left[ V\left( V \right) \right]$.
-We say that $\leq_{\hat{o}}^g$ is a \emph{linearization} of $g$.
+\textbf{Linearizations of a dependency graph.}
+When $\hat{o}$ is a total order over the strongly connected components of $g$ such that $\hat{o} \in \opn{TOs}\left( \hat{g} \right)$, define the total order $l\left(\hat{o},g\right)$ on $V\left( g \right)$ such that $\left(v_1,v_2\right) \in l\left(\hat{o},g\right)$ iff (1) there exists $vs_1,vs_2\in V\left(\hat{g}\right)$ such that $v_1\in vs_1$, $v_2\in vs_2$, and $\left( vs_1,vs_2 \right)\in \hat{o}$, or (2) there exists $vs_1\in V\left(\hat{g}\right)$ such that
+$v_1,v_2\in  vs_1$ and $\left( v_1,v_2 \right)\in f\left[  vs_1 \right]$.
+We say that $l\left(\hat{o},g\right)$ is a \emph{linearization} of $g$.
 
-Property: if $V\left( g_1 \right)\subseteq V\left( g_2 \right)$, $E\left( g_1 \right)\subseteq E\left( g_2 \right)$, $l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$, then $\bot\bullet l1 \preceq \bot\bullet l2$.
+\textbf{Lemma}: 
+Assume that: $g_1$ and $g_2$ are two dependency graphs; 
+for every pair of non-commutative commands $c_1,c_2 \in V\left(g_2\right)$, either $\left( c_1,c_2 \right)\in E\left( g_2 \right)$ or $\left( c_2,c_1 \right)\in E\left( g_2 \right)$;
+$V\left( \hat{g_1} \right)\subseteq V\left( \hat{g_2} \right)$; 
+$E\left( \hat{g_1} \right)\subseteq E\left( \hat{g_2} \right)$; 
+$l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
+Then, $\bot\bullet l1 \preceq \bot\bullet l2$.
 

### 2016-02-11 podc16/paper2.tex
work on graph linearization

@@
 
-Local variables: $\opn{cs}_p$, a map from commands to dependency sets, initially the empty map.
+Local variables: $m_p$, a map from commands to dependency sets, initially the empty map; $\sigma_p$, a c-struct, initially $\bot$.
 \begin{itemize}[noitemsep,nolistsep]
@@
     \item Upon ${\opn{suggest}\left( c,d \right)}^p$: call ${\opn{propose}\left( c,d \right)}^p$.
-    \item Upon ${\opn{decide}\left( c,d \right)}^p$: insert the mapping $c \rightarrow d$ in the map $\opn{cs}_p$.
-    \item Periodically,  every process $p$ processes $\opn{cs}_p$, as described below.
+    \item Upon ${\opn{decide}\left( c,d \right)}^p$: insert the mapping $c \rightarrow d$ in the map $m_p$.
+    \item Periodically, update $\sigma_p$ to $\opn{c-struct}\left( m_p \right)$, where $\opn{c-struct}\left( m_p \right)$ is defined below.
 \end{itemize}
@@
 
-\textbf{Local dependency graph computed by a process.} The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $\opn{cs}_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $v_2 \in \opn{cs}_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$).
+\textbf{Local dependency graph computed by a process.} The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $v_2 \in m_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$).
 For example, if $cs_p = \left[ c_1 \mapsto \aset{c_2,c_3}, c_2 \mapsto \aset{c_1} \right]$ then $V\left(
@@
 \textbf{Linearizations of a dependency graph.}
-When $\hat{o}$ is a total order over the strongly connected components of $g$ such that $\hat{o} \in \opn{TOs}\left( \hat{g} \right)$, define the total order $l\left(\hat{o},g\right)$ on $V\left( g \right)$ such that $\left(v_1,v_2\right) \in l\left(\hat{o},g\right)$ iff (1) there exists $vs_1,vs_2\in V\left(\hat{g}\right)$ such that $v_1\in vs_1$, $v_2\in vs_2$, and $\left( vs_1,vs_2 \right)\in \hat{o}$, or (2) there exists $vs_1\in V\left(\hat{g}\right)$ such that
+When $\hat{o}$ is a total order over the strongly connected components of $g$ such that $\hat{o} \in \opn{TOs}\left( \hat{g} \right)$, we say that a total order $l$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when $\left(v_1,v_2\right) \in l$ iff (1) there exists $vs_1,vs_2\in V\left(\hat{g}\right)$ such that $v_1\in vs_1$, $v_2\in vs_2$, and $\left( vs_1,vs_2 \right)\in \hat{o}$, or (2) there exists $vs_1\in V\left(\hat{g}\right)$ such that
 $v_1,v_2\in  vs_1$ and $\left( v_1,v_2 \right)\in f\left[  vs_1 \right]$.
-We say that $l\left(\hat{o},g\right)$ is a \emph{linearization} of $g$.
 
-\textbf{Lemma}: 
+\textbf{Lemma 1}: 
 Assume that: $g_1$ and $g_2$ are two dependency graphs; 
@@
 $l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
-Then, $\bot\bullet l1 \preceq \bot\bullet l2$.
+Then we have (a), $\bot\bullet l1 \preceq \bot\bullet l2$, and (b), if $g_1 = g_2$ then $\bot\bullet l1 = \bot\bullet l2$.
+
+\textbf{Definition of $\opn{c-struct}\left( m_p \right)$}.
+We define $\opn{c-struct}\left( m_p \right) = \bot\bullet l$ where $l$ is a linearization of $g_p$.
+By lemma 1.b, $\opn{c-struct}\left( m_p \right)$ is well-defined.
 

### 2016-02-11 podc16/paper2.tex
work on graph linearization, added introductory paragraph

@@
 
+The graph processing algorithm must compute $\opn{c-struct}\left( g_p \right)$, as defined below.
+Its main property is Lemma 1; Together with Invariants 1 and 2 (TODO, the invariants guaranteed by the dependency-set algorithm and the map-agreement algorithm), Lemma 1 implies that leaderless algorithms safely implement generalized consensus.
+
 \textbf{Local dependency graph computed by a process.} The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $v_2 \in m_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$).

### 2016-02-11 podc16/paper2.tex
work on graph linearization, added examples

@@
 Note that the reflexive and transitive closure of $V\left(  \hat{g} \right)$ is necessarily a partial order.
-For example, if $V\left( g
-\right)=\aset{c_1,c_2,c_3,c_4}$
-and $E\left( g \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$ then $V\left(\hat{g}\right) = \aset{\aset{c_1,c_2},\aset{c_3},\aset{c_4}}$ and $E\left( \hat{g} \right)=\aset{\left( \aset{c_1,c_2},\aset{c_3} \right)}$.
+For example, if $V\left( g \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( g \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$ then $V\left(\hat{g}\right) = \aset{\aset{c_1,c_2},\aset{c_3},\aset{c_4}}$ and $E\left( \hat{g} \right)=\aset{\left( \aset{c_1,c_2},\aset{c_3} \right)}$.
 
@@
 Let $\opn{TOs}\left( g \right)$ be the set of total orders $o$ such that $\preceq_g \subseteq o$.
+For example, if $V\left( g \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( g \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$, then $\preceq_g = \aset{\left( c_1,c_3 \right)}$ and $\opn{TOs}\left( g \right)$ is the set of all permutations of $\aseq{c_1,c_2,c_3,c_4}$ in which $c_1$ comes before $c_3$.
 
@@
 $v_1,v_2\in  vs_1$ and $\left( v_1,v_2 \right)\in f\left[  vs_1 \right]$.
+For example, if $V\left( g \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( g \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$ and $f\left[ \set{c_1,c_2} \right] = \aseq{c_2,c_1}$ then $\aseq{c_2,c_1,c_3,c_4}$ is a linearization of $g$ but $\aseq{c_1,c_2,c_3,c_4}$ is not and neither is $\aseq{c_3,c_2,c_1,c_4}$
 

### 2016-02-11 podc16/paper2.tex
fixed an error in the explanation of graph processing

@@
 
+\textbf{Partial order on graphs.}
+If $g_1$ and $g_2$ are two graphs, then we write $g_1 \leq g_2$ iff $V\left( g_1 \right)\subseteq V\left( g_2 \right)$, and if $v\in V\left( g_1 \right)$, $e\in E\left( g_2 \right)$, and $v$ is an endpoint of $e$, then $e\in E\left( g_1 \right)$.
+
 \textbf{Lemma 1}:
@@
 for every pair of non-commutative commands $c_1,c_2 \in V\left(g_2\right)$, either $\left( c_1,c_2 \right)\in E\left( g_2 \right)$ or $\left( c_2,c_1 \right)\in E\left( g_2 \right)$;
-$V\left( \hat{g_1} \right)\subseteq V\left( \hat{g_2} \right)$; 
-$E\left( \hat{g_1} \right)\subseteq E\left( \hat{g_2} \right)$; 
+$g_1 \leq g_2$;
 $l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
-Then we have (a), $\bot\bullet l1 \preceq \bot\bullet l2$, and (b), if $g_1 = g_2$ then $\bot\bullet l1 = \bot\bullet l2$.
+Then we have (a), $\bot\bullet l_1 \preceq \bot\bullet l_2$, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.
 

### 2016-02-11 podc16/paper2.tex
examples in the graph processing section

@@
 Note that the reflexive and transitive closure of $V\left(  \hat{g} \right)$ is necessarily a partial order.
-For example, if $V\left( g \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( g \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$ then $V\left(\hat{g}\right) = \aset{\aset{c_1,c_2},\aset{c_3},\aset{c_4}}$ and $E\left( \hat{g} \right)=\aset{\left( \aset{c_1,c_2},\aset{c_3} \right)}$.
+For example, consider $h$ such that $V\left( h \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$.
+Then $V\left(\hat{h}\right) = \aset{\aset{c_1,c_2},\aset{c_3},\aset{c_4}}$ and $E\left( \hat{h} \right)=\aset{\left( \aset{c_1,c_2},\aset{c_3} \right)}$.
+We will use $h$ as an example in the rest of the section.
 
@@
 Let $\opn{TOs}\left( g \right)$ be the set of total orders $o$ such that $\preceq_g \subseteq o$.
-For example, if $V\left( g \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( g \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$, then $\preceq_g = \aset{\left( c_1,c_3 \right)}$ and $\opn{TOs}\left( g \right)$ is the set of all permutations of $\aseq{c_1,c_2,c_3,c_4}$ in which $c_1$ comes before $c_3$.
+For example, $\preceq_h = \aset{\left( c_1,c_3 \right)}$ and $\opn{TOs}\left( h \right)$ is the set of all permutations of $\aseq{c_1,c_2,c_3,c_4}$ in which $c_1$ comes before $c_3$.
 
@@
 $v_1,v_2\in  vs_1$ and $\left( v_1,v_2 \right)\in f\left[  vs_1 \right]$.
-For example, if $V\left( g \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( g \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$ and $f\left[ \set{c_1,c_2} \right] = \aseq{c_2,c_1}$ then $\aseq{c_2,c_1,c_3,c_4}$ is a linearization of $g$ but $\aseq{c_1,c_2,c_3,c_4}$ is not and neither is $\aseq{c_3,c_2,c_1,c_4}$
+For example, if $f\left[ \set{c_1,c_2} \right] = \aseq{c_2,c_1}$ then $\aseq{c_2,c_1,c_3,c_4}$ is a linearization of $h$ but $\aseq{c_1,c_2,c_3,c_4}$ is not and neither is $\aseq{c_3,c_2,c_1,c_4}$
 
@@
 If $g_1$ and $g_2$ are two graphs, then we write $g_1 \leq g_2$ iff $V\left( g_1 \right)\subseteq V\left( g_2 \right)$, and if $v\in V\left( g_1 \right)$, $e\in E\left( g_2 \right)$, and $v$ is an endpoint of $e$, then $e\in E\left( g_1 \right)$.
+For example, if $V\left( h' \right)=\aset{c_1,c_2,c_4}$ and $E\left( h' \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right)}$ then $h'\leq h$.
+However, if $V\left( h'' \right)=\aset{c_1,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_3 \right)}$ then $\left( h'',h \right)\notin \leq$.
 

### 2016-02-11 podc16/paper2.tex
making space

@@
 
-%\title{Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms}
+\title{Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms}
 
-%\author{}
+\author{
+Author 1\\
+ECE, Virginia Tech\\
+email
+\and
+Author 2\\
+ECE, Virginia Tech\\
+email
+\and
+Author 3\\
+ECE, Virginia Tech\\
+email
+}
 
-%\date{}
+\date{}
 
@@
 
-%\maketitle
+\maketitle
 
-\input{cover}
+%\input{cover}
 
@@
 
-Our work sheds light on the common structure of algorithms like EPaxos and Alvin and allows to tune the performance of leaderless algorithms for specific scenarios by changing the underlying generalized-consensus and dependency-set algorithm.
+Our work sheds light on the common structure of algorithms like EPaxos and Alvin and allows to tune the performance of leaderless algorithms for specific scenarios by changing the underlying map-agreement and dependency-set algorithms.
 We expect that these two tunable knobs will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by this new family of algorithms.
@@
 Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}.
-Each process receives proposals for commands of the form $\opn{Propose}(c)$, and it produces responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct.
-A c-struct set is a set, containing the element $\bot$, with an append operator $\bullet$ such that $\sigma \bullet c$ is a c-struct, for any c-struct $\sigma$ and command $c$.
+Each process receives proposals for commands of the form $\opn{Propose}(c)$, and must produce responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct.
+The sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties described below. But first, let us define c-structs and their operations.
+
+%\begin{comment}
+A c-struct set is a set containing the element $\bot$ and with an append operator $\bullet$ such that $\sigma \bullet c$ is a c-struct, for any c-struct $\sigma$ and command $c$. 
 We also write the $\sigma\bullet\aseq{c_1,\cdots,c_n}$ for the c-struct $\sigma\bullet c_1\bullet\cdots\bullet c_n$.
-C-structs are partially ordered by the relation $\preceq$ defined such that $\sigma_1\preceq\sigma_2$ iff there exists a sequence of commands $cs$ such that $\sigma_2=\sigma_1\bullet cs$.
+C-structs are partially ordered by the relation $\sqsubseteq$ defined such that $\sigma_1\sqsubseteq\sigma_2$ iff there exists a sequence of commands $cs$ such that $\sigma_2=\sigma_1\bullet cs$.
+Two c-structs are compatible when they have a common upper bound; when this is the case, two c-structs $\sigma_1$ and $\sigma_2$ have a least upper bound, noted $\bigsqcup\left( \sigma_1,\sigma_2 \right)$.
 We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
@@
 In the rest of the paper the symbol $c$, possibly sub-scripted, ranges over commands, and the symbol $\mathcal{D}$, possibly sub-scripted, ranges over sets of commands.
-\todo{Present generalized consensus here?}
+
+We can now state the generalized-consensus properties: non-triviality requires that any decided c-struct be of the form $\bot\bullet cs$, where $cs$ is a sequence of proposed commands; consistency requires than any two decided c-structs $\sigma_1$ and $\sigma_2$ be compatible; stability requires that when a process $p$ decides a c-struct $\sigma_1$ and time $t_1$ and $\sigma_2$ at time $t_2$, then $t_1 \leq t_2$ implies that $\sigma_1\sqsubseteq \sigma_2$; finally, liveness requires that
+if a command keeps beeing proposed, then a c-struct $\sigma$ containing the command is eventually decided.
+
+%\end{comment}
 
@@
 
-\todo{``Suggest'' can also be used to trigger recovery}.
-
+\todo{``Suggest'' can also be used to trigger recovery}
 \todo{Write a suggestion on about GC of dependencies. Remove any command c such that a committed command depends on c. Also us the GC of the map-agreement component.}
@@
 Moreover, a map-agreement algorithm must ensure that: 
+\begin{comment}
 \begin{enumerate}[noitemsep,nolistsep]
@@
 \todo{OR MORE FORMALLY}
+
+\end{comment}
 

### 2016-02-11 podc16/paper2.tex
typos

@@
 
-The graph processing algorithm must compute $\opn{c-struct}\left( g_p \right)$, as defined below.
+The graph processing algorithm is executed locally by a node. It must compute $\opn{c-struct}\left( g_p \right)$, as defined below.
 Its main property is Lemma 1; Together with Invariants 1 and 2 (TODO, the invariants guaranteed by the dependency-set algorithm and the map-agreement algorithm), Lemma 1 implies that leaderless algorithms safely implement generalized consensus.
+To execute the graph processing
 
@@
 For example, consider $h$ such that $V\left( h \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$.
-Then $V\left(\hat{h}\right) = \aset{\aset{c_1,c_2},\aset{c_3},\aset{c_4}}$ and $E\left( \hat{h} \right)=\aset{\left( \aset{c_1,c_2},\aset{c_3} \right)}$.
+Then $V(\hat{h}) = \aset{\aset{c_1,c_2},\aset{c_3},\aset{c_4}}$ and $E( \hat{h} )=\aset{\left( \aset{c_1,c_2},\aset{c_3} \right)}$.
 We will use $h$ as an example in the rest of the section.
@@
 $v_1,v_2\in  vs_1$ and $\left( v_1,v_2 \right)\in f\left[  vs_1 \right]$.
-For example, if $f\left[ \set{c_1,c_2} \right] = \aseq{c_2,c_1}$ then $\aseq{c_2,c_1,c_3,c_4}$ is a linearization of $h$ but $\aseq{c_1,c_2,c_3,c_4}$ is not and neither is $\aseq{c_3,c_2,c_1,c_4}$
+For example, if $f\left[ \aset{c_1,c_2} \right] = \aseq{c_2,c_1}$ then $\aseq{c_2,c_1,c_3,c_4}$ is a linearization of $h$ but $\aseq{c_1,c_2,c_3,c_4}$ is not and neither is $\aseq{c_3,c_2,c_1,c_4}$
 
@@
 For example, if $V\left( h' \right)=\aset{c_1,c_2,c_4}$ and $E\left( h' \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right)}$ then $h'\leq h$.
-However, if $V\left( h'' \right)=\aset{c_1,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_3 \right)}$ then $\left( h'',h \right)\notin \leq$.
+However, if $V\left( h'' \right)=\aset{c_1,c_3,c_4}$ and $E\left( h'' \right)=\aset{\left( c_1,c_3 \right)}$ then $\left( h'',h \right)\notin \leq$.
 

### 2016-02-11 podc16/paper2.tex
shortened the graph processing section

@@
 Its main property is Lemma 1; Together with Invariants 1 and 2 (TODO, the invariants guaranteed by the dependency-set algorithm and the map-agreement algorithm), Lemma 1 implies that leaderless algorithms safely implement generalized consensus.
-To execute the graph processing
+We assume that processes initially agree on a \emph{linearization function} $f$ mapping sets of commands to sequences of commands such that $f\left[ cs \right]$ is a total order on the set $cs$.
 
-\textbf{Local dependency graph computed by a process.} The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $v_2 \in m_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$).
+%\textbf{Local dependency graph computed by a process.} 
+The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $v_2 \in m_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$).
 For example, if $cs_p = \left[ c_1 \mapsto \aset{c_2,c_3}, c_2 \mapsto \aset{c_1} \right]$ then $V\left(
 g_p \right) = \aset{c_1,c_2}$ and $E\left( g_p \right) = \aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right)}$.
+
+%\textbf{Orders induced by a graph.}
+A directed graph $g$ \emph{induces a partial order} $\preceq_g$ on its vertices defined such that $v_1 \preceq_g v_2$ iff there is a path from $v_1$ to $v_2$ and there is no path from $v_2$ to $v_1$.
+%Let $\opn{TOs}\left( g \right)$ be the set of total orders $o$ such that $\preceq_g \subseteq o$.
+For example, $\preceq_h = \aset{\left( c_1,c_3 \right)}$. % and $\opn{TOs}\left( h \right)$ is the set of all permutations of $\aseq{c_1,c_2,c_3,c_4}$ in which $c_1$ comes before $c_3$.
+%\textbf{Equivalence relation induced by a graph.}
+A directed graph $g$ \emph{induces an equivalence relation} $\sim_g$ on its vertices defined such that $v_1 \sim_g v_2$ iff $v_1$ and $v_2$ belong to the same strongly connected component.
+%\textbf{Linearizations of a dependency graph.}
+We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $v_1,v_2\in V\left( g \right)$, $v_1 \preceq_g v_2$ implies $v_1 \ll v_2$, and if $v_1 \sim_g v_2$ and $\left( v_1,v_2 \right)\in f$ hold then $v_1 \ll v_2$.
+%\textbf{Partial order on graphs.}
+We now define a partial order on graphs.
+If $g_1$ and $g_2$ are two graphs, then we write $g_1 \leq g_2$ iff $V\left( g_1 \right)\subseteq V\left( g_2 \right)$, and if $v\in V\left( g_1 \right)$, $e\in E\left( g_2 \right)$, and $v$ is an endpoint of $e$, then $e\in E\left( g_1 \right)$.
+For example, consider the graphs $h'$ and $h''$ such that $V\left( h' \right)=\aset{c_1,c_2,c_4}$, $E\left( h' \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right)}$, $V\left( h'' \right)=\aset{c_1,c_3,c_4}$, and $E\left( h'' \right)=\aset{\left( c_1,c_3 \right)}$.
+We have that $h'\leq h$, but $ h''\not\leq h$.
+
+\textbf{Lemma 1}
+Assume that: $g_1$ and $g_2$ are two dependency graphs; 
+for every pair of non-commutative commands $c_1,c_2 \in V\left(g_2\right)$, either $\left( c_1,c_2 \right)\in E\left( g_2 \right)$ or $\left( c_2,c_1 \right)\in E\left( g_2 \right)$;
+$g_1 \leq g_2$;
+$l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
+Then we have (a), $\bot\bullet l_1 \preceq \bot\bullet l_2$, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.
+
+\textbf{Definition 1}
+We define $\opn{c-struct}\left( m_p \right) = \bot\bullet l$ where $l$ is a linearization of $g_p$.
+By lemma 1.b, $\opn{c-struct}\left( m_p \right)$ is well-defined.
+
+\begin{comment}
 
@@
 We will use $h$ as an example in the rest of the section.
-
-\textbf{Orders induced by a graph.} A directed graph $g$ induces a partial order $\preceq_g$ on its vertices defined such that $v_1 \preceq_g v_2$ iff there is a path from $v_1$ to $v_2$ and there is no path from $v_2$ to $v_1$.
-Let $\opn{TOs}\left( g \right)$ be the set of total orders $o$ such that $\preceq_g \subseteq o$.
-For example, $\preceq_h = \aset{\left( c_1,c_3 \right)}$ and $\opn{TOs}\left( h \right)$ is the set of all permutations of $\aseq{c_1,c_2,c_3,c_4}$ in which $c_1$ comes before $c_3$.
 
@@
 
-\textbf{Partial order on graphs.}
-If $g_1$ and $g_2$ are two graphs, then we write $g_1 \leq g_2$ iff $V\left( g_1 \right)\subseteq V\left( g_2 \right)$, and if $v\in V\left( g_1 \right)$, $e\in E\left( g_2 \right)$, and $v$ is an endpoint of $e$, then $e\in E\left( g_1 \right)$.
-For example, if $V\left( h' \right)=\aset{c_1,c_2,c_4}$ and $E\left( h' \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right)}$ then $h'\leq h$.
-However, if $V\left( h'' \right)=\aset{c_1,c_3,c_4}$ and $E\left( h'' \right)=\aset{\left( c_1,c_3 \right)}$ then $\left( h'',h \right)\notin \leq$.
-
-\textbf{Lemma 1}:
-Assume that: $g_1$ and $g_2$ are two dependency graphs; 
-for every pair of non-commutative commands $c_1,c_2 \in V\left(g_2\right)$, either $\left( c_1,c_2 \right)\in E\left( g_2 \right)$ or $\left( c_2,c_1 \right)\in E\left( g_2 \right)$;
-$g_1 \leq g_2$;
-$l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
-Then we have (a), $\bot\bullet l_1 \preceq \bot\bullet l_2$, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.
-
-\textbf{Definition of $\opn{c-struct}\left( m_p \right)$}.
-We define $\opn{c-struct}\left( m_p \right) = \bot\bullet l$ where $l$ is a linearization of $g_p$.
-By lemma 1.b, $\opn{c-struct}\left( m_p \right)$ is well-defined.
+\end{comment}
 

### 2016-02-11 podc16/paper2.tex
margins
fixes in the graph processing section

@@
 
-The graph processing algorithm is executed locally by a node. It must compute $\opn{c-struct}\left( g_p \right)$, as defined below.
-Its main property is Lemma 1; Together with Invariants 1 and 2 (TODO, the invariants guaranteed by the dependency-set algorithm and the map-agreement algorithm), Lemma 1 implies that leaderless algorithms safely implement generalized consensus.
-We assume that processes initially agree on a \emph{linearization function} $f$ mapping sets of commands to sequences of commands such that $f\left[ cs \right]$ is a total order on the set $cs$.
+The graph processing algorithm is executed locally by a node. It must compute the value of $\opn{c-struct}\left( g_p \right)$, as defined below.
+Its main property is Lemma 1; together with Invariants 1 and 2 (TODO, the invariants guaranteed by the dependency-set algorithm and the map-agreement algorithm), Lemma 1 implies that leaderless algorithms safely implement generalized consensus.
+We assume that processes initially agree, for each set of commands $cs$, on a total order $<_{cs}$ on $cs$.
+In practice, if commands are attached a unique identifier among a totally ordered set, $<_{cs}$ can be the restriction of this total order on $cs$.
 
@@
 %\textbf{Linearizations of a dependency graph.}
-We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $v_1,v_2\in V\left( g \right)$, $v_1 \preceq_g v_2$ implies $v_1 \ll v_2$, and if $v_1 \sim_g v_2$ and $\left( v_1,v_2 \right)\in f$ hold then $v_1 \ll v_2$.
+We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $v_1,v_2\in V\left( g \right)$, $v_1 \preceq_g v_2$ implies $v_1 \ll v_2$, and if $v_1 \sim_g v_2$ and $\left( v_1,v_2 \right)\in f\left( \left[ v_1 \right] \right)$ hold then $v_1 \ll v_2$.
 %\textbf{Partial order on graphs.}

### 2016-02-11 podc16/paper2.tex
fixes in the graph processing section

@@
 
-\title{Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms}
+\title{\vspace{-0.8in}Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms}
 
@@
 A directed graph $g$ \emph{induces an equivalence relation} $\sim_g$ on its vertices defined such that $v_1 \sim_g v_2$ iff $v_1$ and $v_2$ belong to the same strongly connected component.
+We write $\left[ v \right]_g$ for the equivalence class of $v$ under $\sim_g$.
 %\textbf{Linearizations of a dependency graph.}
-We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $v_1,v_2\in V\left( g \right)$, $v_1 \preceq_g v_2$ implies $v_1 \ll v_2$, and if $v_1 \sim_g v_2$ and $\left( v_1,v_2 \right)\in f\left( \left[ v_1 \right] \right)$ hold then $v_1 \ll v_2$.
+We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $v_1,v_2\in V\left( g \right)$, $v_1 \preceq_g v_2$ implies $v_1 \ll v_2$, and if $v_1 \sim_g v_2$ and $ v_1 <_{\left[ v_1 \right]} v_2$ hold then $v_1 \ll v_2$.
 %\textbf{Partial order on graphs.}
+
 We now define a partial order on graphs.

### 2016-02-11 podc16/paper2.tex
typo in section on dependency-set algo

--- 2016-02-11 podc16/paper2.tex
@@
 The idea is to let every process broadcast $c$ to all the processes in the system, in order to announce $c$.
-Then whenever a process $p$ receives $c$, it replies to the sender with the set $\mathcal{D}$ of commands that $p$ has observed so far, and such that $c$ and any command in $\mathcal{D}$ are \textit{non-commutative}.
+Then whenever a process $p$ receives $c$, it replies to the sender with the set $\mathcal{D}$ of commands that $p$ has observed so far, and such that $c$ and any command in $\mathcal{D}$ do not commute.
 Afterwards, the process announcing $c$ collects the sets $\mathcal{D}$ for $c$ from a quorum of $\mathcal{Q}$ processes, and it locally suggests the union of those sets as final dependency-set for $c$.
 
-The provided implementation satisfies the two properties of a dependency-set algorithm for the following reasons.
+The provided implementation satisfies the specification of a dependency-set algorithm for the following reasons.
 Since any command in a dependency-set that is suggested by a process must be in the set $cmds_q$ of the commands received by some process $q$, and a process receives a command if and only if there exists some process that previously announced the command, then it follows that any command in any suggested dependency-set has been previously announced.

### 2016-02-11 podc16/paper2.tex
fixes in the graph processing section, again

--- 2016-02-11 podc16/paper2.tex
@@
 %\textbf{Linearizations of a dependency graph.}
-We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $v_1,v_2\in V\left( g \right)$, $v_1 \preceq_g v_2$ implies $v_1 \ll v_2$, and if $v_1 \sim_g v_2$ and $ v_1 <_{\left[ v_1 \right]} v_2$ hold then $v_1 \ll v_2$.
+We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $v_1,v_2\in V\left( g \right)$, $v_1 \preceq_g v_2$ implies $v_1 \ll v_2$, and if $v_1 \sim_g v_2$ and $ v_1 <_{\left[ v_1 \right]_g} v_2$ hold then $v_1 \ll v_2$.
 %\textbf{Partial order on graphs.}
@@
 $l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
-Then we have (a), $\bot\bullet l_1 \preceq \bot\bullet l_2$, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.
+Then we have (a), $\bot\bullet l_1 \sqsubseteq \bot\bullet l_2$, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.
 

### 2016-02-11 podc16/paper2.tex
fixes in the graph processing section, again again

@@
 %\textbf{Local dependency graph computed by a process.} 
-The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $v_2 \in m_p\left[ v_1 \right]$ (i.e., $v_1$ depends on $v_2$).
+The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $c_2 \in m_p\left[ c_1 \right]$ (i.e., $c_1$ depends on $c_2$).
 For example, if $cs_p = \left[ c_1 \mapsto \aset{c_2,c_3}, c_2 \mapsto \aset{c_1} \right]$ then $V\left(
@@
 %\textbf{Orders induced by a graph.}
-A directed graph $g$ \emph{induces a partial order} $\preceq_g$ on its vertices defined such that $v_1 \preceq_g v_2$ iff there is a path from $v_1$ to $v_2$ and there is no path from $v_2$ to $v_1$.
+A directed graph $g$ \emph{induces a partial order} $\preceq_g$ on its vertices defined such that $c_1 \preceq_g c_2$ iff there is a path from $c_1$ to $c_2$ and there is no path from $c_2$ to $c_1$.
 %Let $\opn{TOs}\left( g \right)$ be the set of total orders $o$ such that $\preceq_g \subseteq o$.
@@
 %\textbf{Equivalence relation induced by a graph.}
-A directed graph $g$ \emph{induces an equivalence relation} $\sim_g$ on its vertices defined such that $v_1 \sim_g v_2$ iff $v_1$ and $v_2$ belong to the same strongly connected component.
+A directed graph $g$ \emph{induces an equivalence relation} $\sim_g$ on its vertices defined such that $c_1 \sim_g c_2$ iff $c_1$ and $c_2$ belong to the same strongly connected component.
 We write $\left[ v \right]_g$ for the equivalence class of $v$ under $\sim_g$.
 %\textbf{Linearizations of a dependency graph.}
-We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $v_1,v_2\in V\left( g \right)$, $v_1 \preceq_g v_2$ implies $v_1 \ll v_2$, and if $v_1 \sim_g v_2$ and $ v_1 <_{\left[ v_1 \right]_g} v_2$ hold then $v_1 \ll v_2$.
+We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1 \sim_g c_2$ and $ c_1 <_{\left[ c_1 \right]_g} c_2$ hold then $c_1 \ll c_2$.
 %\textbf{Partial order on graphs.}

### 2016-02-11 podc16/paper2.tex
removed the def of generalized consensus and refered to Lamport instead
shortened the graph processing section a bit more

@@
 Each process receives proposals for commands of the form $\opn{Propose}(c)$, and must produce responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct.
-The sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties described below. But first, let us define c-structs and their operations.
+The sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties described below.
+We refer the reader to~\cite{Lamport05GeneralizeConsensus} for the definition of c-structs and generalized consensus. Note that we use exactly the same notation.
 
-%\begin{comment}
+\begin{comment}
+
 A c-struct set is a set containing the element $\bot$ and with an append operator $\bullet$ such that $\sigma \bullet c$ is a c-struct, for any c-struct $\sigma$ and command $c$. 
@@
 
-%\end{comment}
+\end{comment}
 
@@
 For example, $\preceq_h = \aset{\left( c_1,c_3 \right)}$. % and $\opn{TOs}\left( h \right)$ is the set of all permutations of $\aseq{c_1,c_2,c_3,c_4}$ in which $c_1$ comes before $c_3$.
+If $c\in V\left( g \right)$, we write $\left[ c \right]_g$ for the strongly connected component of $g$ to which $c$ belongs. 
 %\textbf{Equivalence relation induced by a graph.}
-A directed graph $g$ \emph{induces an equivalence relation} $\sim_g$ on its vertices defined such that $c_1 \sim_g c_2$ iff $c_1$ and $c_2$ belong to the same strongly connected component.
+%A directed graph $g$ \emph{induces an equivalence relation} $\sim_g$ on its vertices defined such that $c_1 \sim_g c_2$ iff $c_1$ and $c_2$ belong to the same strongly connected component.
 We write $\left[ v \right]_g$ for the equivalence class of $v$ under $\sim_g$.

### 2016-02-11 podc16/paper2.tex
fixes in the graph processing section, examples were missing
changes in the intro and minor other things

@@
 
-Many agreement algorithms inspired by Paxos and MultiPaxos~\cite{lamport2001paxos} rely on the existence of a unique leader replica which enforces an ordering on commands during fault-free periods.
+Many agreement algorithms inspired by Paxos and MultiPaxos~\cite{lamport2001paxos} rely on the existence of a unique leader process which enforces an ordering on commands during fault-free periods.
 In practice, the unique leader can easily become a performance bottleneck for the system.
 Therefore, recently proposed agreement algorithms like Mencius~\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine}, EPaxos~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments}, or Alvin~\cite{TurcuETAL14BeGeneralDontGiveUpConsistency} strive to increase performance by avoiding the use of a unique leader.
-Mencius rotates the leader among the replicas but its structure remains close to MultiPaxos. 
+Mencius rotates the leader among the processes but its structure remains close to MultiPaxos. 
 However, EPaxos and Alvin are based on a novel algorithmic idea first introduced in EPaxos.
 
-In this paper, we show that the core algorithmic idea underlying EPaxos can be used to build a family of leaderless algorithms parameterized by (1) a per-command dependency-set computation algorithm, a concept that we define in the paper, and (2) an algorithm ensuring agreement on a mapping from commands to dependency sets.
-%(2) an algorithm ensuring a per-command agreement on a dependency set and which is oblivious of any ordering between commands.
-An algorithm in the family is obtained by instantiating the two parameters to a concrete agreement algorithm and a concrete dependency-set computation algorithm, which are then used as sub-routines. 
-Note that any generalized consensus algorithm can be used to instantiate parameter 2, but that only a weaker problem than generalized consensus needs to be solved because no ordering constraints need to be enforced.
+We show that the core idea underlying EPaxos can be captured in an abstract algorithm that uses two abstractions: (1), a per-command dependency-set computation algorithm, a concept that we define in the paper, and (2), an algorithm ensuring, for each command proposed to it, agreement on a dependency set.
+Both abstractions lend themselves well to implementations that avoid the use of a unique leader process.
+
+Our abstract algorithm gives rise to a family of algorithms whose members are obtained by using a concrete map-agreement algorithm and a concrete dependency-set computation algorithm. 
+Note that any generalized consensus algorithm can be used to implement (2), but only a weaker problem than generalized consensus needs to be solved because no ordering among proposal is required.
+
+Our work has two main advantages: it enables modular correctness proofs of algorithms like EPaxos and it allows to tune the performance of implementations by changing the underlying map-agreement and dependency-set algorithms.
+We expect that these two tunable knobs will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by this new family of algorithms.
 
 We prove that all the algorithms in the family are safe and live, and we have formalized our ideas in TLA+ and thoroughly model-check our claims with the TLC model-checker.
-
-Our work sheds light on the common structure of algorithms like EPaxos and Alvin and allows to tune the performance of leaderless algorithms for specific scenarios by changing the underlying map-agreement and dependency-set algorithms.
-We expect that these two tunable knobs will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by this new family of algorithms.
 
@@
 Second, the process $p$ proposes the mapping  $c\mapsto cs$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto cs'$ for $c$, in which case $cs'$ becomes the set of dependencies of $c$, and we say that $c$ is executable when its dependency set is empty or when all its dependencies are executable.
-Finally, when $c$ is executable, the process $p$ may execute $c$ by deciding the c-struct obtained by running the \emph{graph-processing} algorithm described in \Cref{graph-proc}.
-Any other process can run the three phases independently of $p$ and eventually execute the command.
-However, we show that a leaderless  generalized consensus algorithm is always safe and is live when at least a majority of the processes are correct.
+Finally, when $c$ is executable, the process $p$ may execute $c$ by deciding the c-struct obtained by running the \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
+Any other process can run the three phases independently of $p$ and eventually execute the command despite a crash of $p$.
+However, we show that a leaderless  generalized consensus algorithm is always safe and is live when its underlying dependency-set and map-agreement algorithms are safe and live.
 
@@
 \label{dep-algo}
+
+\todo{Why not add the decide call to the interface, in order to guarantee liveness?}
 
@@
 
-\subsection{Implementing Leaderless Generalized Consensus}
-
-Genearalized consensus~\cite{Lamport05GeneralizeConsensus} is a specification for state-machine replication algorithms. 
-A generalized-consensus algorithm exposes the input actions $propose\left( c \right)$, for a command $c$, and $learn\left( \sigma \right)$, for a c-struct $\sigma$.
-Moreover, a generalized-consensus algorithm must ensure that TODO.
-
-We now propose an algorithm that uses the dependency-set computation abstraction and the map-agreement abstraction to implement generalized consensus.
-We therefore obtain a family of generalized consensus algorithms where each member of the family is obtained by using concrete dependency-set and map-agreement algorithms.
-
-\todo{Explain the algorithm with pseudocode?}
-
+\subsection{The Abstract Generalized Consensus Algorithm}
+\label{abstract-algo}
+We now propose describe our abstract generalized consensus algorithm. 
 Local variables: $m_p$, a map from commands to dependency sets, initially the empty map; $\sigma_p$, a c-struct, initially $\bot$.
@@
 \end{itemize}
-
-\subsection{Graph Processing}
-\label{graph-proc}
 
@@
 %Let $\opn{TOs}\left( g \right)$ be the set of total orders $o$ such that $\preceq_g \subseteq o$.
-For example, $\preceq_h = \aset{\left( c_1,c_3 \right)}$. % and $\opn{TOs}\left( h \right)$ is the set of all permutations of $\aseq{c_1,c_2,c_3,c_4}$ in which $c_1$ comes before $c_3$.
+For example, consider $h$ such that $V\left( h \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$. 
+We have that $\preceq_h = \aset{\left( c_1,c_3 \right)}$. % and $\opn{TOs}\left( h \right)$ is the set of all permutations of $\aseq{c_1,c_2,c_3,c_4}$ in which $c_1$ comes before $c_3$.
 If $c\in V\left( g \right)$, we write $\left[ c \right]_g$ for the strongly connected component of $g$ to which $c$ belongs. 
@@
 We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1 \sim_g c_2$ and $ c_1 <_{\left[ c_1 \right]_g} c_2$ hold then $c_1 \ll c_2$.
+For example, assuming that $c_2 \leq_{\aset{c_1,c_2}} c_1$, the linearizations of $h$ are $\aseq{c_2,c_1,c_3,c_4}$, $\aseq{c_2,c_1,c_4,c_3}$, $\aseq{c_2,c_4,c_1,c_3}$, and $\aseq{c_4,c_2,c_1,c_3}$.
 %\textbf{Partial order on graphs.}
@@
 
-\subsection{Correctenss Proof}
+\subsection{Correctness Proof}
 

### 2016-02-11 podc16/paper2.tex
added a decide input to the dependency-set algorithm

@@
 
-A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, and the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
+A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{decide}\left( c,\mathcal{D} \right)$ (meant to receive inputs from the agreement sub-algorithm), where $p\in P$ is the process executing the call.
 Moreover, a dependency-set algorithm must ensure that:
@@
     \item For any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$ (with $p$ possibly equal to $q$), if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
+    \item If a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{decide}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
 \end{enumerate}

### 2016-02-11 podc16/paper2.tex
new pseudocode

--- 2016-02-11 podc16/paper2.tex
@@
 The sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties described below.
-We refer the reader to~\cite{Lamport05GeneralizeConsensus} for the definition of c-structs and generalized consensus. Note that we use exactly the same notation.
+We refer the read to~\cite{Lamport05GeneralizeConsensus} for the definition of c-structs and generalized consensus. Note that we use exactly the same notation.
 
@@
 An example of implementation of such an algorithm is depicted by the pseudocode of \Cref{fig:deps-set-algo}.
-The idea is to let every process broadcast $c$ to all the processes in the system, in order to announce $c$.
-Then whenever a process $p$ receives $c$, it replies to the sender with the set $\mathcal{D}$ of commands that $p$ has observed so far, and such that $c$ and any command in $\mathcal{D}$ do not commute.
-Afterwards, the process announcing $c$ collects the sets $\mathcal{D}$ for $c$ from a quorum of $\mathcal{Q}$ processes, and it locally suggests the union of those sets as final dependency-set for $c$.
+The idea is to let every process broadcast $c$ to all the processes in the system, in order to announce $c$ (lines \ref{alg:announce}--\ref{alg:send-new-cmd}).
+Then whenever a process $p$ receives $c$, it replies to the sender with the set $\mathcal{D}$ of commands that $p$ has observed so far, and such that $c$ and any command in $\mathcal{D}$ do not commute (lines \ref{alg:receive-new-cmd}--\ref{alg:send-deps}).
+Afterwards, the process announcing $c$ collects the sets $\mathcal{D}$ for $c$ from a \textit{quorum} of $\mathcal{Q}$ processes, and it locally suggests the union of those sets as final dependency-set for $c$ (lines \ref{alg:receive-deps}--\ref{alg:call-suggest}).
+For the purpose of this example we allow for $\mathcal{Q}$ to be equal to any positive value that is not greater than the number of processes, as long as the following requirement is met: the intersection of any two quorums that are selected by processes to collect the sets of dependencies is not empty.
 
@@
 Since any command in a dependency-set that is suggested by a process must be in the set $cmds_q$ of the commands received by some process $q$, and a process receives a command if and only if there exists some process that previously announced the command, then it follows that any command in any suggested dependency-set has been previously announced.
+Further, since a process can suggest a set of dependencies $deps_p[c]$ for a command $c$ after it has received a reply for $c$ from at least a quorum of processes, we have that $\opn{suggest}_p \left( c,deps_p[c]\right)$ is called only if a quorum of processes $q$ have $c$ in their $cmds_q$.
+Therefore, by the requirement that we enforced on the quorums, we can conclude that for any two suggestions $\opn{suggest}_p \left( c_1,deps_p[c_1]\right)$ and $\opn{suggest}_q \left( c_2,deps_q[c_2]\right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_q[c_2]$ or $c_2 \in deps_p[c_1]$.
 
@@
 		\begin{algorithmic}[1]
-			{\small
+			{\scriptsize
 			\START[\texttt{init$_p$}]{{\bf upon}}{$()$} \label{alg:init}
@@
 }	
-\hspace*{0.3in}\subfigure[receive of \texttt{cmd} and \texttt{deps} messages, and suggestion of dependencies for a command on process $p$.]{
+\hspace*{0.1in}\subfigure[receive of \texttt{cmd} and \texttt{deps} messages, and suggestion of dependencies for a command on process $p$.]{
 \noindent\fbox{%
@@
 		\begin{algorithmic}[1]
-			{\small
+			\makeatletter
+			\setcounter{ALG@line}{8}
+			\makeatother
+			{\scriptsize
 			\START[$\langle \texttt{cmd}$ $|$ $c \rangle$]{{\bf upon receive}}{ {\bf from} $q$} \label{alg:receive-new-cmd}
-			\State $\mathcal{D} \gets \{ d : d \in cmds_p \land d\neq c \land d \asymp c \}$
+			\State $\mathcal{D} \gets \{ d : d \in cmds_p \land$
+			\Statex \hspace{2cm}$\land d\neq c \land d \asymp c \}$
 			\State $cmds_p\gets cmds_p\cup\{c\}$
@@
 			\Statex
-			\START[]{{\bf upon}}{$\exists c$ $:$ $ | heard_p[c] | \geq \mathcal{Q}$ $\land$ $suggested_p[c]=false$} \label{alg:receive-deps}
+			\START[]{{\bf upon}}{$\exists c$ $:$ $ | heard_p[c] | \geq \mathcal{Q}$ $\land$} \label{alg:receive-deps}
+			\Statex \hspace{1cm}$\land$ $suggested_p[c]=false$
 			\State $suggested_p[c]\gets true$
-			\State {\bf call} $\texttt{suggest}_p(c,deps_p[c])$
+			\State {\bf call} $\texttt{suggest}_p(c,deps_p[c])$ \label{alg:call-suggest}
 			\END
@@
 }
-\caption{Example of dependency-set algorithm implementation.}
+\hspace*{0.1in}\subfigure[Leaderless generalized consensus implementation.]{
+\noindent\fbox{%
+	\begin{varwidth}{\dimexpr\linewidth-2\fboxsep-2\fboxrule\relax}
+
+		\begin{algorithmic}[1]
+			{\scriptsize
+			\START[$\opn{init}_p$]{{\bf upon}}{$()$}
+			\State $m_p\gets \emptyset$
+			\State $\sigma_p\gets \bot$
+			\END
+			\START[$\opn{GC-propose}_p$]{{\bf upon}}{$(c)$}
+			\State {\bf call} $\opn{announce}_p\left( c \right)$
+			\END
+			\Statex
+			\START[$\opn{suggest}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
+			\State  $\opn{propose}_p( c,\mathcal{D})$ 			
+			\END
+			\Statex
+			\START[$\opn{decide}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
+			\State  $m_p \gets m_p \oplus \left[ c \mapsto \mathcal{D}\right]$	
+			\State $\sigma_p \gets  \opn{c-struct}\left(m_p\right)$
+			\State $\opn{GC-decide}_p(\sigma_p)$		
+			\END
+			}
+		\end{algorithmic}
+	\end{varwidth}%
+}
+\label{fig:deps-set-algo3}
+}
+\caption{Example of dependency-set algorithm and leaderless generalized consensus implementations.}
 \label{fig:deps-set-algo}
@@
 For example, $\preceq_h = \aset{\left( c_1,c_3 \right)}$. % and $\opn{TOs}\left( h \right)$ is the set of all permutations of $\aseq{c_1,c_2,c_3,c_4}$ in which $c_1$ comes before $c_3$.
-If $c\in V\left( g \right)$, we write $\left[ c \right]_g$ for the strongly connected component of $g$ to which $c$ belongs. 
 %\textbf{Equivalence relation induced by a graph.}
-%A directed graph $g$ \emph{induces an equivalence relation} $\sim_g$ on its vertices defined such that $c_1 \sim_g c_2$ iff $c_1$ and $c_2$ belong to the same strongly connected component.
+A directed graph $g$ \emph{induces an equivalence relation} $\sim_g$ on its vertices defined such that $c_1 \sim_g c_2$ iff $c_1$ and $c_2$ belong to the same strongly connected component.
 We write $\left[ v \right]_g$ for the equivalence class of $v$ under $\sim_g$.

### 2016-02-12 podc16/paper2.tex
spotted a mistake in graph processing. Left a few comments.

@@
 
-\begin{comment}
-
-A c-struct set is a set containing the element $\bot$ and with an append operator $\bullet$ such that $\sigma \bullet c$ is a c-struct, for any c-struct $\sigma$ and command $c$. 
-We also write the $\sigma\bullet\aseq{c_1,\cdots,c_n}$ for the c-struct $\sigma\bullet c_1\bullet\cdots\bullet c_n$.
-C-structs are partially ordered by the relation $\sqsubseteq$ defined such that $\sigma_1\sqsubseteq\sigma_2$ iff there exists a sequence of commands $cs$ such that $\sigma_2=\sigma_1\bullet cs$.
-Two c-structs are compatible when they have a common upper bound; when this is the case, two c-structs $\sigma_1$ and $\sigma_2$ have a least upper bound, noted $\bigsqcup\left( \sigma_1,\sigma_2 \right)$.
-We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
-Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands (see~\cite{Lamport05GeneralizeConsensus} for a complete discussion on c-struct).
-In the rest of the paper the symbol $c$, possibly sub-scripted, ranges over commands, and the symbol $\mathcal{D}$, possibly sub-scripted, ranges over sets of commands.
-
-We can now state the generalized-consensus properties: non-triviality requires that any decided c-struct be of the form $\bot\bullet cs$, where $cs$ is a sequence of proposed commands; consistency requires than any two decided c-structs $\sigma_1$ and $\sigma_2$ be compatible; stability requires that when a process $p$ decides a c-struct $\sigma_1$ and time $t_1$ and $\sigma_2$ at time $t_2$, then $t_1 \leq t_2$ implies that $\sigma_1\sqsubseteq \sigma_2$; finally, liveness requires that
-if a command keeps beeing proposed, then a c-struct $\sigma$ containing the command is eventually decided.
-
-\end{comment}
-
 We now present our abstract leaderless generalized consensus algorithm.
@@
 However, we show that a leaderless  generalized consensus algorithm is always safe and is live when its underlying dependency-set and map-agreement algorithms are safe and live.
+The leaderless generalized consensus algorithm orchestrates the two dependency-set and map-agreement sub-algorithms as shown in \cref{fig:deps-set-algo3}.
 
@@
 
-\subsection{The Abstract Generalized Consensus Algorithm}
+\subsection{Processing the Dependency Graph}
 \label{abstract-algo}
-We now propose describe our abstract generalized consensus algorithm. 
+
+\begin{comment}
 Local variables: $m_p$, a map from commands to dependency sets, initially the empty map; $\sigma_p$, a c-struct, initially $\bot$.
@@
 
-The graph processing algorithm is executed locally by a node. It must compute the value of $\opn{c-struct}\left( g_p \right)$, as defined below.
-Its main property is Lemma 1; together with Invariants 1 and 2 (TODO, the invariants guaranteed by the dependency-set algorithm and the map-agreement algorithm), Lemma 1 implies that leaderless algorithms safely implement generalized consensus.
+\end{comment}
+
+The graph processing algorithm is executed locally by a node (line 10 of \cref{fig:deps-set-algo3}). It must compute the value of $\opn{c-struct}\left( g_p \right)$, as defined below.
+Its main property is Lemma 1, which with the properties of the sub-algorithms imply safety and liveness. 
 We assume that processes initially agree, for each set of commands $cs$, on a total order $<_{cs}$ on $cs$.
@@
 
+\todo{Partial orders may not be what we need (see Lemma 2), instead:}
+Define the intersection of two graphs as $V\left( g_1 \cap g_2 \right)=V\left( g_1 \right)\cap V\left( g_2 \right)$ and $E\left( g_1 \cap g_2 \right)= E\left( g_1 \right)\cap E\left( g_2 \right)$, and define their union similarly.
+Two graphs $g_1$ and $g_2$ are \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$.
+\todo{Does this make any sense?}
+
 \textbf{Lemma 1}
@@
 for every pair of non-commutative commands $c_1,c_2 \in V\left(g_2\right)$, either $\left( c_1,c_2 \right)\in E\left( g_2 \right)$ or $\left( c_2,c_1 \right)\in E\left( g_2 \right)$;
-$g_1 \leq g_2$;
+$g_1 \leq g_2$; \todo{Instead of the previous one: $g_1$ and $g_2$ are compatible};
 $l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
@@
 
-Define the set of linearizations $\opn{Lins}\left( g \right)$ of a graph $g$ a the set of sequences whose length is the cardinality of $V\left( g \right)$, which contain each element of $V\left( g \right)$ exactly once, and ...
+Define the set of linearizations $\opn{Lins}\left( g \right)$ of a graph $g$ as the set of sequences whose length is the cardinality of $V\left( g \right)$, which contain each element of $V\left( g \right)$ exactly once, and ...
 \todo{or first define the partial order induced by a graph and then its compatible total orders.}
@@
 
-\subsection{Correctness Proof}
+\subsection{Consistency and Liveness}
+\textbf{Lemma 2}
+If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, and $t_1 \leq t_2$, then $g_p^1 \leq g_q^2$.
+\todo{This is wrong\ldots}
+What about: If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, and $t_1 \leq t_2$, then $g_p^1$ and $g_q^2$ are compatible.
 
@@
 
+\appendix
+
+\section{Generalized Consensus}
+
+A c-struct set is a set containing the element $\bot$ and with an append operator $\bullet$ such that $\sigma \bullet c$ is a c-struct, for any c-struct $\sigma$ and command $c$. 
+We also write the $\sigma\bullet\aseq{c_1,\cdots,c_n}$ for the c-struct $\sigma\bullet c_1\bullet\cdots\bullet c_n$.
+C-structs are partially ordered by the relation $\sqsubseteq$ defined such that $\sigma_1\sqsubseteq\sigma_2$ iff there exists a sequence of commands $cs$ such that $\sigma_2=\sigma_1\bullet cs$.
+Two c-structs are compatible when they have a common upper bound; when this is the case, two c-structs $\sigma_1$ and $\sigma_2$ have a least upper bound, noted $\bigsqcup\left( \sigma_1,\sigma_2 \right)$.
+We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
+Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands (see~\cite{Lamport05GeneralizeConsensus} for a complete discussion on c-struct).
+In the rest of the paper the symbol $c$, possibly sub-scripted, ranges over commands, and the symbol $\mathcal{D}$, possibly sub-scripted, ranges over sets of commands.
+
+We can now state the generalized-consensus properties: non-triviality requires that any decided c-struct be of the form $\bot\bullet cs$, where $cs$ is a sequence of proposed commands; consistency requires than any two decided c-structs $\sigma_1$ and $\sigma_2$ be compatible; stability requires that when a process $p$ decides a c-struct $\sigma_1$ and time $t_1$ and $\sigma_2$ at time $t_2$, then $t_1 \leq t_2$ implies that $\sigma_1\sqsubseteq \sigma_2$; finally, liveness requires that
+if a command keeps beeing proposed, then a c-struct $\sigma$ containing the command is eventually decided.
+
 \end{document}

### 2016-02-12 podc16/paper2.tex
minor

@@
 If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, and $t_1 \leq t_2$, then $g_p^1 \leq g_q^2$.
-\todo{This is wrong\ldots}
-What about: If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, and $t_1 \leq t_2$, then $g_p^1$ and $g_q^2$ are compatible.
+\todo{This is wrong\ldots
+What about: If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, and $t_1 \leq t_2$, then $g_p^1$ and $g_q^2$ are compatible.}
 

### 2016-02-12 podc16/paper2.tex
Checked with TLC that the graph processing formalization in the brief announcement is correct.

@@
 Then we have (a), $\bot\bullet l_1 \sqsubseteq \bot\bullet l_2$, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.
+\todo{No, then we have that $\bot\bullet l_1$ and $\bot\bullet l_2$ are compatible.}
 
@@
 \todo{This is wrong\ldots
-What about: If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, and $t_1 \leq t_2$, then $g_p^1$ and $g_q^2$ are compatible.}
+What about: If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible.}
+
+\textbf{Lemma 3}
+Consider $g_1$ and $g_2$. Assume that if $v\in V\left( g_1 \right)$ and $v\in V\left( g_2 \right)$ then $n\left( g_1 \right)=n\left( g2 \right)$.
+
+\textbf{Lemma 4}
+If $m_p\subseteq m_q$
 

### 2016-02-12 podc16/paper2.tex
fixed the graph processing section after checking it with TLC

@@
 
+\begin{comment}
 We now define a partial order on graphs.
@@
 We have that $h'\leq h$, but $ h''\not\leq h$.
+\end{comment}
 
-\todo{Partial orders may not be what we need (see Lemma 2), instead:}
 Define the intersection of two graphs as $V\left( g_1 \cap g_2 \right)=V\left( g_1 \right)\cap V\left( g_2 \right)$ and $E\left( g_1 \cap g_2 \right)= E\left( g_1 \right)\cap E\left( g_2 \right)$, and define their union similarly.
-Two graphs $g_1$ and $g_2$ are \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$.
-\todo{Does this make any sense?}
+Define two graphs $g_1$ and $g_2$ as being \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$.
 
 \textbf{Lemma 1}
-Assume that: $g_1$ and $g_2$ are two dependency graphs; 
-for every pair of non-commutative commands $c_1,c_2 \in V\left(g_2\right)$, either $\left( c_1,c_2 \right)\in E\left( g_2 \right)$ or $\left( c_2,c_1 \right)\in E\left( g_2 \right)$;
-$g_1 \leq g_2$; \todo{Instead of the previous one: $g_1$ and $g_2$ are compatible};
-$l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
-Then we have (a), $\bot\bullet l_1 \sqsubseteq \bot\bullet l_2$, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.
-\todo{No, then we have that $\bot\bullet l_1$ and $\bot\bullet l_2$ are compatible.}
+Assume that $g_1$ and $g_2$ are two dependency graphs, 
+that for every pair of non-commutative commands $c_1,c_2 \in V\left(g_2\right)$, either $\left( c_1,c_2 \right)\in E\left( g_2 \right)$ or $\left( c_2,c_1 \right)\in E\left( g_2 \right)$,
+that $g_1$ and $g_2$ are compatible,
+and that $l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
+Then we have (a), $\bot\bullet l_1$ and $\bot\bullet l_2$ are compatible c-structs, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.
 
@@
 \subsection{Consistency and Liveness}
+
 \textbf{Lemma 2}
-If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, and $t_1 \leq t_2$, then $g_p^1 \leq g_q^2$.
-\todo{This is wrong\ldots
-What about: If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible.}
+If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible.
 
+\begin{comment}
 \textbf{Lemma 3}
@@
 If $m_p\subseteq m_q$
+
+\end{comment}
 

### 2016-02-12 podc16/paper2.tex
gc and comments

--- 2016-02-12 podc16/paper2.tex
@@
 A process $p$ receiving a command $c$ proceeds in three phases.
-First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $cs$ that the command $c$ may depend on.
+First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $cs$ \todo{are we sure about the notation cs?} that the command $c$ may depend on.
 Second, the process $p$ proposes the mapping  $c\mapsto cs$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto cs'$ for $c$, in which case $cs'$ becomes the set of dependencies of $c$, and we say that $c$ is executable when its dependency set is empty or when all its dependencies are executable.
@@
 However, we show that a leaderless  generalized consensus algorithm is always safe and is live when its underlying dependency-set and map-agreement algorithms are safe and live.
-The leaderless generalized consensus algorithm orchestrates the two dependency-set and map-agreement sub-algorithms as shown in \cref{fig:deps-set-algo3}.
+The leaderless generalized consensus algorithm orchestrates the two dependency-set and map-agreement sub-algorithms as shown in \cref{fig:deps-set-algo3}.\todo{This pseudocode must be in figure (a).
+Also we have to better describe the pseudocode.}
 
@@
 
-A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{decide}\left( c,\mathcal{D} \right)$ (meant to receive inputs from the agreement sub-algorithm), where $p\in P$ is the process executing the call.
+A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and \todo{this is not clear: an input interface $\opn{decide}\left( c,\mathcal{D} \right)$ (meant to receive inputs from the agreement sub-algorithm)}, where $p\in P$ is the process executing the call.
 Moreover, a dependency-set algorithm must ensure that:
@@
 
-An example of implementation of such an algorithm is depicted by the pseudocode of \Cref{fig:deps-set-algo}.
+An example of implementation of such an algorithm is depicted by the pseudocode of Figures~\ref{fig:deps-set-algo1} and~\ref{fig:deps-set-algo2}.
 The idea is to let every process broadcast $c$ to all the processes in the system, in order to announce $c$ (lines \ref{alg:announce}--\ref{alg:send-new-cmd}).
@@
 Afterwards, the process announcing $c$ collects the sets $\mathcal{D}$ for $c$ from a \textit{quorum} of $\mathcal{Q}$ processes, and it locally suggests the union of those sets as final dependency-set for $c$ (lines \ref{alg:receive-deps}--\ref{alg:call-suggest}).
-For the purpose of this example we allow for $\mathcal{Q}$ to be equal to any positive value that is not greater than the number of processes, as long as the following requirement is met: the intersection of any two quorums that are selected by processes to collect the sets of dependencies is not empty.
+For the purpose of this example we allow for $\mathcal{Q}$ to be equal to any positive integer that is not greater than the number of processes, as long as the following requirement is met: the intersection of any two quorums that are selected by processes to collect the sets of dependencies is not empty.
 
 The provided implementation satisfies the specification of a dependency-set algorithm for the following reasons.
-Since any command in a dependency-set that is suggested by a process must be in the set $cmds_q$ of the commands received by some process $q$, and a process receives a command if and only if there exists some process that previously announced the command, then it follows that any command in any suggested dependency-set has been previously announced.
-Further, since a process can suggest a set of dependencies $deps_p[c]$ for a command $c$ after it has received a reply for $c$ from at least a quorum of processes, we have that $\opn{suggest}_p \left( c,deps_p[c]\right)$ is called only if a quorum of processes $q$ have $c$ in their $cmds_q$.
+Since any command in a dependency-set that is suggested by a process (lines~\ref{alg:update-deps-p} and~\ref{alg:call-suggest}) must be in the set $cmds_q$ of the commands received by some process $q$ (line~\ref{alg:update-d}), and a process receives a command if and only if there exists some process that previously announced the command (line~\ref{alg:send-new-cmd}), then it follows that any command in any suggested dependency-set has been previously announced.
+Further, since a process can suggest a set of dependencies $deps_p[c]$ for a command $c$ after it has received a reply for $c$ from at least a quorum of processes (lines~\ref{alg:receive-deps}--\ref{alg:call-suggest}), we have that $\opn{suggest}_p \left( c,deps_p[c]\right)$ is called only if a quorum of processes $q$ have $c$ in their $cmds_q$ (line~\ref{alg:update-cmds-p}).
 Therefore, by the requirement that we enforced on the quorums, we can conclude that for any two suggestions $\opn{suggest}_p \left( c_1,deps_p[c_1]\right)$ and $\opn{suggest}_q \left( c_2,deps_q[c_2]\right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_q[c_2]$ or $c_2 \in deps_p[c_1]$.
@@
 			\START[$\langle \texttt{cmd}$ $|$ $c \rangle$]{{\bf upon receive}}{ {\bf from} $q$} \label{alg:receive-new-cmd}
-			\State $\mathcal{D} \gets \{ d : d \in cmds_p \land$
+			\State $\mathcal{D} \gets \{ d : d \in cmds_p \land$\label{alg:update-d}
 			\Statex \hspace{2cm}$\land d\neq c \land d \asymp c \}$
-			\State $cmds_p\gets cmds_p\cup\{c\}$
+			\State $cmds_p\gets cmds_p\cup\{c\}$\label{alg:update-cmds-p}
 			\State {\bf send} $\langle \texttt{deps}$ $|$ $c,\mathcal{D} \rangle$ {\bf to} $q$\label{alg:send-deps}
@@
 			\START[$\langle \texttt{deps}$ $|$ $c,\mathcal{D} \rangle$]{{\bf upon receive}}{ {\bf from} $q$} \label{alg:receive-deps}
-			\State $deps_p[c]\gets deps_p[c]\cup \mathcal{D}$
+			\State $deps_p[c]\gets deps_p[c]\cup \mathcal{D}$\label{alg:update-deps-p}
 			\State $heard_p[c]\gets heard_p[c]\cup \{q\}$
@@
 
-\todo{``Suggest'' can also be used to trigger recovery}
-\todo{Write a suggestion on about GC of dependencies. Remove any command c such that a committed command depends on c. Also us the GC of the map-agreement component.}
+Any implementation of dependency-set algorithm can rely on the garbage collection of the map-agreement module to get rid of redundant information associated with specific commands in its data structures, with the purpose of preventing the dependency sets to grow indefinitely.
+
+\todo{``Suggest'' can also be used to trigger recovery. Describe better this part.}
+%\todo{Write a suggestion on about GC of dependencies. Remove any command c such that a committed command depends on c. Also us the GC of the map-agreement component.}
 
@@
 
-Note that any SMR algorithm, like MultiPaxos, can be used to implementing map-agreement by uniquely associating commands to MultiPaxos instance numbers.
+An example of implementation of the map-agreement algorithm can be any state machine replication algorithm, like MultiPaxos~\cite{paxos-made-simple}, where a pair of command $c$ and dependency-set $\mathcal{D}$ can be proposed to a instance of MultiPaxos that is identified by the command $c$ itself.
+%used to implementing map-agreement by uniquely associating commands to MultiPaxos instance numbers.
 
-\todo{Cite Generalized Paxos as an example for bounding the data structures.}
+\todo{Cite Generalized Paxos as an example for bounding the data structures. Problem: we do not describe any particular data structure here. It is not clear why we should write about GC here.}
 

### 2016-02-12 podc16/paper2.tex

@@
 We have that $h'\leq h$, but $ h''\not\leq h$.
+\end{comment}
 
-\todo{Partial orders may not be what we need (see Lemma 2), instead:}
 Define the intersection of two graphs as $V\left( g_1 \cap g_2 \right)=V\left( g_1 \right)\cap V\left( g_2 \right)$ and $E\left( g_1 \cap g_2 \right)= E\left( g_1 \right)\cap E\left( g_2 \right)$, and define their union similarly.
-Two graphs $g_1$ and $g_2$ are \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$.
-\todo{Does this make any sense?}
+Define two graphs $g_1$ and $g_2$ as being \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$.
 
 \textbf{Lemma 1}
-Assume that: $g_1$ and $g_2$ are two dependency graphs; 
-for every pair of non-commutative commands $c_1,c_2 \in V\left(g_2\right)$, either $\left( c_1,c_2 \right)\in E\left( g_2 \right)$ or $\left( c_2,c_1 \right)\in E\left( g_2 \right)$;
-$g_1 \leq g_2$; \todo{Instead of the previous one: $g_1$ and $g_2$ are compatible};
-$l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
-Then we have (a), $\bot\bullet l_1 \sqsubseteq \bot\bullet l_2$, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.
+Assume that $g_1$ and $g_2$ are two dependency graphs, 
+that for every pair of non-commutative commands $c_1,c_2 \in V\left(g_2\right)$, either $\left( c_1,c_2 \right)\in E\left( g_2 \right)$ or $\left( c_2,c_1 \right)\in E\left( g_2 \right)$,
+that $g_1$ and $g_2$ are compatible,
+and that $l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
+Then we have (a), $\bot\bullet l_1$ and $\bot\bullet l_2$ are compatible c-structs, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.
 
@@
 \subsection{Consistency and Liveness}
+
 \textbf{Lemma 2}
-If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, and $t_1 \leq t_2$, then $g_p^1 \leq g_q^2$.
-\todo{This is wrong\ldots
-What about: If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, and $t_1 \leq t_2$, then $g_p^1$ and $g_q^2$ are compatible.}
+If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible.
+
+\begin{comment}
+\textbf{Lemma 3}
+Consider $g_1$ and $g_2$. Assume that if $v\in V\left( g_1 \right)$ and $v\in V\left( g_2 \right)$ then $n\left( g_1 \right)=n\left( g2 \right)$.
+
+\textbf{Lemma 4}
+If $m_p\subseteq m_q$
+
+\end{comment}
 

### 2016-02-12 podc16/paper2.tex
a few fixes in graph processing

@@
 g_p \right) = \aset{c_1,c_2}$ and $E\left( g_p \right) = \aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right)}$.
+\todo{This is not accurate, see the TLA spec.}
 
@@
 For example, consider $h$ such that $V\left( h \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$. 
-We have that $\preceq_h = \aset{\left( c_1,c_3 \right)}$. % and $\opn{TOs}\left( h \right)$ is the set of all permutations of $\aseq{c_1,c_2,c_3,c_4}$ in which $c_1$ comes before $c_3$.
-If $c\in V\left( g \right)$, we write $\left[ c \right]_g$ for the strongly connected component of $g$ to which $c$ belongs. 
-%\textbf{Equivalence relation induced by a graph.}
-%A directed graph $g$ \emph{induces an equivalence relation} $\sim_g$ on its vertices defined such that $c_1 \sim_g c_2$ iff $c_1$ and $c_2$ belong to the same strongly connected component.
-We write $\left[ v \right]_g$ for the equivalence class of $v$ under $\sim_g$.
+We have that $\preceq_h = \aset{\left( c_1,c_3 \right)}$.
+%If $c\in V\left( g \right)$, we write $\left[ c \right]_g$ for the strongly connected component of $g$ to which $c$ belongs. 
 %\textbf{Linearizations of a dependency graph.}
-We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1 \sim_g c_2$ and $ c_1 <_{\left[ c_1 \right]_g} c_2$ hold then $c_1 \ll c_2$.
+We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1$ and $c_2$ belong to the same strongly connected component $cs$ and $ c_1 <_{cs} c_2$ hold, then $c_1 \ll c_2$.
 For example, assuming that $c_2 \leq_{\aset{c_1,c_2}} c_1$, the linearizations of $h$ are $\aseq{c_2,c_1,c_3,c_4}$, $\aseq{c_2,c_1,c_4,c_3}$, $\aseq{c_2,c_4,c_1,c_3}$, and $\aseq{c_4,c_2,c_1,c_3}$.
@@
 
-Define the intersection of two graphs as $V\left( g_1 \cap g_2 \right)=V\left( g_1 \right)\cap V\left( g_2 \right)$ and $E\left( g_1 \cap g_2 \right)= E\left( g_1 \right)\cap E\left( g_2 \right)$, and define their union similarly.
+Define the intersection of two graphs such that $V\left( g_1 \cap g_2 \right)=V\left( g_1 \right)\cap V\left( g_2 \right)$ and $E\left( g_1 \cap g_2 \right)= E\left( g_1 \right)\cap E\left( g_2 \right)$, and define their union such that $V\left( g_1 \cup g_2 \right)=V\left( g_1 \right)\cup V\left( g_2 \right)$ and $E\left( g_1 \cup g_2 \right)= E\left( g_1 \right)\cup E\left( g_2 \right)$.
 Define two graphs $g_1$ and $g_2$ as being \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$.

### 2016-02-12 podc16/paper2.tex
reintroduced a few facts about c-structs, made a few fixes

@@
 The sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties described below.
-We refer the reader to~\cite{Lamport05GeneralizeConsensus} for the definition of c-structs and generalized consensus. Note that we use exactly the same notation.
+We refer the reader to~\cite{Lamport05GeneralizeConsensus}, and we use the same notation, for the definition of c-structs and generalized consensus.
+Briefly, the c-structs are a set of elements, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct.
+C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $cs$ such that $\sigma_2 = \sigma_1\bullet cs$ and $\bot\sqsubseteq \sigma$ for any
+$\sigma$. Two c-structs are compatible when they have a
+common upper bound.
+Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands.
+The consistency property of generalized consensus requires that any two decided c-structs be compatible.
 
@@
 
-An example of implementation of the map-agreement algorithm can be any state machine replication algorithm, like MultiPaxos~\cite{paxos-made-simple}, where a pair of command $c$ and dependency-set $\mathcal{D}$ can be proposed to a instance of MultiPaxos that is identified by the command $c$ itself.
+An example of implementation of the map-agreement algorithm can be any state machine replication algorithm, like MultiPaxos~\cite{lamport2001paxos}, where a pair of command $c$ and dependency-set $\mathcal{D}$ can be proposed to a instance of MultiPaxos that is identified by the command $c$ itself.
 %used to implementing map-agreement by uniquely associating commands to MultiPaxos instance numbers.
@@
 that $g_1$ and $g_2$ are compatible,
-and that $l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
+that $l_1$ is a linearization of $g_1$, and that $l_2$ is a linearization of $g_2$. 
 Then we have (a), $\bot\bullet l_1$ and $\bot\bullet l_2$ are compatible c-structs, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.

### 2016-02-12 podc16/paper2.tex
more changes in graph processing, referencable definition and lemma

@@
 
-The graph processing algorithm is executed locally by a node (line 10 of \cref{fig:deps-set-algo3}). It must compute the value of $\opn{c-struct}\left( g_p \right)$, as defined below.
-Its main property is Lemma 1, which with the properties of the sub-algorithms imply safety and liveness. 
+The graph processing algorithm is executed locally by a node (line 10 of \cref{fig:deps-set-algo3}). It must compute the value of $\opn{c-struct}\left( g_p \right)$, as defined in \cref{lindef}.
+Its key property is \cref{lemma1}. 
 We assume that processes initially agree, for each set of commands $cs$, on a total order $<_{cs}$ on $cs$.
@@
 %\textbf{Local dependency graph computed by a process.} 
-The local variable $cs_p$ of a process $p$ can be seen as describing a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the domain of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only both $c_1$ and $c_2$ are in $V\left( g_p \right)$ and $c_2 \in m_p\left[ c_1 \right]$ (i.e., $c_1$ depends on $c_2$).
-For example, if $cs_p = \left[ c_1 \mapsto \aset{c_2,c_3}, c_2 \mapsto \aset{c_1} \right]$ then $V\left(
-g_p \right) = \aset{c_1,c_2}$ and $E\left( g_p \right) = \aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right)}$.
-\todo{This is not accurate, see the TLA spec.}
+We say that a command $c$ in the domain of $m_p$ is \emph{complete} when no command $c$ which is not in the domain of $m_p$ can be reached from $c$ by recursively following $c$'s dependencies.
+The local variable $m_p$ denotes a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the complete commands of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ iff $c_2 \in m_p\left[ c_1 \right]$ (i.e., $c_1$ depends on $c_2$).
+For example, if $m_p = \left[ c_1 \mapsto \aset{c_2,c_4}, c_2 \mapsto \aset{c_3}, c_3 \mapsto \aset{c_2}\right]$ then $V\left( g_p \right) = \aset{c_2,c_3}$ and $E\left( g_p \right) = \aset{\left( c_2,c_3 \right),\left( c_3,c_2 \right)}$.
 
 %\textbf{Orders induced by a graph.}
-A directed graph $g$ \emph{induces a partial order} $\preceq_g$ on its vertices defined such that $c_1 \preceq_g c_2$ iff there is a path from $c_1$ to $c_2$ and there is no path from $c_2$ to $c_1$.
+A directed graph $g$ \emph{induces a partial order} $\preceq_g$ on its vertices defined such that $c_1 \preceq_g c_2$ iff there is a path from $c_1$ to $c_2$ and none from $c_2$ to $c_1$.
 %Let $\opn{TOs}\left( g \right)$ be the set of total orders $o$ such that $\preceq_g \subseteq o$.
-For example, consider $h$ such that $V\left( h \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$. 
+For example, consider $h$ where $V\left( h \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$. 
 We have that $\preceq_h = \aset{\left( c_1,c_3 \right)}$.
@@
 %\textbf{Linearizations of a dependency graph.}
-We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every pair of vertices $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1$ and $c_2$ belong to the same strongly connected component $cs$ and $ c_1 <_{cs} c_2$ hold, then $c_1 \ll c_2$.
+We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1$ and $c_2$ belong to the same strongly connected component $cs$ and $ c_1 <_{cs} c_2$ hold, then $c_1 \ll c_2$.
 For example, assuming that $c_2 \leq_{\aset{c_1,c_2}} c_1$, the linearizations of $h$ are $\aseq{c_2,c_1,c_3,c_4}$, $\aseq{c_2,c_1,c_4,c_3}$, $\aseq{c_2,c_4,c_1,c_3}$, and $\aseq{c_4,c_2,c_1,c_3}$.
@@
 
-\textbf{Lemma 1}
+\begin{mylemma}\label{lemma1}
 Assume that $g_1$ and $g_2$ are two dependency graphs, 
@@
 Then we have (a), $\bot\bullet l_1$ and $\bot\bullet l_2$ are compatible c-structs, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.
+\end{mylemma}
 
-\textbf{Definition 1}
+\begin{mydef}\label{lindef}
 We define $\opn{c-struct}\left( m_p \right) = \bot\bullet l$ where $l$ is a linearization of $g_p$.
 By lemma 1.b, $\opn{c-struct}\left( m_p \right)$ is well-defined.
+\end{mydef}
 

### 2016-02-12 podc16/paper2.tex
replaced all the cs by \mathcal{C}

@@
 Briefly, the c-structs are a set of elements, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct.
-C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $cs$ such that $\sigma_2 = \sigma_1\bullet cs$ and $\bot\sqsubseteq \sigma$ for any
+C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $\mathcal{C}$ such that $\sigma_2 = \sigma_1\bullet \mathcal{C}$ and $\bot\sqsubseteq \sigma$ for any
 $\sigma$. Two c-structs are compatible when they have a
@@
 A process $p$ receiving a command $c$ proceeds in three phases.
-First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $cs$ \todo{are we sure about the notation cs?} that the command $c$ may depend on.
-Second, the process $p$ proposes the mapping  $c\mapsto cs$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto cs'$ for $c$, in which case $cs'$ becomes the set of dependencies of $c$, and we say that $c$ is executable when its dependency set is empty or when all its dependencies are executable.
+First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $\mathcal{C}$ that the command $c$ may depend on.
+Second, the process $p$ proposes the mapping  $c\mapsto \mathcal{C}$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto \mathcal{C}'$ for $c$, in which case $\mathcal{C}'$ becomes the set of dependencies of $c$, and we say that $c$ is executable when its dependency set is empty or when all its dependencies are executable.
 Finally, when $c$ is executable, the process $p$ may execute $c$ by deciding the c-struct obtained by running the \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
@@
 Its key property is \cref{lemma1}. 
-We assume that processes initially agree, for each set of commands $cs$, on a total order $<_{cs}$ on $cs$.
-In practice, if commands are attached a unique identifier among a totally ordered set, $<_{cs}$ can be the restriction of this total order on $cs$.
+We assume that processes initially agree, for each set of commands $\mathcal{C}$, on a total order $<_{\mathcal{C}}$ on $\mathcal{C}$.
+In practice, if commands are attached a unique identifier among a totally ordered set, $<_{\mathcal{C}}$ can be the restriction of this total order on $\mathcal{C}$.
 
@@
 %\textbf{Linearizations of a dependency graph.}
-We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1$ and $c_2$ belong to the same strongly connected component $cs$ and $ c_1 <_{cs} c_2$ hold, then $c_1 \ll c_2$.
+We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1$ and $c_2$ belong to the same strongly connected component $\mathcal{C}$ and $ c_1 <_{\mathcal{C}} c_2$ hold, then $c_1 \ll c_2$.
 For example, assuming that $c_2 \leq_{\aset{c_1,c_2}} c_1$, the linearizations of $h$ are $\aseq{c_2,c_1,c_3,c_4}$, $\aseq{c_2,c_1,c_4,c_3}$, $\aseq{c_2,c_4,c_1,c_3}$, and $\aseq{c_4,c_2,c_1,c_3}$.
@@
 
-\textbf{Linearization function.} Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ cs \right]$ is a total order on the set $cs$.
+\textbf{Linearization function.} Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ \mathcal{C} \right]$ is a total order on the set $\mathcal{C}$.
 
@@
 
-Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ cs \right]$ contains each command in $cs$ exactly once.
+Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ \mathcal{C} \right]$ contains each command in $\mathcal{C}$ exactly once.
 A sequence of commands $cseq$ is an $f$-linearization of the graph $g$ if and only there is a sequence whose element of strongly connected
@@
 We also write the $\sigma\bullet\aseq{c_1,\cdots,c_n}$ for the c-struct $\sigma\bullet c_1\bullet\cdots\bullet c_n$.
-C-structs are partially ordered by the relation $\sqsubseteq$ defined such that $\sigma_1\sqsubseteq\sigma_2$ iff there exists a sequence of commands $cs$ such that $\sigma_2=\sigma_1\bullet cs$.
+C-structs are partially ordered by the relation $\sqsubseteq$ defined such that $\sigma_1\sqsubseteq\sigma_2$ iff there exists a sequence of commands $\mathcal{C}$ such that $\sigma_2=\sigma_1\bullet \mathcal{C}$.
 Two c-structs are compatible when they have a common upper bound; when this is the case, two c-structs $\sigma_1$ and $\sigma_2$ have a least upper bound, noted $\bigsqcup\left( \sigma_1,\sigma_2 \right)$.
@@
 
-We can now state the generalized-consensus properties: non-triviality requires that any decided c-struct be of the form $\bot\bullet cs$, where $cs$ is a sequence of proposed commands; consistency requires than any two decided c-structs $\sigma_1$ and $\sigma_2$ be compatible; stability requires that when a process $p$ decides a c-struct $\sigma_1$ and time $t_1$ and $\sigma_2$ at time $t_2$, then $t_1 \leq t_2$ implies that $\sigma_1\sqsubseteq \sigma_2$; finally, liveness requires that
+We can now state the generalized-consensus properties: non-triviality requires that any decided c-struct be of the form $\bot\bullet \mathcal{C}$, where $\mathcal{C}$ is a sequence of proposed commands; consistency requires than any two decided c-structs $\sigma_1$ and $\sigma_2$ be compatible; stability requires that when a process $p$ decides a c-struct $\sigma_1$ and time $t_1$ and $\sigma_2$ at time $t_2$, then $t_1 \leq t_2$ implies that $\sigma_1\sqsubseteq \sigma_2$; finally, liveness requires that
 if a command keeps beeing proposed, then a c-struct $\sigma$ containing the command is eventually decided.

### 2016-02-12 podc16/paper2.tex
few changes in the intro, description of the generic algo, changed decide to commit, \mathcal{C} to \mathcal{D}

--- 2016-02-12 podc16/paper2.tex
@@
 
-We show that the core idea underlying EPaxos can be captured in an abstract algorithm that uses two abstractions: (1), a per-command dependency-set computation algorithm, a concept that we define in the paper, and (2), an algorithm ensuring, for each command proposed to it, agreement on a dependency set.
+We show that the core idea underlying EPaxos can be captured in a generic algorithm that uses two abstractions: (1), a per-command dependency-set computation algorithm, a concept that we define in the paper, and (2), an algorithm ensuring, for each command proposed to it, agreement on a dependency set.
 Both abstractions lend themselves well to implementations that avoid the use of a unique leader process.
@@
 
-We prove that all the algorithms in the family are safe and live, and we have formalized our ideas in TLA+ and thoroughly model-check our claims with the TLC model-checker.
+We prove that all the algorithms in the family are safe and live, and we have formalized our ideas in TLA+ and have thoroughly model-checked our safety claims with the TLC model-checker.
 
@@
 Briefly, the c-structs are a set of elements, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct.
-C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $cs$ such that $\sigma_2 = \sigma_1\bullet cs$ and $\bot\sqsubseteq \sigma$ for any
+C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $\mathcal{D}$ such that $\sigma_2 = \sigma_1\bullet \mathcal{D}$ and $\bot\sqsubseteq \sigma$ for any
 $\sigma$. Two c-structs are compatible when they have a
@@
 A process $p$ receiving a command $c$ proceeds in three phases.
-First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $cs$ \todo{are we sure about the notation cs?} that the command $c$ may depend on.
-Second, the process $p$ proposes the mapping  $c\mapsto cs$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto cs'$ for $c$, in which case $cs'$ becomes the set of dependencies of $c$, and we say that $c$ is executable when its dependency set is empty or when all its dependencies are executable.
-Finally, when $c$ is executable, the process $p$ may execute $c$ by deciding the c-struct obtained by running the \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
+First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $\mathcal{D}$ that the command $c$ may depend on.
+Second, $p$ proposes the mapping  $c\mapsto \mathcal{D}$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto \mathcal{D}'$ for $c$, in which case $\mathcal{D}'$ becomes the dependency set of $c$;
+we say that $c$ is executable when all its transitive dependencies are known.
+%no command $c'$ for which no mapping has been committed is reachable from $c$ by recursively following committed mappings.
+\todo{Change the name of the the decide action to commit.}
+Finally, when $c$ is executable, the $p$ may execute $c$ by deciding the c-struct obtained by \emph{locally} running a \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
 Any other process can run the three phases independently of $p$ and eventually execute the command despite a crash of $p$.
@@
 
-\todo{Why not add the decide call to the interface, in order to guarantee liveness?}
-
-A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and \todo{this is not clear: an input interface $\opn{decide}\left( c,\mathcal{D} \right)$ (meant to receive inputs from the agreement sub-algorithm)}, where $p\in P$ is the process executing the call.
+A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and \todo{this is not clear: an input interface $\opn{commit}\left( c,\mathcal{D} \right)$ (meant to receive inputs from the agreement sub-algorithm)}, where $p\in P$ is the process executing the call.
 Moreover, a dependency-set algorithm must ensure that:
@@
     \item For any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$ (with $p$ possibly equal to $q$), if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
-    \item If a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{decide}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
+    \item If a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
 \end{enumerate}
@@
 			\Statex
-			\START[$\opn{decide}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
+			\START[$\opn{commit}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
 			\State  $m_p \gets m_p \oplus \left[ c \mapsto \mathcal{D}\right]$	
@@
 
-A map-agreement algorithm exposes the input interface $\opn{propose}_p\left( c, \mathcal{D} \right)$, to propose a set of dependencies $\mathcal{D}$ for a command $c$, and the output interface $\opn{decide}_p\left( c, \mathcal{D} \right)$, to decide $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
+A map-agreement algorithm exposes the input interface $\opn{propose}_p\left( c, \mathcal{D} \right)$, to propose a set of dependencies $\mathcal{D}$ for a command $c$, and the output interface $\opn{commit}_p\left( c, \mathcal{D} \right)$, to commit $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
 Moreover, a map-agreement algorithm must ensure that: 
@@
 \begin{enumerate}[noitemsep,nolistsep]
-    \item If the set of dependencies $\mathcal{D}$ is decided for a command $c$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
-    \item If $\mathcal{D}_1$ is decided for a command $c$ and $\mathcal{D}_2$ is decided for the same command $c$, then $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
+    \item If the set of dependencies $\mathcal{D}$ is committed for a command $c$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
+    \item If $\mathcal{D}_1$ is committed for a command $c$ and $\mathcal{D}_2$ is committed for the same command $c$, then $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
  \end{enumerate}
@@
 \begin{enumerate}[noitemsep,nolistsep]
-    \item For any decision $\opn{decide}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
-    \item For any two decisions $\opn{decide}_p\left( c, \mathcal{D}_1 \right)$, $\opn{decide}_q\left( c, \mathcal{D}_2 \right)$ (with $p$ possibly equal to $q$), then $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
+    \item For any decision $\opn{commit}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
+    \item For any two decisions $\opn{commit}_p\left( c, \mathcal{D}_1 \right)$, $\opn{commit}_q\left( c, \mathcal{D}_2 \right)$ (with $p$ possibly equal to $q$), then $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
 
@@
 Its key property is \cref{lemma1}. 
-We assume that processes initially agree, for each set of commands $cs$, on a total order $<_{cs}$ on $cs$.
-In practice, if commands are attached a unique identifier among a totally ordered set, $<_{cs}$ can be the restriction of this total order on $cs$.
+We assume that processes initially agree, for each set of commands $\mathcal{D}$, on a total order $<_{\mathcal{D}}$ on $\mathcal{D}$.
+In practice, if commands are attached a unique identifier among a totally ordered set, $<_{\mathcal{D}}$ can be the restriction of this total order on $\mathcal{D}$.
 
 %\textbf{Local dependency graph computed by a process.} 
-We say that a command $c$ in the domain of $m_p$ is \emph{complete} when no command $c$ which is not in the domain of $m_p$ can be reached from $c$ by recursively following $c$'s dependencies.
-The local variable $m_p$ denotes a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the complete commands of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ iff $c_2 \in m_p\left[ c_1 \right]$ (i.e., $c_1$ depends on $c_2$).
+We say that a command $c$ in the domain of $m_p$ is \emph{executable} when no command $c$ which is not in the domain of $m_p$ can be reached from $c$ by recursively following $c$'s dependencies.
+The local variable $m_p$ denotes a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the executable commands of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ iff $c_2 \in m_p\left[ c_1 \right]$ (i.e., $c_1$ depends on $c_2$).
 For example, if $m_p = \left[ c_1 \mapsto \aset{c_2,c_4}, c_2 \mapsto \aset{c_3}, c_3 \mapsto \aset{c_2}\right]$ then $V\left( g_p \right) = \aset{c_2,c_3}$ and $E\left( g_p \right) = \aset{\left( c_2,c_3 \right),\left( c_3,c_2 \right)}$.
@@
 %\textbf{Linearizations of a dependency graph.}
-We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1$ and $c_2$ belong to the same strongly connected component $cs$ and $ c_1 <_{cs} c_2$ hold, then $c_1 \ll c_2$.
+We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1$ and $c_2$ belong to the same strongly connected component $\mathcal{D}$ and $ c_1 <_{\mathcal{D}} c_2$ hold, then $c_1 \ll c_2$.
 For example, assuming that $c_2 \leq_{\aset{c_1,c_2}} c_1$, the linearizations of $h$ are $\aseq{c_2,c_1,c_3,c_4}$, $\aseq{c_2,c_1,c_4,c_3}$, $\aseq{c_2,c_4,c_1,c_3}$, and $\aseq{c_4,c_2,c_1,c_3}$.
@@
 
-\textbf{Linearization function.} Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ cs \right]$ is a total order on the set $cs$.
+\textbf{Linearization function.} Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ \mathcal{D} \right]$ is a total order on the set $\mathcal{D}$.
 
@@
 
-Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ cs \right]$ contains each command in $cs$ exactly once.
+Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ \mathcal{D} \right]$ contains each command in $\mathcal{D}$ exactly once.
 A sequence of commands $cseq$ is an $f$-linearization of the graph $g$ if and only there is a sequence whose element of strongly connected
@@
 We also write the $\sigma\bullet\aseq{c_1,\cdots,c_n}$ for the c-struct $\sigma\bullet c_1\bullet\cdots\bullet c_n$.
-C-structs are partially ordered by the relation $\sqsubseteq$ defined such that $\sigma_1\sqsubseteq\sigma_2$ iff there exists a sequence of commands $cs$ such that $\sigma_2=\sigma_1\bullet cs$.
+C-structs are partially ordered by the relation $\sqsubseteq$ defined such that $\sigma_1\sqsubseteq\sigma_2$ iff there exists a sequence of commands $\mathcal{D}$ such that $\sigma_2=\sigma_1\bullet \mathcal{D}$.
 Two c-structs are compatible when they have a common upper bound; when this is the case, two c-structs $\sigma_1$ and $\sigma_2$ have a least upper bound, noted $\bigsqcup\left( \sigma_1,\sigma_2 \right)$.
@@
 
-We can now state the generalized-consensus properties: non-triviality requires that any decided c-struct be of the form $\bot\bullet cs$, where $cs$ is a sequence of proposed commands; consistency requires than any two decided c-structs $\sigma_1$ and $\sigma_2$ be compatible; stability requires that when a process $p$ decides a c-struct $\sigma_1$ and time $t_1$ and $\sigma_2$ at time $t_2$, then $t_1 \leq t_2$ implies that $\sigma_1\sqsubseteq \sigma_2$; finally, liveness requires that
+We can now state the generalized-consensus properties: non-triviality requires that any decided c-struct be of the form $\bot\bullet \mathcal{D}$, where $\mathcal{D}$ is a sequence of proposed commands; consistency requires than any two decided c-structs $\sigma_1$ and $\sigma_2$ be compatible; stability requires that when a process $p$ decides a c-struct $\sigma_1$ and time $t_1$ and $\sigma_2$ at time $t_2$, then $t_1 \leq t_2$ implies that $\sigma_1\sqsubseteq \sigma_2$; finally, liveness requires that
 if a command keeps beeing proposed, then a c-struct $\sigma$ containing the command is eventually decided.

### 2016-02-12 podc16/paper2.tex
formatting changes

@@
 
-We now present our abstract leaderless generalized consensus algorithm.
+Our leaderless generalized consensus algorithm orchestrates a dependency-set sub-algorithm and map-agreement sub-algorithm as shown in \cref{fig:deps-set-algo3}.
 A process $p$ receiving a command $c$ proceeds in three phases.
@@
 %no command $c'$ for which no mapping has been committed is reachable from $c$ by recursively following committed mappings.
-\todo{Change the name of the the decide action to commit.}
 Finally, when $c$ is executable, the $p$ may execute $c$ by deciding the c-struct obtained by \emph{locally} running a \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
@@
 However, we show that a leaderless  generalized consensus algorithm is always safe and is live when its underlying dependency-set and map-agreement algorithms are safe and live.
-The leaderless generalized consensus algorithm orchestrates the two dependency-set and map-agreement sub-algorithms as shown in \cref{fig:deps-set-algo3}.\todo{This pseudocode must be in figure (a).
-Also we have to better describe the pseudocode.}
 
@@
 
-A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and \todo{this is not clear: an input interface $\opn{commit}\left( c,\mathcal{D} \right)$ (meant to receive inputs from the agreement sub-algorithm)}, where $p\in P$ is the process executing the call.
+A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{commit}\left( c,\mathcal{D} \right)$ (meant to observe the map-agreement sub-algorithm), where $p\in P$ is the process executing the call.
 Moreover, a dependency-set algorithm must ensure that:
-\begin{enumerate}[noitemsep,nolistsep]
-    \item For any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before.
-    \item For any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$ (with $p$ possibly equal to $q$), if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
-    \item If a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
-\end{enumerate}
-
-An example of implementation of such an algorithm is depicted by the pseudocode of Figures~\ref{fig:deps-set-algo1} and~\ref{fig:deps-set-algo2}.
-The idea is to let every process broadcast $c$ to all the processes in the system, in order to announce $c$ (lines \ref{alg:announce}--\ref{alg:send-new-cmd}).
-Then whenever a process $p$ receives $c$, it replies to the sender with the set $\mathcal{D}$ of commands that $p$ has observed so far, and such that $c$ and any command in $\mathcal{D}$ do not commute (lines \ref{alg:receive-new-cmd}--\ref{alg:send-deps}).
-Afterwards, the process announcing $c$ collects the sets $\mathcal{D}$ for $c$ from a \textit{quorum} of $\mathcal{Q}$ processes, and it locally suggests the union of those sets as final dependency-set for $c$ (lines \ref{alg:receive-deps}--\ref{alg:call-suggest}).
-For the purpose of this example we allow for $\mathcal{Q}$ to be equal to any positive integer that is not greater than the number of processes, as long as the following requirement is met: the intersection of any two quorums that are selected by processes to collect the sets of dependencies is not empty.
-
-The provided implementation satisfies the specification of a dependency-set algorithm for the following reasons.
-Since any command in a dependency-set that is suggested by a process (lines~\ref{alg:update-deps-p} and~\ref{alg:call-suggest}) must be in the set $cmds_q$ of the commands received by some process $q$ (line~\ref{alg:update-d}), and a process receives a command if and only if there exists some process that previously announced the command (line~\ref{alg:send-new-cmd}), then it follows that any command in any suggested dependency-set has been previously announced.
-Further, since a process can suggest a set of dependencies $deps_p[c]$ for a command $c$ after it has received a reply for $c$ from at least a quorum of processes (lines~\ref{alg:receive-deps}--\ref{alg:call-suggest}), we have that $\opn{suggest}_p \left( c,deps_p[c]\right)$ is called only if a quorum of processes $q$ have $c$ in their $cmds_q$ (line~\ref{alg:update-cmds-p}).
-Therefore, by the requirement that we enforced on the quorums, we can conclude that for any two suggestions $\opn{suggest}_p \left( c_1,deps_p[c_1]\right)$ and $\opn{suggest}_q \left( c_2,deps_q[c_2]\right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_q[c_2]$ or $c_2 \in deps_p[c_1]$.
+\begin{itemize}[noitemsep,nolistsep]
+    \item[-] Safety: (a) For any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before, and (b), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
+%    \item If $c$ was announced by $p$, then eventually a dependency set is suggested for $c$ to $p$.
+    \item[-] Liveness: (a) If a suggestion $\opn{suggest}_p\left( c \right)$ was made, then eventually a suggestion $\opn{suggest}\left( c,\mathcal{D} \right)_p$ is made, and (b), if a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
+\end{itemize}
+\vspace{5cm}
 
@@
 \centering
+\subfigure[Leaderless generalized consensus implementation.]{
+\noindent\fbox{%
+	\begin{varwidth}{\dimexpr\linewidth-2\fboxsep-2\fboxrule\relax}
+
+		\begin{algorithmic}[1]
+			{\scriptsize
+			\START[$\opn{init}_p$]{{\bf upon}}{$()$}
+			\State $m_p\gets \emptyset$
+			\State $\sigma_p\gets \bot$
+			\END
+			\START[$\opn{GC-propose}_p$]{{\bf upon}}{$(c)$}
+			\State {\bf call} $\opn{announce}_p\left( c \right)$
+			\END
+			\Statex
+			\START[$\opn{suggest}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
+			\State  $\opn{propose}_p( c,\mathcal{D})$ 			
+			\END
+			\Statex
+			\START[$\opn{commit}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
+			\State  $m_p \gets m_p \oplus \left[ c \mapsto \mathcal{D}\right]$	
+			\State $\sigma_p \gets  \opn{c-struct}\left(m_p\right)$
+			\State $\opn{GC-decide}_p(\sigma_p)$		
+			\END
+			}
+		\end{algorithmic}
+	\end{varwidth}%
+}
+\label{fig:deps-set-algo3}
+}
 \subfigure[$\texttt{init}_p$ and $\texttt{announce}_p$ implementations.]{
@@
 }
-\hspace*{0.1in}\subfigure[Leaderless generalized consensus implementation.]{
-\noindent\fbox{%
-	\begin{varwidth}{\dimexpr\linewidth-2\fboxsep-2\fboxrule\relax}
-
-		\begin{algorithmic}[1]
-			{\scriptsize
-			\START[$\opn{init}_p$]{{\bf upon}}{$()$}
-			\State $m_p\gets \emptyset$
-			\State $\sigma_p\gets \bot$
-			\END
-			\START[$\opn{GC-propose}_p$]{{\bf upon}}{$(c)$}
-			\State {\bf call} $\opn{announce}_p\left( c \right)$
-			\END
-			\Statex
-			\START[$\opn{suggest}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
-			\State  $\opn{propose}_p( c,\mathcal{D})$ 			
-			\END
-			\Statex
-			\START[$\opn{commit}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
-			\State  $m_p \gets m_p \oplus \left[ c \mapsto \mathcal{D}\right]$	
-			\State $\sigma_p \gets  \opn{c-struct}\left(m_p\right)$
-			\State $\opn{GC-decide}_p(\sigma_p)$		
-			\END
-			}
-		\end{algorithmic}
-	\end{varwidth}%
-}
-\label{fig:deps-set-algo3}
-}
 \caption{Example of dependency-set algorithm and leaderless generalized consensus implementations.}
@@
 Moreover, a map-agreement algorithm must ensure that:
-\begin{comment}
-\begin{enumerate}[noitemsep,nolistsep]
-    \item If the set of dependencies $\mathcal{D}$ is committed for a command $c$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
-    \item If $\mathcal{D}_1$ is committed for a command $c$ and $\mathcal{D}_2$ is committed for the same command $c$, then $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
- \end{enumerate}
-
-\todo{OR MORE FORMALLY}
-
-\end{comment}
 
@@
 %used to implementing map-agreement by uniquely associating commands to MultiPaxos instance numbers.
-
-\todo{Cite Generalized Paxos as an example for bounding the data structures. Problem: we do not describe any particular data structure here. It is not clear why we should write about GC here.}
 
@@
 
+\subsection{Description map-agreement implementation}
+
+An example of implementation of such an algorithm is depicted by the pseudocode of Figures~\ref{fig:deps-set-algo1} and~\ref{fig:deps-set-algo2}.
+The idea is to let every process broadcast $c$ to all the processes in the system, in order to announce $c$ (lines \ref{alg:announce}--\ref{alg:send-new-cmd}).
+Then whenever a process $p$ receives $c$, it replies to the sender with the set $\mathcal{D}$ of commands that $p$ has observed so far, and such that $c$ and any command in $\mathcal{D}$ do not commute (lines \ref{alg:receive-new-cmd}--\ref{alg:send-deps}).
+Afterwards, the process announcing $c$ collects the sets $\mathcal{D}$ for $c$ from a \textit{quorum} of $\mathcal{Q}$ processes, and it locally suggests the union of those sets as final dependency-set for $c$ (lines \ref{alg:receive-deps}--\ref{alg:call-suggest}).
+For the purpose of this example we allow for $\mathcal{Q}$ to be equal to any positive integer that is not greater than the number of processes, as long as the following requirement is met: the intersection of any two quorums that are selected by processes to collect the sets of dependencies is not empty.
+
+The provided implementation satisfies the specification of a dependency-set algorithm for the following reasons.
+Since any command in a dependency-set that is suggested by a process (lines~\ref{alg:update-deps-p} and~\ref{alg:call-suggest}) must be in the set $cmds_q$ of the commands received by some process $q$ (line~\ref{alg:update-d}), and a process receives a command if and only if there exists some process that previously announced the command (line~\ref{alg:send-new-cmd}), then it follows that any command in any suggested dependency-set has been previously announced.
+Further, since a process can suggest a set of dependencies $deps_p[c]$ for a command $c$ after it has received a reply for $c$ from at least a quorum of processes (lines~\ref{alg:receive-deps}--\ref{alg:call-suggest}), we have that $\opn{suggest}_p \left( c,deps_p[c]\right)$ is called only if a quorum of processes $q$ have $c$ in their $cmds_q$ (line~\ref{alg:update-cmds-p}).
+Therefore, by the requirement that we enforced on the quorums, we can conclude that for any two suggestions $\opn{suggest}_p \left( c_1,deps_p[c_1]\right)$ and $\opn{suggest}_q \left( c_2,deps_q[c_2]\right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_q[c_2]$ or $c_2 \in deps_p[c_1]$.
+
 \end{document}

### 2016-02-12 podc16/paper2.tex
more compact items

@@
 
-\subsection{Agreeing on Dependency Sets}
-\label{map-algo}
-
+\subsection{Agreeing on Dependency Sets}\label{map-algo}
 A map-agreement algorithm exposes the input interface $\opn{propose}_p\left( c, \mathcal{D} \right)$, to propose a set of dependencies $\mathcal{D}$ for a command $c$, and the output interface $\opn{commit}_p\left( c, \mathcal{D} \right)$, to commit $\mathcal{D}$ for $c$, where $p\in P$ is the process executing the call.
@@
 \begin{enumerate}[noitemsep,nolistsep]
-    \item For any decision $\opn{commit}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
-    \item For any two decisions $\opn{commit}_p\left( c, \mathcal{D}_1 \right)$, $\opn{commit}_q\left( c, \mathcal{D}_2 \right)$ (with $p$ possibly equal to $q$), then $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
+    \item[-] For any decision $\opn{commit}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
+    \item[-] For any two decisions $\opn{commit}_p\left( c, \mathcal{D}_1 \right)$, $\opn{commit}_q\left( c, \mathcal{D}_2 \right)$, $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
 

### 2016-02-12 podc16/paper2.tex
added again a bit more to the presentation of c-structs (needed later in the paper)

@@
 Each process receives proposals for commands of the form $\opn{Propose}(c)$, and must produce responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct.
-The sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties described below.
-We refer the reader to~\cite{Lamport05GeneralizeConsensus}, and we use the same notation, for the definition of c-structs and generalized consensus.
-Briefly, the c-structs are a set of elements, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct.
+The sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
+We use the same notation and refer the reader to~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of c-structs and generalized consensus.
+Briefly, the c-structs are a set, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct.
+Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands.
+We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
 C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $\mathcal{D}$ such that $\sigma_2 = \sigma_1\bullet \mathcal{D}$ and $\bot\sqsubseteq \sigma$ for any
 $\sigma$. Two c-structs are compatible when they have a
-common upper bound.
-Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands.
-The consistency property of generalized consensus requires that any two decided c-structs be compatible.
+common upper bound.  The consistency property of generalized consensus requires that any two decided c-structs be compatible.
 
@@
 \begin{itemize}[noitemsep,nolistsep]
-    \item[-] Safety: (a) For any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before, and (b), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
+    \item[-] Safety: (a), for any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before; (b), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
 %    \item If $c$ was announced by $p$, then eventually a dependency set is suggested for $c$ to $p$.

### 2016-02-12 podc16/paper2.tex
added a liveness property to map-agreement

@@
 \begin{enumerate}[noitemsep,nolistsep]
-    \item[-] For any decision $\opn{commit}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time.
-    \item[-] For any two decisions $\opn{commit}_p\left( c, \mathcal{D}_1 \right)$, $\opn{commit}_q\left( c, \mathcal{D}_2 \right)$, $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
-
+    \item[-] Safety: (a), for any decision $\opn{commit}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time; (b), for any two decisions $\opn{commit}_p\left( c, \mathcal{D}_1 \right)$, $\opn{commit}_q\left( c, \mathcal{D}_2 \right)$, $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
+    \item[-] Liveness: if a proposal $propose_p\left( c,\mathcal{D} \right)$ is made, then eventually a decision $\opn{commit}_p\left( c, \mathcal{D} \right)$ is made.
 \end{enumerate}
@@
 
-\begin{comment}
-Local variables: $m_p$, a map from commands to dependency sets, initially the empty map; $\sigma_p$, a c-struct, initially $\bot$.
-\begin{itemize}[noitemsep,nolistsep]
-    \item Upon ${\opn{propose}\left( c \right)}^p$: call ${\opn{announce}\left( c \right)}^p$.
-    \item Upon ${\opn{suggest}\left( c,d \right)}^p$: call ${\opn{propose}\left( c,d \right)}^p$.
-    \item Upon ${\opn{decide}\left( c,d \right)}^p$: insert the mapping $c \rightarrow d$ in the map $m_p$.
-    \item Periodically, update $\sigma_p$ to $\opn{c-struct}\left( m_p \right)$, where $\opn{c-struct}\left( m_p \right)$ is defined below.
-\end{itemize}
-
-\end{comment}
-
 The graph processing algorithm is executed locally by a node (line 10 of \cref{fig:deps-set-algo3}). It must compute the value of $\opn{c-struct}\left( g_p \right)$, as defined in \cref{lindef}.
@@
 We assume that processes initially agree, for each set of commands $\mathcal{D}$, on a total order $<_{\mathcal{D}}$ on $\mathcal{D}$.
-In practice, if commands are attached a unique identifier among a totally ordered set, $<_{\mathcal{D}}$ can be the restriction of this total order on $\mathcal{D}$.
+In practice, if commands are attached a unique identifier among a totally ordered set, $<_{\mathcal{D}}$ can be the restriction of this total order to $\mathcal{D}$.
 
-%\textbf{Local dependency graph computed by a process.} 
 We say that a command $c$ in the domain of $m_p$ is \emph{executable} when no command $c$ which is not in the domain of $m_p$ can be reached from $c$ by recursively following $c$'s dependencies.
@@
 
-%\textbf{Orders induced by a graph.}
 A directed graph $g$ \emph{induces a partial order} $\preceq_g$ on its vertices defined such that $c_1 \preceq_g c_2$ iff there is a path from $c_1$ to $c_2$ and none from $c_2$ to $c_1$.
-%Let $\opn{TOs}\left( g \right)$ be the set of total orders $o$ such that $\preceq_g \subseteq o$.
 For example, consider $h$ where $V\left( h \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$. 
 We have that $\preceq_h = \aset{\left( c_1,c_3 \right)}$.
-%If $c\in V\left( g \right)$, we write $\left[ c \right]_g$ for the strongly connected component of $g$ to which $c$ belongs. 
-%\textbf{Linearizations of a dependency graph.}
 We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1$ and $c_2$ belong to the same strongly connected component $\mathcal{D}$ and $ c_1 <_{\mathcal{D}} c_2$ hold, then $c_1 \ll c_2$.
 For example, assuming that $c_2 \leq_{\aset{c_1,c_2}} c_1$, the linearizations of $h$ are $\aseq{c_2,c_1,c_3,c_4}$, $\aseq{c_2,c_1,c_4,c_3}$, $\aseq{c_2,c_4,c_1,c_3}$, and $\aseq{c_4,c_2,c_1,c_3}$.
-%\textbf{Partial order on graphs.}
 
@@
 
-\begin{comment}
-
-\textbf{Graph of strongly connected components.} Given a graph $g$, let $\hat{g}$ be the graph whose vertices are the strongly connected components of $g$ (represented as sets of vertices) and where there is an edge from $g_1$ to $g_2$ if and only if there are two vertices $v_1$ and $v_2$ of $g$ such that $v_1 \in V\left( g_1 \right)$, $v_2 \in V\left( g_2 \right)$, and $\left( v_1,v_2 \right)\in E\left( g \right)$.
-Note that the reflexive and transitive closure of $V\left(  \hat{g} \right)$ is necessarily a partial order.
-For example, consider $h$ such that $V\left( h \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$.
-Then $V(\hat{h}) = \aset{\aset{c_1,c_2},\aset{c_3},\aset{c_4}}$ and $E( \hat{h} )=\aset{\left( \aset{c_1,c_2},\aset{c_3} \right)}$.
-We will use $h$ as an example in the rest of the section.
-
-\textbf{Linearization function.} Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ \mathcal{D} \right]$ is a total order on the set $\mathcal{D}$.
-
-\textbf{Linearizations of a dependency graph.}
-When $\hat{o}$ is a total order over the strongly connected components of $g$ such that $\hat{o} \in \opn{TOs}\left( \hat{g} \right)$, we say that a total order $l$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when $\left(v_1,v_2\right) \in l$ iff (1) there exists $vs_1,vs_2\in V\left(\hat{g}\right)$ such that $v_1\in vs_1$, $v_2\in vs_2$, and $\left( vs_1,vs_2 \right)\in \hat{o}$, or (2) there exists $vs_1\in V\left(\hat{g}\right)$ such that
-$v_1,v_2\in  vs_1$ and $\left( v_1,v_2 \right)\in f\left[  vs_1 \right]$.
-For example, if $f\left[ \aset{c_1,c_2} \right] = \aseq{c_2,c_1}$ then $\aseq{c_2,c_1,c_3,c_4}$ is a linearization of $h$ but $\aseq{c_1,c_2,c_3,c_4}$ is not and neither is $\aseq{c_3,c_2,c_1,c_4}$
-
-\end{comment}
-
-\begin{comment}
-Define $\opn{Lin}_g\left( \leq \right) = f\left[ V_1 \right] \cdot F\left[ V_2 \right] \cdot F\left[ V_n \right]$, where $V_i\in V\left( \hat{g} \right)$
-
-Define the set of linearizations $\opn{Lins}\left( g \right)$ of a graph $g$ as the set of sequences whose length is the cardinality of $V\left( g \right)$, which contain each element of $V\left( g \right)$ exactly once, and ...
-\todo{or first define the partial order induced by a graph and then its compatible total orders.}
-
-Consider a function $f$ mapping sets of commands to sequences of commands such that $f\left[ \mathcal{D} \right]$ contains each command in $\mathcal{D}$ exactly once.
-A sequence of commands $cseq$ is an $f$-linearization of the graph $g$ if and only there is a sequence whose element of strongly connected
-
-\todo{Explain the graph processing algo independently? Its key property is that when $G1 \subseteq G2$ then $lin\left( c_1,G1 \right)$ is a prefix of $lin\left( c_2,G2 \right)$ or vice-versa.}
-
-\end{comment}
-
 \subsection{Consistency and Liveness}
 
-\textbf{Lemma 2}
+\begin{mylemma}\label{lemma-compat}
 If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible.
+\end{mylemma}
+\Cref{lemma-compat} follows from the safety property (b) of the map-agreement algorithm and definition of the graph $g_p$.
+From (b) it follows that if $c$ is in the domain of both $m_p$ and $m_q$ then $m_p\left[ c \right] = m_q\left[ c \right]$.
 
@@
 The provided implementation satisfies the specification of a dependency-set algorithm for the following reasons.
-Since any command in a dependency-set that is suggested by a process (lines~\ref{alg:update-deps-p} and~\ref{alg:call-suggest}) must be in the set $cmds_q$ of the commands received by some process $q$ (line~\ref{alg:update-d}), and a process receives a command if and only if there exists some process that previously announced the command (line~\ref{alg:send-new-cmd}), then it follows that any command in any suggested dependency-set has been previously announced.
+Since any command in a dependency-set that is suggested by a process (lines~\ref{alg:update-deps-p} and~\ref{alg:call-suggest}) must be in the set $cmds_q$ of the commands received by some process $q$ (line~\ref{alg:update-d}), and a process receives a command iff there exists some process that previously announced the command (line~\ref{alg:send-new-cmd}), then it follows that any command in any suggested dependency-set has been previously announced.
 Further, since a process can suggest a set of dependencies $deps_p[c]$ for a command $c$ after it has received a reply for $c$ from at least a quorum of processes (lines~\ref{alg:receive-deps}--\ref{alg:call-suggest}), we have that $\opn{suggest}_p \left( c,deps_p[c]\right)$ is called only if a quorum of processes $q$ have $c$ in their $cmds_q$ (line~\ref{alg:update-cmds-p}).

### 2016-02-12 podc16/paper2.tex
started writing proofs, made them compact

@@
 The graph processing algorithm is executed locally by a node (line 10 of \cref{fig:deps-set-algo3}). It must compute the value of $\opn{c-struct}\left( g_p \right)$, as defined in \cref{lindef}.
-Its key property is \cref{lemma1}. 
+Its key property is \cref{lemma:graph}. 
 We assume that processes initially agree, for each set of commands $\mathcal{D}$, on a total order $<_{\mathcal{D}}$ on $\mathcal{D}$.
@@
 
-\begin{mylemma}\label{lemma1}
+\begin{mylemma}\label{lemma:graph}
 Assume that $g_1$ and $g_2$ are two dependency graphs, 
@@
 \subsection{Consistency and Liveness}
-
-\begin{mylemma}\label{lemma-compat}
+Let us now sketch why our generic leaderless algorithm meets the consistency and liveness requirements of generalized consensus. Consider an execution of the algorithm.
+\begin{mylemma}\label{lemma:compat}
 If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible.
 \end{mylemma}
-\Cref{lemma-compat} follows from the safety property (b) of the map-agreement algorithm and definition of the graph $g_p$.
+%\vspace{-5pt}
+\begin{proof*}
+Follows from the safety property (b) of the map-agreement algorithm and definition of the graph $g_p$.
 From (b) it follows that if $c$ is in the domain of both $m_p$ and $m_q$ then $m_p\left[ c \right] = m_q\left[ c \right]$.
+\end{proof*}
+%\vspace{-5pt}
+\begin{mylemma}
+    The consistency property always holds.
+    \label{lemma:consistency}
+\end{mylemma}
+%\vspace{-5pt}
+\begin{proof*}
+    By \cref{lemma:compat} and the safety property (b) of the dependency-set algorithm, $g_p^1$ and $g_q^2$ satisfy the premises of \cref{lemma:graph}.
+Therefore we get that $g_p^1$ and $g_q^2$ are compatible.
+\end{proof*}
 

### 2016-02-12 podc16/paper2.tex
new description of dependency-set algo

--- 2016-02-12 podc16/paper2.tex
@@
 Each process receives proposals for commands of the form $\opn{Propose}(c)$, and must produce responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct.
-The sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
-We use the same notation and refer the reader to~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of c-structs and generalized consensus.
-Briefly, the c-structs are a set, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct.
-Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands.
-We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
+The sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties described below.
+We refer the reader to~\cite{Lamport05GeneralizeConsensus}, and we use the same notation, for the definition of c-structs and generalized consensus.
+Briefly, the c-structs are a set of elements, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct.
 C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $\mathcal{D}$ such that $\sigma_2 = \sigma_1\bullet \mathcal{D}$ and $\bot\sqsubseteq \sigma$ for any
 $\sigma$. Two c-structs are compatible when they have a
-common upper bound.  The consistency property of generalized consensus requires that any two decided c-structs be compatible.
+common upper bound.
+Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands.
+The consistency property of generalized consensus requires that any two decided c-structs be compatible.
 
@@
 \begin{itemize}[noitemsep,nolistsep]
-    \item[-] Safety: (a), for any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before; (b), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
+    \item[-] Safety: (a) For any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before, and (b), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
 %    \item If $c$ was announced by $p$, then eventually a dependency set is suggested for $c$ to $p$.
@@
 \end{itemize}
-\vspace{5cm}
+
+An example of implementation of such an algorithm is depicted by the pseudocode of Figures~\ref{fig:deps-set-algo1} and~\ref{fig:deps-set-algo2}.
+Due to space constraints, our example focuses on satisfying the safety property only.
+The main idea is the following: a process $p$ announcing a command $c$ broadcasts $c$ to all the processes, and it suggests the dependency-set $deps_p[c]$ for $c$ as the union of the sets $\mathcal{D}$ that it receives as replies from a quorum of $\mathcal{Q}$ processes; on the other hand, a process that receives an announcement for $c$, computes its reply $\mathcal{D}$ for $c$ as the set of commands $d$ that it received so far, and such that $c$ and $d$ do not commute.
+The choice of the value of $\mathcal{Q}$ must be such that the intersection of any two quorums is not empty.
+About correctness, we can easily prove that such an implementation satisfies the safety property by noticing that:  any process $p$ inserts a command $d$ in a dependency-set $\mathcal{D}$ that $p$ sent as a reply for a command $c$, only if $p$ previously received $d$.
+Therefore, every command suggested in a dependency-set must have been announced before.
+Further, since a process can suggest a set of dependencies $deps_p[c]$ for a command $c$ after it has received a reply for $c$ from at least a quorum of processes, and thanks to the requirement that we enforced on the intersection of quorums, we can conclude that for any two suggestions $\opn{suggest}_p \left( c_1,deps_p[c_1]\right)$ and $\opn{suggest}_q \left( c_2,deps_q[c_2]\right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_q[c_2]$ or $c_2 \in deps_p[c_1]$.
 

### 2016-02-12 podc16/paper2.tex
new ids for safety and liveness properties

@@
 \begin{itemize}[noitemsep,nolistsep]
-    \item[-] Safety: (a), for any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before; (b), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
+    \item[-] Safety: (S1), for any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before; (S2), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
 %    \item If $c$ was announced by $p$, then eventually a dependency set is suggested for $c$ to $p$.
-    \item[-] Liveness: (a) If a suggestion $\opn{suggest}_p\left( c \right)$ was made, then eventually a suggestion $\opn{suggest}\left( c,\mathcal{D} \right)_p$ is made, and (b), if a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
+    \item[-] Liveness: (L1) If a suggestion $\opn{suggest}_p\left( c \right)$ was made, then eventually a suggestion $\opn{suggest}\left( c,\mathcal{D} \right)_p$ is made, and (L2), if a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
 \end{itemize}
@@
 \begin{enumerate}[noitemsep,nolistsep]
-    \item[-] Safety: (a), for any decision $\opn{commit}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time; (b), for any two decisions $\opn{commit}_p\left( c, \mathcal{D}_1 \right)$, $\opn{commit}_q\left( c, \mathcal{D}_2 \right)$, $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
-    \item[-] Liveness: if a proposal $propose_p\left( c,\mathcal{D} \right)$ is made, then eventually a decision $\opn{commit}_p\left( c, \mathcal{D} \right)$ is made.
+    \item[-] Safety: (S3), for any decision $\opn{commit}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time; (S4), for any two decisions $\opn{commit}_p\left( c, \mathcal{D}_1 \right)$, $\opn{commit}_q\left( c, \mathcal{D}_2 \right)$, $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
+    \item[-] Liveness: (L3) if a proposal $propose_p\left( c,\mathcal{D} \right)$ is made, then eventually a decision $\opn{commit}_p\left( c, \mathcal{D} \right)$ is made.
 \end{enumerate}

### 2016-02-12 podc16/paper2.tex
working on correctness

--- 2016-02-12 podc16/paper2.tex
@@
 $\sigma$. Two c-structs are compatible when they have a
-common upper bound.  The consistency property of generalized consensus requires that any two decided c-structs be compatible.
+common upper bound.
+A c-struct contains a command $c$ when it is of the form $\bot\bullet cs$ where $c$ appears in the sequence of commands $cs$.
+The consistency property of generalized consensus requires that any two decided c-structs be compatible.
 
@@
 \end{itemize}
-
-An example of implementation of such an algorithm is depicted by the pseudocode of Figures~\ref{fig:deps-set-algo1} and~\ref{fig:deps-set-algo2}.
-Due to space constraints, our example focuses on satisfying the safety property only.
-The main idea is the following: a process $p$ announcing a command $c$ broadcasts $c$ to all the processes, and it suggests the dependency-set $deps_p[c]$ for $c$ as the union of the sets $\mathcal{D}$ that it receives as replies from a quorum of $\mathcal{Q}$ processes; on the other hand, a process that receives an announcement for $c$, computes its reply $\mathcal{D}$ for $c$ as the set of commands $d$ that it received so far, and such that $c$ and $d$ do not commute.
-The choice of the value of $\mathcal{Q}$ must be such that the intersection of any two quorums is not empty.
-About correctness, we can easily prove that such an implementation satisfies the safety property by noticing that:  any process $p$ inserts a command $d$ in a dependency-set $\mathcal{D}$ that $p$ sent as a reply for a command $c$, only if $p$ previously received $d$.
-Therefore, every command suggested in a dependency-set must have been announced before.
-Further, since a process can suggest a set of dependencies $deps_p[c]$ for a command $c$ after it has received a reply for $c$ from at least a quorum of processes, and thanks to the requirement that we enforced on the intersection of quorums, we can conclude that for any two suggestions $\opn{suggest}_p \left( c_1,deps_p[c_1]\right)$ and $\opn{suggest}_q \left( c_2,deps_q[c_2]\right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_q[c_2]$ or $c_2 \in deps_p[c_1]$.
+\vspace{5cm}
 
@@
 
-\subsection{Consistency and Liveness}
-Let us now sketch why our generic leaderless algorithm meets the consistency and liveness requirements of generalized consensus. Consider an execution of the algorithm.
+\subsection{Satisfying Consistency and Liveness of Generalized Consensus}
+Consider an execution of the algorithm.
 \begin{mylemma}\label{lemma:compat}
@@
 \end{mylemma}
-%\vspace{-5pt}
 \begin{proof*}
@@
 \end{proof*}
-%\vspace{-5pt}
-\begin{mylemma}
+\begin{mythm}
     The consistency property always holds.
     \label{lemma:consistency}
-\end{mylemma}
-%\vspace{-5pt}
+\end{mythm}
 \begin{proof*}
@@
 Therefore we get that $g_p^1$ and $g_q^2$ are compatible.
+\end{proof*}
+\begin{mythm}
+    If a command $c$ which is GC-proposed is then repeatedly GC-proposed, then a c-struct containing $c$ is eventually decided.
+\end{mythm}
+\begin{proof*}
+    The liveness properties (a) and (b) of the dependency-set algorithm ensures that some dependency sets are repeatedly proposed for $c$ to the map-agreement algorithm, as long as a commit is not observed.
+Moreover, the map-agreement algorithm ensures that when dependency sets are repeatedly proposed for $c$, then a dependency set will eventually be committed for $c$.
+Similarly, any dependency of $c$ is eventually committed and $c$ becomes executable, at which points it is included in the next decided c-struct.
 \end{proof*}

### 2016-02-12 podc16/paper2.tex
fixes in the graph processing section, and added wordrefs

@@
 \begin{itemize}[noitemsep,nolistsep]
-    \item[-] Safety: (S1), for any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before; (S2), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
+    \item[-] Safety: (\setword{S1}{dep-valid}), for any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before; (\setword{S2}{suggestion-agreement}), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
 %    \item If $c$ was announced by $p$, then eventually a dependency set is suggested for $c$ to $p$.
-    \item[-] Liveness: (L1) If a suggestion $\opn{suggest}_p\left( c \right)$ was made, then eventually a suggestion $\opn{suggest}\left( c,\mathcal{D} \right)_p$ is made, and (L2), if a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
+    \item[-] Liveness: (\setword{L1}{eventual-suggestion}) If a suggestion $\opn{suggest}_p\left( c \right)$ was made, then eventually a suggestion $\opn{suggest}\left( c,\mathcal{D} \right)_p$ is made, and (\setword{L2}{re-suggest}), if a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
 \end{itemize}
@@
 \begin{enumerate}[noitemsep,nolistsep]
-    \item[-] Safety: (S3), for any decision $\opn{commit}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time; (S4), for any two decisions $\opn{commit}_p\left( c, \mathcal{D}_1 \right)$, $\opn{commit}_q\left( c, \mathcal{D}_2 \right)$, $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
-    \item[-] Liveness: (L3) if a proposal $propose_p\left( c,\mathcal{D} \right)$ is made, then eventually a decision $\opn{commit}_p\left( c, \mathcal{D} \right)$ is made.
+    \item[-] Safety: (\setword{S3}{commit-validity}), for any commit $\opn{commit}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time; (\setword{S4}{commit-agreement}), for any two decisions $\opn{commit}_p\left( c, \mathcal{D}_1 \right)$, $\opn{commit}_q\left( c, \mathcal{D}_2 \right)$, $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
+    \item[-] Liveness: (\setword{L3}{eventual-commit}) if a proposal $propose_p\left( c,\mathcal{D} \right)$ is made, then eventually a decision $\opn{commit}_p\left( c, \mathcal{D} \right)$ is made.
 \end{enumerate}
@@
 
-A directed graph $g$ \emph{induces a partial order} $\preceq_g$ on its vertices defined such that $c_1 \preceq_g c_2$ iff there is a path from $c_1$ to $c_2$ and none from $c_2$ to $c_1$.
+A directed graph $g$ \emph{induces a partial order} $\preceq_g$ on its vertices defined such that $c_1 \preceq_g c_2$ iff there is a path from $c_2$ to $c_1$ and none from $c_1$ to $c_2$.
 For example, consider $h$ where $V\left( h \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$. 
-We have that $\preceq_h = \aset{\left( c_1,c_3 \right)}$.
+We have that $\preceq_h = \aset{\left( c_3,c_1 \right)}$.
 We say that a total order $\ll$ on $V\left( g \right)$ is a \emph{linearization} of $g$ when for every $c_1,c_2\in V\left( g \right)$, $c_1 \preceq_g c_2$ implies $c_1 \ll c_2$, and if $c_1$ and $c_2$ belong to the same strongly connected component $\mathcal{D}$ and $ c_1 <_{\mathcal{D}} c_2$ hold, then $c_1 \ll c_2$.
-For example, assuming that $c_2 \leq_{\aset{c_1,c_2}} c_1$, the linearizations of $h$ are $\aseq{c_2,c_1,c_3,c_4}$, $\aseq{c_2,c_1,c_4,c_3}$, $\aseq{c_2,c_4,c_1,c_3}$, and $\aseq{c_4,c_2,c_1,c_3}$.
+For example, assuming that $c_1 <_{\aset{c_1,c_2}} c_2$, the linearizations of $h$ are $\aseq{c_3,c_1,c_2,c_4}$, $\aseq{c_3,c_1,c_4,c_2}$, $\aseq{c_3,c_4,c_1,c_2}$,
+$\aseq{c_4,c_3,c_1,c_2}$.
 

### 2016-02-12 podc16/paper2.tex
minor changes to pseudocode

@@
 }
-\subfigure[$\texttt{init}_p$ and $\texttt{announce}_p$ implementations.]{
+\hspace*{0.1in}\subfigure[Example of dependency-set algorithm implementation - part 1.]{
 \noindent\fbox{%
@@
 }	
-\hspace*{0.1in}\subfigure[receive of \texttt{cmd} and \texttt{deps} messages, and suggestion of dependencies for a command on process $p$.]{
+\hspace*{0.1in}\subfigure[Example of dependency-set algorithm implementation - part 2.]{
 \noindent\fbox{%
@@
 }
-\caption{Example of dependency-set algorithm and leaderless generalized consensus implementations.}
+\caption{Leaderless generalized consensus, and example of dependency-set algorithm implementation.}
 \label{fig:deps-set-algo}

### 2016-02-12 podc16/paper2.tex
added def of vertice-induced subgraph

--- 2016-02-12 podc16/paper2.tex
@@
 
-The graph processing algorithm is executed locally by a node (line 10 of \cref{fig:deps-set-algo3}). It must compute the value of $\opn{c-struct}\left( g_p \right)$, as defined in \cref{lindef}.
+The graph processing algorithm is executed locally by a process (line 10 of \cref{fig:deps-set-algo3}). It must compute the value of $\opn{c-struct}\left( g_p \right)$, as defined in \cref{lindef}.
 Its key property is \cref{lemma:graph}. 
@@
 
-Define the intersection of two graphs such that $V\left( g_1 \cap g_2 \right)=V\left( g_1 \right)\cap V\left( g_2 \right)$ and $E\left( g_1 \cap g_2 \right)= E\left( g_1 \right)\cap E\left( g_2 \right)$, and define their union such that $V\left( g_1 \cup g_2 \right)=V\left( g_1 \right)\cup V\left( g_2 \right)$ and $E\left( g_1 \cup g_2 \right)= E\left( g_1 \right)\cup E\left( g_2 \right)$.
-Define two graphs $g_1$ and $g_2$ as being \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$.
+Define graph intersection such that $V\left( g_1 \cap g_2 \right)=V\left( g_1 \right)\cap V\left( g_2 \right)$ and $E\left( g_1 \cap g_2 \right)= E\left( g_1 \right)\cap E\left( g_2 \right)$, and graph union such that $V\left( g_1 \cup g_2 \right)=V\left( g_1 \right)\cup V\left( g_2 \right)$ and $E\left( g_1 \cup g_2 \right)= E\left( g_1 \right)\cup E\left( g_2 \right)$.
+Define two graphs $g_1$ and $g_2$ as being \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$ ($g$ is a vertex-induced subgraph of $g'$ iff $V\left( g \right)\subseteq V\left( g' \right)$ and if both endpoints of $e\in E\left( g' \right)$ are in $V\left( g \right)$, then $e\in E\left( g \right)$).
 

### 2016-02-12 podc16/paper2.tex
small fix

@@
 Define graph intersection such that $V\left( g_1 \cap g_2 \right)=V\left( g_1 \right)\cap V\left( g_2 \right)$ and $E\left( g_1 \cap g_2 \right)= E\left( g_1 \right)\cap E\left( g_2 \right)$, and graph union such that $V\left( g_1 \cup g_2 \right)=V\left( g_1 \right)\cup V\left( g_2 \right)$ and $E\left( g_1 \cup g_2 \right)= E\left( g_1 \right)\cup E\left( g_2 \right)$.
-Define two graphs $g_1$ and $g_2$ as being \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$ ($g$ is a vertex-induced subgraph of $g'$ iff $V\left( g \right)\subseteq V\left( g' \right)$ and if both endpoints of $e\in E\left( g' \right)$ are in $V\left( g \right)$, then $e\in E\left( g \right)$).
+Define two graphs $g_1$ and $g_2$ as being \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$ ($g$ is a vertex-induced subgraph of $g'$ iff $V\left( g \right)\subseteq V\left( g' \right)$, and for every $e\in E\left( g' \right)$, if both endpoints of $e$ are in $V\left( g \right)$, then $e\in E\left( g \right)$).
 

### 2016-02-12 podc16/paper2.tex
small fix

@@
 
-We say that a command $c$ in the domain of $m_p$ is \emph{executable} when no command $c$ which is not in the domain of $m_p$ can be reached from $c$ by recursively following $c$'s dependencies.
+We say that a command $c$ in the domain of $m_p$ is \emph{executable} when no command $c'$ which is not in the domain of $m_p$ can be reached from $c$ by recursively following $c$'s dependencies.
 The local variable $m_p$ denotes a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the executable commands of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ iff $c_2 \in m_p\left[ c_1 \right]$ (i.e., $c_1$ depends on $c_2$).

### 2016-02-12 podc16/paper2.tex
minor changes to pseudocode, and better description of leaderless generalized consensus algorithm

--- 2016-02-12 podc16/paper2.tex
@@
 
-Our leaderless generalized consensus algorithm orchestrates a dependency-set sub-algorithm and map-agreement sub-algorithm as shown in \cref{fig:deps-set-algo3}.
-A process $p$ receiving a command $c$ proceeds in three phases.
-First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $\mathcal{D}$ that the command $c$ may depend on.
-Second, $p$ proposes the mapping  $c\mapsto \mathcal{D}$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto \mathcal{D}'$ for $c$, in which case $\mathcal{D}'$ becomes the dependency set of $c$;
-we say that $c$ is executable when all its transitive dependencies are known.
+Our leaderless generalized consensus algorithm orchestrates a dependency-set algorithm and map-agreement algorithm as shown in \cref{fig:deps-set-algo3}.
+A process $p$ maintains two local variables $m_p$ (a map from commands to sets of commands) and $\sigma_p$ (a c-struct).
+Upon  receiving a command $c$, $p$ proceeds in three phases.
+First, it calls a \textit{dependency-set} algorithm (\Cref{dep-algo}) to determine a set of commands $\mathcal{D}$ that the command $c$ may depend on.
+Second, $p$ proposes the mapping  $c\mapsto \mathcal{D}$ to a \textit{map-agreement} algorithm (\Cref{map-algo}); the map-agreement algorithm may subsequently commit a mapping $c\mapsto \mathcal{D}'$ for $c$, in which case $p$ inserts the mapping $\left[c \mapsto \mathcal{D}'\right]$ in $m_p$, and we say that $\mathcal{D}'$ is the dependency set of $c$.
+We say that $c$ is executable in $m_p$ when all its transitive dependencies are known.
 %no command $c'$ for which no mapping has been committed is reachable from $c$ by recursively following committed mappings.
-Finally, when $c$ is executable, the $p$ may execute $c$ by deciding the c-struct obtained by \emph{locally} running a \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
+Finally, when $c$ is executable in $m_p$, then $p$ may execute $c$ by deciding the c-struct obtained by \emph{locally} running a \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
 Any other process can run the three phases independently of $p$ and eventually execute the command despite a crash of $p$.
-However, we show that a leaderless  generalized consensus algorithm is always safe and is live when its underlying dependency-set and map-agreement algorithms are safe and live.
+We show that the leaderless  generalized consensus algorithm guarantees the safety and liveness properties of generalized consensus when its underlying dependency-set and map-agreement algorithms satisfy their safety and liveness properties.
 
@@
 
-A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{commit}\left( c,\mathcal{D} \right)$ (meant to observe the map-agreement sub-algorithm), where $p\in P$ is the process executing the call.
+A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{commit}\left( c,\mathcal{D} \right)$ (meant to observe the map-agreement algorithm), where $p\in P$ is the process executing the call.
 Moreover, a dependency-set algorithm must ensure that:
@@
 
-An example of implementation of such an algorithm is depicted by the pseudocode of Figures~\ref{fig:deps-set-algo1} and~\ref{fig:deps-set-algo2}.
+An example of implementation of such an algorithm is depicted by the pseudocode of \cref{fig:deps-set-algo1,fig:deps-set-algo2}.
 Due to space constraints, our example focuses on satisfying the safety property only.
@@
 			\END
+			\line(1,0){80}
 			\START[$\opn{GC-propose}_p$]{{\bf upon}}{$(c)$}
@@
 			\END
-			\Statex
+			\line(1,0){80}
 			\START[$\opn{suggest}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
@@
 			\END
-			\Statex
+			\line(1,0){80}
 			\START[$\opn{commit}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
@@
 			\END
-			\Statex
+			\line(1,0){108}
 			\START[\texttt{announce$_p$}]{{\bf upon}}{$(c)$} \label{alg:announce}
@@
 			\END
-			\Statex
+			\line(1,0){125}
 			\START[$\langle \texttt{deps}$ $|$ $c,\mathcal{D} \rangle$]{{\bf upon receive}}{ {\bf from} $q$} \label{alg:receive-deps}
@@
 			\END
-			\Statex
+			\line(1,0){125}
 			\START[]{{\bf upon}}{$\exists c$ $:$ $ | heard_p[c] | \geq \mathcal{Q}$ $\land$} \label{alg:receive-deps}

### 2016-02-12 podc16/paper2.tex
no GC

@@
 
-Any implementation of dependency-set algorithm can rely on the garbage collection of the map-agreement module to get rid of redundant information associated with specific commands in its data structures, with the purpose of preventing the dependency sets to grow indefinitely.
+%Any implementation of dependency-set algorithm can rely on the garbage collection of the map-agreement module to get rid of redundant information associated with specific commands in its data structures, with the purpose of preventing the dependency sets to grow indefinitely.
 
-\todo{``Suggest'' can also be used to trigger recovery. Describe better this part.}
+%\todo{``Suggest'' can also be used to trigger recovery. Describe better this part.}
 %\todo{Write a suggestion on about GC of dependencies. Remove any command c such that a committed command depends on c. Also us the GC of the map-agreement component.}

### 2016-02-12 podc16/paper2.tex
work on the intro

--- 2016-02-12 podc16/paper2.tex
@@
 
-We show that the core idea underlying EPaxos can be captured in a generic algorithm that uses two abstractions: (1), a per-command dependency-set computation algorithm, a concept that we define in the paper, and (2), an algorithm ensuring, for each command proposed to it, agreement on a dependency set.
+We show that the core idea underlying EPaxos can be captured in a generic algorithm that uses two abstractions: a dependency-set algorithm, which suggests dependencies for commands, and a map-agreement algorithm, which ensures that, for each submitted command, processes agree on a dependency set.
 Both abstractions lend themselves well to implementations that avoid the use of a unique leader process.
+Our generic algorithm gives rise to a family of algorithms whose members are obtained by using a concrete agreement algorithm and a concrete dependency-set algorithm. 
+%Note that any generalized consensus algorithm can be used to implement (2), but only a weaker problem than generalized consensus needs to be solved because no ordering among proposal is required.
 
-Our abstract algorithm gives rise to a family of algorithms whose members are obtained by using a concrete map-agreement algorithm and a concrete dependency-set computation algorithm. 
-Note that any generalized consensus algorithm can be used to implement (2), but only a weaker problem than generalized consensus needs to be solved because no ordering among proposal is required.
-
-Our work has two main advantages: it enables modular correctness proofs of algorithms like EPaxos and it allows to tune the performance of implementations by changing the underlying map-agreement and dependency-set algorithms.
-We expect that these two tunable knobs will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by this new family of algorithms.
+On top of enabling modular correctness proofs of algorithms like EPaxos, we expect that the modular structure of our generic leaderless algorithm will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by EPaxos-like algorithms.
 
@@
 
-Our leaderless generalized consensus algorithm orchestrates a dependency-set algorithm and map-agreement algorithm as shown in \cref{fig:deps-set-algo3}.
-A process $p$ maintains two local variables $m_p$ (a map from commands to sets of commands) and $\sigma_p$ (a c-struct).
-Upon  receiving a command $c$, $p$ proceeds in three phases.
-First, it calls a \textit{dependency-set} algorithm (\Cref{dep-algo}) to determine a set of commands $\mathcal{D}$ that the command $c$ may depend on.
-Second, $p$ proposes the mapping  $c\mapsto \mathcal{D}$ to a \textit{map-agreement} algorithm (\Cref{map-algo}); the map-agreement algorithm may subsequently commit a mapping $c\mapsto \mathcal{D}'$ for $c$, in which case $p$ inserts the mapping $\left[c \mapsto \mathcal{D}'\right]$ in $m_p$, and we say that $\mathcal{D}'$ is the dependency set of $c$.
-We say that $c$ is executable in $m_p$ when all its transitive dependencies are known.
+Our leaderless generalized consensus algorithm orchestrates a dependency-set sub-algorithm and map-agreement sub-algorithm as shown in \cref{fig:deps-set-algo3}.
+A process $p$ receiving a command $c$ proceeds in three phases.
+First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $\mathcal{D}$ that the command $c$ may depend on.
+Second, $p$ proposes the mapping  $c\mapsto \mathcal{D}$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto \mathcal{D}'$ for $c$, in which case $\mathcal{D}'$ becomes the dependency set of $c$;
+we say that $c$ is executable when all its transitive dependencies are known.
 %no command $c'$ for which no mapping has been committed is reachable from $c$ by recursively following committed mappings.
-Finally, when $c$ is executable in $m_p$, then $p$ may execute $c$ by deciding the c-struct obtained by \emph{locally} running a \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
+Finally, when $c$ is executable, the $p$ may execute $c$ by deciding the c-struct obtained by \emph{locally} running a \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
 Any other process can run the three phases independently of $p$ and eventually execute the command despite a crash of $p$.
-We show that the leaderless  generalized consensus algorithm guarantees the safety and liveness properties of generalized consensus when its underlying dependency-set and map-agreement algorithms satisfy their safety and liveness properties.
+However, we show that a leaderless  generalized consensus algorithm is always safe and is live when its underlying dependency-set and map-agreement algorithms are safe and live.
 
@@
 
-A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{commit}\left( c,\mathcal{D} \right)$ (meant to observe the map-agreement algorithm), where $p\in P$ is the process executing the call.
+A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{commit}\left( c,\mathcal{D} \right)$ (meant to observe the map-agreement sub-algorithm), where $p\in P$ is the process executing the call.
 Moreover, a dependency-set algorithm must ensure that:
@@
 
-An example of implementation of such an algorithm is depicted by the pseudocode of \cref{fig:deps-set-algo1,fig:deps-set-algo2}.
+An example of implementation of such an algorithm is depicted by the pseudocode of Figures~\ref{fig:deps-set-algo1} and~\ref{fig:deps-set-algo2}.
 Due to space constraints, our example focuses on satisfying the safety property only.
@@
 			\END
-			\line(1,0){80}
 			\START[$\opn{GC-propose}_p$]{{\bf upon}}{$(c)$}
@@
 			\END
-			\line(1,0){80}
+			\Statex
 			\START[$\opn{suggest}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
@@
 			\END
-			\line(1,0){80}
+			\Statex
 			\START[$\opn{commit}_p$]{{\bf upon}}{$(c,\mathcal{D})$}
@@
 			\END
-			\line(1,0){108}
+			\Statex
 			\START[\texttt{announce$_p$}]{{\bf upon}}{$(c)$} \label{alg:announce}
@@
 			\END
-			\line(1,0){125}
+			\Statex
 			\START[$\langle \texttt{deps}$ $|$ $c,\mathcal{D} \rangle$]{{\bf upon receive}}{ {\bf from} $q$} \label{alg:receive-deps}
@@
 			\END
-			\line(1,0){125}
+			\Statex
 			\START[]{{\bf upon}}{$\exists c$ $:$ $ | heard_p[c] | \geq \mathcal{Q}$ $\land$} \label{alg:receive-deps}

### 2016-02-12 podc16/paper2.tex
authors
work on the intro

@@
 
-On top of enabling modular correctness proofs of algorithms like EPaxos, we expect that the modular structure of our generic leaderless algorithm will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by EPaxos-like algorithms.
-
-We prove that all the algorithms in the family are safe and live, and we have formalized our ideas in TLA+ and have thoroughly model-checked our safety claims with the TLC model-checker.
+On top of enabling modular correctness proofs of algorithms like EPaxos, we expect that the modular structure of our generic leaderless algorithm will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by different implementations of our two abstractions, as well as optimizations to their orchestration.
+We have formalized our generic leaderless algorithm in TLA+ and have thoroughly model-checked our safety claims with the TLC model-checker.
 

### 2016-02-12 podc16/paper2.tex
no appendix
property references

@@
 
+\begin{comment}
 \author{
-Author 1\\
+{\small Author 1}\\
+{\small ECE, Virginia Tech}\\
+{\small email}
+\and
+{\small Author 2}\\
+{\small ECE, Virginia Tech}\\
+{\small email}
+\and
+{\small Author 3}\\
+{\small ECE, Virginia Tech}\\
+{\small email}
+}
+\end{comment}
+
+%\begin{comment}
+
+\author{
+Giuliano Losa\\
 ECE, Virginia Tech\\
-email
+giuli4no@vt.edu
 \and
-Author 2\\
+Sebastiano Peluso\\
 ECE, Virginia Tech\\
-email
+peluso@vt.edu
 \and
-Author 3\\
+Binoy Ravindran\\
 ECE, Virginia Tech\\
-email
+binoy@vt.edu
 }
+%\end{comment}
 
@@
 
+\vspace{-0.5cm}
 \section{Introduction}
@@
 \begin{proof*}
-Follows from the safety property (b) of the map-agreement algorithm and definition of the graph $g_p$.
-From (b) it follows that if $c$ is in the domain of both $m_p$ and $m_q$ then $m_p\left[ c \right] = m_q\left[ c \right]$.
+TODO
+%Follows from the safety property (b) of the map-agreement algorithm and definition of the graph $g_p$.
+%From (b) it follows that if $c$ is in the domain of both $m_p$ and $m_q$ then $m_p\left[ c \right] = m_q\left[ c \right]$.
 \end{proof*}
 \begin{mythm}
-    The consistency property always holds.
+    The consistency property of generalized consensus always holds.
     \label{lemma:consistency}
@@
 \begin{proof*}
-    By \cref{lemma:compat} and the safety property (b) of the dependency-set algorithm, $g_p^1$ and $g_q^2$ satisfy the premises of \cref{lemma:graph}.
-Therefore we get that $g_p^1$ and $g_q^2$ are compatible.
+    By \cref{lemma:compat} and \ref{suggestion-agreement}, $g_p^1$ and $g_q^2$ satisfy the premises of \cref{lemma:graph}. Therefore we get that $g_p^1$ and $g_q^2$ are compatible.
 \end{proof*}
@@
 \begin{proof*}
-    The liveness properties (a) and (b) of the dependency-set algorithm ensures that some dependency sets are repeatedly proposed for $c$ to the map-agreement algorithm, as long as a commit is not observed.
+    The properties \ref{eventual-suggestion} and \ref{re-suggest} ensure that dependency sets are repeatedly proposed for $c$ to the map-agreement algorithm as long as a commit is not observed.
 Moreover, the map-agreement algorithm ensures that when dependency sets are repeatedly proposed for $c$, then a dependency set will eventually be committed for $c$.

### 2016-02-12 podc16/paper2.tex
Working on the implementation section.

@@
 We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
-C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $\mathcal{D}$ such that $\sigma_2 = \sigma_1\bullet \mathcal{D}$ and $\bot\sqsubseteq \sigma$ for any
+C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $s$ such that $\sigma_2 = \sigma_1\bullet s$ and $\bot\sqsubseteq \sigma$ for any
 $\sigma$. Two c-structs are compatible when they have a
@@
 First, it calls a \textit{dependency-set} algorithm (\Cref{dep-algo}) to determine a set of commands $\mathcal{D}$ that the command $c$ may depend on.
-Second, $p$ proposes the mapping  $c\mapsto \mathcal{D}$ to a \textit{map-agreement} algorithm (\Cref{map-algo}); the map-agreement algorithm may subsequently commit a mapping $c\mapsto \mathcal{D}'$ for $c$, in which case $p$ inserts the mapping $\left[c \mapsto \mathcal{D}'\right]$ in $m_p$, and we say that $\mathcal{D}'$ is the dependency set of $c$.
+Second, $p$ proposes the mapping  $c\mapsto \mathcal{D}$ to a \textit{map-agreement} algorithm (\Cref{map-algo}); when the map-agreement algorithm
+commits a mapping $c\mapsto \mathcal{D}'$ for $c$, $p$ inserts the mapping $\left[c \mapsto \mathcal{D}'\right]$ in $m_p$, and we say that $\mathcal{D}'$ is the dependency set of $c$.
 We say that $c$ is executable in $m_p$ when all its transitive dependencies are known.
@@
 Finally, when $c$ is executable in $m_p$, then $p$ may execute $c$ by deciding the c-struct obtained by \emph{locally} running a \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
-Any other process can run the three phases independently of $p$ and eventually execute the command despite a crash of $p$.
-We show that the leaderless  generalized consensus algorithm guarantees the safety and liveness properties of generalized consensus when its underlying dependency-set and map-agreement algorithms satisfy their safety and liveness properties.
+Any other process can run the three phases independently of $p$ and eventually execute the command, possibly despite a crash of $p$.
+%We show that the leaderless  generalized consensus algorithm guarantees the safety and liveness properties of generalized consensus when its underlying dependency-set and map-agreement algorithms satisfy their safety and liveness properties.
 
@@
 
-A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{commit}\left( c,\mathcal{D} \right)$ (meant to observe the map-agreement algorithm), where $p\in P$ is the process executing the call.
+A dependency-set algorithm exposes at each process $p$ the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{commit}\left( c,\mathcal{D} \right)$,  to observe which command is committed by the other abstraction.
 Moreover, a dependency-set algorithm must ensure that:
@@
 \end{itemize}
-
-An example of implementation of such an algorithm is depicted by the pseudocode of \cref{fig:deps-set-algo1,fig:deps-set-algo2}.
-Due to space constraints, our example focuses on satisfying the safety property only.
-The main idea is the following: a process $p$ announcing a command $c$ broadcasts $c$ to all the processes, and it suggests the dependency-set $deps_p[c]$ for $c$ as the union of the sets $\mathcal{D}$ that it receives as replies from a quorum of $\mathcal{Q}$ processes; on the other hand, a process that receives an announcement for $c$, computes its reply $\mathcal{D}$ for $c$ as the set of commands $d$ that it received so far, and such that $c$ and $d$ do not commute.
-The choice of the value of $\mathcal{Q}$ must be such that the intersection of any two quorums is not empty.
-About correctness, we can easily prove that such an implementation satisfies the safety property by noticing that:  any process $p$ inserts a command $d$ in a dependency-set $\mathcal{D}$ that $p$ sent as a reply for a command $c$, only if $p$ previously received $d$.
-Therefore, every command suggested in a dependency-set must have been announced before.
-Further, since a process can suggest a set of dependencies $deps_p[c]$ for a command $c$ after it has received a reply for $c$ from at least a quorum of processes, and thanks to the requirement that we enforced on the intersection of quorums, we can conclude that for any two suggestions $\opn{suggest}_p \left( c_1,deps_p[c_1]\right)$ and $\opn{suggest}_q \left( c_2,deps_q[c_2]\right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_q[c_2]$ or $c_2 \in deps_p[c_1]$.
 
@@
 
-An example of implementation of the map-agreement algorithm can be any state machine replication algorithm, like MultiPaxos~\cite{lamport2001paxos}, where a pair of command $c$ and dependency-set $\mathcal{D}$ can be proposed to a instance of MultiPaxos that is identified by the command $c$ itself.
-%used to implementing map-agreement by uniquely associating commands to MultiPaxos instance numbers.
-
-\subsection{Processing the Dependency Graph}
+\subsection{Local Dependency Graph Processing}
 \label{abstract-algo}
@@
 
-\subsection{Obtaining EPaxos}
+\subsection{Implementing the Abstractions}
+Both the dependency-set and map-agreement abstractions are well suited to leaderless implementations.
+
+The dependency-set abstraction can be implemented as shown in \cref{fig:deps-set-algo1,fig:deps-set-algo2}.
+Due to space constraints, our example focuses on satisfying the safety property only.
+A process $p$ announcing a command $c$ asks all the processes for the set of commands they have seen so far and that do not commute with $c$.
+Then, it suggests the union of the dependency sets received from a \emph{quorum} of processes.
+Quorums are sets of processes such that the intersection of any two
+quorums is not empty. Since $p$ receives sets of commands seen by other processes, the implementation trivially ensures \ref{dep-valid}. 
+Further, consider two suggestions $\opn{suggest}\left(c_1,\matthcal{D}_1\right)$ and $\opn{suggest}\left(c_2,\matthcal{D}_2\right)$ where $c_1$ and $c_2$ do not commute.
+There are two quorums $Q_1$ and $Q_2$ such that $\matthcal{D}_1$ was computed from $Q_1$ and $\matthcal{D}_2$ was computed from $Q_2$.
+By the definition of quorums, there is a process $q$ belonging to both $Q_1$ and $Q_2$.
+This process $q$ received either $c_1$ before $c_2$, in which case $c_1\in
+\mathcal{D}_2$, or $c_2$ before $c_1$, in which case $c_2\in\mathcal{D}_1$. Therefore, property \ref{suggestion-agreement} holds.
+
+The map-agreement algorithm can be implemented by any state machine replication algorithm, like MultiPaxos~\cite{lamport2001paxos}, where a command $c$ is uniquely associated with a position in the sequence of consensus instances.
+%used to implementing map-agreement by uniquely associating commands to MultiPaxos instance numbers.
+However, MultiPaxos is too strong because we do not care about the ordering among instances.
+Instead, one can uniquely associate one incarnation of MultiPaxos with each process, where each process is initially the leader of its MultiPaxos incarnation.  
+We assume that a command is associated with a unique natural number $\opn{id}\left( c \right)$ and a process id $\opn{pid}\left( c \right)$, corresponding to the first process that received $c$.
+A process receiving a proposal $\opn{propose}_p\left( c, \mathcal{D}) $  proposes $ c\mapsto\mathcal{D} $ in instance $\opn{id}\left( c \right)$ of the MultiPaxos incarnation of process $\opn{pid}\left( c \right)$.
+Upon a MultiPaxos decision $ c\mapsto\mathcal{D'} $, $p$ calls $\opn{commit}_p\left( c,\mathcal{D} \right)$.
+Note that in a failure-free case, a command is committed in two message delays by its originator.
+
+EPaxos improves upon our MultiPaxos scheme in two ways.
+First, EPaxos uses Fast Paxos\cite{Lamport05GeneralizeConsensus} incarnations instead of MultiPaxos incarnations.
+Second, EPaxos combines the dependency-set algorithm outlined above, but using fast quorums, with the first phase of Fast Paxos: the state of the dependency-set algorithm and of the Fast Paxos incarnations are merged so as to allow the reception of identical dependency sets from a fast quorum of processes
+(after an $\opn{announce}_p\left( c \right)$) to act as a fast-path decision of Fast Paxos.
+
+%A process receiving a proposal $ c\mapsto\mathcal{D} $ attaches a unique natural number $\opn{id}\left( c \right)$ as command identifier and its own id $\opn{pid}\left( c \right)$.
+proposes $ c\mapsto\mathcal{D} $ in instance $\opn{id}\left( c \right)$ of its MultiPaxos incarnation.
+Upon a MultiPaxos decision $ c\mapsto\mathcal{D'} $, $p$ commits $
+%c\mapsto\mathcal{D'} $.
 
 \printbibliography%
-
+\begin{comment}
 \appendix
@@
 
+\end{comment}
 \end{document}

### 2016-02-12 podc16/paper2.tex
compilation fix

@@
 quorums is not empty. Since $p$ receives sets of commands seen by other processes, the implementation trivially ensures \ref{dep-valid}. 
-Further, consider two suggestions $\opn{suggest}\left(c_1,\matthcal{D}_1\right)$ and $\opn{suggest}\left(c_2,\matthcal{D}_2\right)$ where $c_1$ and $c_2$ do not commute.
-There are two quorums $Q_1$ and $Q_2$ such that $\matthcal{D}_1$ was computed from $Q_1$ and $\matthcal{D}_2$ was computed from $Q_2$.
+Further, consider two suggestions $\opn{suggest}\left(c_1,\mathcal{D}_1\right)$ and $\opn{suggest}\left(c_2,\mathcal{D}_2\right)$ where $c_1$ and $c_2$ do not commute.
+There are two quorums $Q_1$ and $Q_2$ such that $\mathcal{D}_1$ was computed from $Q_1$ and $\mathcal{D}_2$ was computed from $Q_2$.
 By the definition of quorums, there is a process $q$ belonging to both $Q_1$ and $Q_2$.
@@
 We assume that a command is associated with a unique natural number $\opn{id}\left( c \right)$ and a process id $\opn{pid}\left( c \right)$, corresponding to the first process that received $c$.
-A process receiving a proposal $\opn{propose}_p\left( c, \mathcal{D}) $  proposes $ c\mapsto\mathcal{D} $ in instance $\opn{id}\left( c \right)$ of the MultiPaxos incarnation of process $\opn{pid}\left( c \right)$.
+A process receiving a proposal $\opn{propose}_p\left( c, \mathcal{D}\right) $  proposes $ c\mapsto\mathcal{D} $ in instance $\opn{id}\left( c \right)$ of the MultiPaxos incarnation of process $\opn{pid}\left( c \right)$.
 Upon a MultiPaxos decision $ c\mapsto\mathcal{D'} $, $p$ calls $\opn{commit}_p\left( c,\mathcal{D} \right)$.

### 2016-02-12 podc16/paper2.tex
minor fixes

@@
 We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network.
-Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}.
-Each process receives proposals for commands of the form $\opn{Propose}(c)$, and must produce responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct.
-The sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
+Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}, in which each process receives proposals for commands of the form $\opn{Propose}(c)$, and must produce responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct, and the sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
 We use the same notation and refer the reader to~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of c-structs and generalized consensus.
@@
 
-\subsection{Local Dependency Graph Processing}
+\subsection{Local Dependency-Graph Processing}
 \label{abstract-algo}
@@
 Define graph intersection such that $V\left( g_1 \cap g_2 \right)=V\left( g_1 \right)\cap V\left( g_2 \right)$ and $E\left( g_1 \cap g_2 \right)= E\left( g_1 \right)\cap E\left( g_2 \right)$, and graph union such that $V\left( g_1 \cup g_2 \right)=V\left( g_1 \right)\cup V\left( g_2 \right)$ and $E\left( g_1 \cup g_2 \right)= E\left( g_1 \right)\cup E\left( g_2 \right)$.
-Define two graphs $g_1$ and $g_2$ as being \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$ ($g$ is a vertex-induced subgraph of $g'$ iff $V\left( g \right)\subseteq V\left( g' \right)$, and for every $e\in E\left( g' \right)$, if both endpoints of $e$ are in $V\left( g \right)$, then $e\in E\left( g \right)$).
+We say that $g$ is a vertex-induced subgraph of $g'$ iff $V\left( g \right)\subseteq V\left( g' \right)$, and for every $e\in E\left( g' \right)$, if both endpoints of $e$ are in $V\left( g \right)$, then $e\in E\left( g \right)$).
+Define two graphs $g_1$ and $g_2$ as \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$.
 

### 2016-02-12 podc16/paper2.tex
compact proofs and theorems

@@
 
-\subsection{Computing Dependency-Sets}
+\subsection{Computing Potential Dependency Sets}
 \label{dep-algo}
@@
 }
-\hspace*{0.1in}\subfigure[Example of dependency-set algorithm implementation - part 1.]{
+\hspace*{0.1in}\subfigure[Dependency-set algorithm implementation - part 1.]{
 \noindent\fbox{%
@@
 
-\subsection{Satisfying Consistency and Liveness of Generalized Consensus}
+\subsection{Correctness}
 Consider an execution of the algorithm.
-\begin{mylemma}\label{lemma:compat}
-If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible.
-\end{mylemma}
-\begin{proof*}
+%\begin{mylemma}\label{lemma:compat}
+%If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible.
+%\end{mylemma}
+%\begin{proof*}
 TODO
@@
 \end{proof*}
-\begin{mythm}
+
+\textbf{Lemma 2.} If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible.  
+\emph{Proof.} TODO
+
+\textbf{Theorem 1.}
 The consistency property of generalized consensus always holds.
-    \label{lemma:consistency}
-\end{mythm}
-\begin{proof*}
-    By \cref{lemma:compat} and \ref{suggestion-agreement}, $g_p^1$ and $g_q^2$ satisfy the premises of \cref{lemma:graph}. Therefore we get that $g_p^1$ and $g_q^2$ are compatible.
-\end{proof*}
-\begin{mythm}
+\emph{Proof.} By lemma 2 and property \ref{suggestion-agreement}, $g_p^1$ and $g_q^2$ satisfy the premises of \cref{lemma:graph}. Therefore we get that $g_p^1$ and $g_q^2$ are compatible.
+%\begin{mythm}
+%    The consistency property of generalized consensus always holds.
+%    \label{lemma:consistency}
+%\end{mythm}
+%\begin{proof*}
+%    By \cref{lemma:compat} and \ref{suggestion-agreement}, $g_p^1$ and $g_q^2$ satisfy the premises of \cref{lemma:graph}. Therefore we get that $g_p^1$ and $g_q^2$ are compatible.
+%\end{proof*}
+
+\textbf{Theorem 2.}
 If a command $c$ which is GC-proposed is then repeatedly GC-proposed, then a c-struct containing $c$ is eventually decided.
-\end{mythm}
-\begin{proof*}
+\emph{Proof.}
 The properties \ref{eventual-suggestion} and \ref{re-suggest} ensure that dependency sets are repeatedly proposed for $c$ to the map-agreement algorithm as long as a commit is not observed.
@@
 Similarly, any dependency of $c$ is eventually committed and $c$ becomes executable, at which points it is included in the next decided c-struct.
-\end{proof*}
+
+%\begin{mythm}
+%    If a command $c$ which is GC-proposed is then repeatedly GC-proposed, then a c-struct containing $c$ is eventually decided.
+%\end{mythm}
+%\begin{proof*}
+%    The properties \ref{eventual-suggestion} and \ref{re-suggest} ensure that dependency sets are repeatedly proposed for $c$ to the map-agreement algorithm as long as a commit is not observed.
+Moreover, the map-agreement algorithm ensures that when dependency sets are repeatedly proposed for $c$, then a dependency set will eventually be committed for $c$.
+Similarly, any dependency of $c$ is eventually committed and $c$ becomes executable, at which points it is included in the next decided c-struct.
+%\end{proof*}
 

### 2016-02-12 podc16/paper2.tex
small fix
fitting everything in 3 pages

@@
 
-Many agreement algorithms inspired by Paxos and MultiPaxos~\cite{lamport2001paxos} rely on the existence of a unique leader process which enforces an ordering on commands during fault-free periods.
+Most agreement algorithms inspired by Paxos and MultiPaxos~\cite{lamport2001paxos} rely on the existence of a unique leader process which enforces an ordering on commands during fault-free periods.
 In practice, the unique leader can easily become a performance bottleneck for the system.
@@
 \begin{itemize}[noitemsep,nolistsep]
-    \item[-] Safety: (\setword{S1}{dep-valid}), for any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before; (\setword{S2}{suggestion-agreement}), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
+    \item[-] Safety: (\setword{S1}{dep-valid}), for any call $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before; (\setword{S2}{suggestion-agreement}), for any two calls $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then $c_1\in \mathcal{D}_2\vee c_2 \in \mathcal{D}_1$.
 %    \item If $c$ was announced by $p$, then eventually a dependency set is suggested for $c$ to $p$.
@@
 \begin{mylemma}\label{lemma:graph}
-Assume that $g_1$ and $g_2$ are two dependency graphs, 
-that for every pair of non-commutative commands $c_1,c_2 \in V\left(g_2\right)$, either $\left( c_1,c_2 \right)\in E\left( g_2 \right)$ or $\left( c_2,c_1 \right)\in E\left( g_2 \right)$,
-that $g_1$ and $g_2$ are compatible,
-that $l_1$ is a linearization of $g_1$, and that $l_2$ is a linearization of $g_2$. 
-Then we have (a), $\bot\bullet l_1$ and $\bot\bullet l_2$ are compatible c-structs, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.
+Assume that: $l_1$ and $l_2$ are linearizations, respectively, of two compatible dependency graphs $g_1$ and $g_2$; 
+for every pair of non-commutative commands $c_1,c_2 \in V\left(g_2\right)$, either $\left( c_1,c_2 \right)\in E\left( g_2 \right)$ or $\left( c_2,c_1 \right)\in E\left( g_2 \right)$.
+Then we have: (a), $\bot\bullet l_1$ and $\bot\bullet l_2$ are compatible c-structs; (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.
 \end{mylemma}
@@
 We define $\opn{c-struct}\left( m_p \right) = \bot\bullet l$ where $l$ is a linearization of $g_p$.
-By lemma 1.b, $\opn{c-struct}\left( m_p \right)$ is well-defined.
 \end{mydef}
+By lemma 1(b), $\opn{c-struct}\left( m_p \right)$ is well-defined.
 
 \subsection{Correctness}
-Consider an execution of the algorithm.
+%Consider an execution of the algorithm.
 %\begin{mylemma}\label{lemma:compat}
@@
 %\begin{proof*}
-TODO
 %Follows from the safety property (b) of the map-agreement algorithm and definition of the graph $g_p$.
 %From (b) it follows that if $c$ is in the domain of both $m_p$ and $m_q$ then $m_p\left[ c \right] = m_q\left[ c \right]$.
-\end{proof*}
+%\end{proof*}
 
-\textbf{Lemma 2.} If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible.  
-\emph{Proof.} TODO
+\noindent\textbf{Lemma 2.} \emph{If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible. }
 
-\textbf{Theorem 1.}
-The consistency property of generalized consensus always holds.
-\emph{Proof.} By lemma 2 and property \ref{suggestion-agreement}, $g_p^1$ and $g_q^2$ satisfy the premises of \cref{lemma:graph}. Therefore we get that $g_p^1$ and $g_q^2$ are compatible.
+Lemma 2 follows from property \ref{commit-agreement} of the map-agreement algorithm and from the definition of $g_p$.
+
+\noindent\textbf{Theorem 1.}
+\emph{The consistency property of generalized consensus always holds.}
+
+Theorem 1 follows from lemma 2 and property \ref{suggestion-agreement}, $g_p^1$ and $g_q^2$ satisfy the premises of \cref{lemma:graph}. Therefore we get that $g_p^1$ and $g_q^2$ are compatible.
 %\begin{mythm}
@@
 
-\textbf{Theorem 2.}
-If a command $c$ which is GC-proposed is then repeatedly GC-proposed, then a c-struct containing $c$ is eventually decided.
-\emph{Proof.}
+\noindent\textbf{Theorem 2.}
+\emph{If a command $c$ which is GC-proposed is then repeatedly GC-proposed, then a c-struct containing $c$ is eventually decided.}
+
 The properties \ref{eventual-suggestion} and \ref{re-suggest} ensure that dependency sets are repeatedly proposed for $c$ to the map-agreement algorithm as long as a commit is not observed.
@@
 
-The map-agreement algorithm can be implemented by any state machine replication algorithm, like MultiPaxos~\cite{lamport2001paxos}, where a command $c$ is uniquely associated with a position in the sequence of consensus instances.
+Any state machine replication algorithm, like MultiPaxos~\cite{lamport2001paxos}, can implement the map-agreement algorithm by uniquely associating commands with positions in the sequence of consensus instances.
 %used to implementing map-agreement by uniquely associating commands to MultiPaxos instance numbers.
-However, MultiPaxos is too strong because we do not care about the ordering among instances.
-Instead, one can uniquely associate one incarnation of MultiPaxos with each process, where each process is initially the leader of its MultiPaxos incarnation.  
-We assume that a command is associated with a unique natural number $\opn{id}\left( c \right)$ and a process id $\opn{pid}\left( c \right)$, corresponding to the first process that received $c$.
+However, MultiPaxos provides unnecessary guarantees on the ordering among instances.
+Instead, one can uniquely associate one incarnation of MultiPaxos with each process, which is initially the leader of its MultiPaxos incarnation.  
+We assume that a command is associated with a unique natural number $\opn{id}\left( c \right)$ and with the process who first received $c$, noted $\opn{pid}\left( c \right)$.
 A process receiving a proposal $\opn{propose}_p\left( c, \mathcal{D}\right) $  proposes $ c\mapsto\mathcal{D} $ in instance $\opn{id}\left( c \right)$ of the MultiPaxos incarnation of process $\opn{pid}\left( c \right)$.
-Upon a MultiPaxos decision $ c\mapsto\mathcal{D'} $, $p$ calls $\opn{commit}_p\left( c,\mathcal{D} \right)$.
-Note that in a failure-free case, a command is committed in two message delays by its originator.
+Upon a MultiPaxos decision $ c\mapsto\mathcal{D'} $, $p$ calls $\opn{commit}_p\left( c,\mathcal{D'} \right)$.
+Note that in a failure-free case, a command is committed in one round-trip by its originator.
 
-EPaxos improves upon our MultiPaxos scheme in two ways.
+EPaxos improves upon this MultiPaxos scheme in two ways.
 First, EPaxos uses Fast Paxos\cite{Lamport05GeneralizeConsensus} incarnations instead of MultiPaxos incarnations.
 Second, EPaxos combines the dependency-set algorithm outlined above, but using fast quorums, with the first phase of Fast Paxos: the state of the dependency-set algorithm and of the Fast Paxos incarnations are merged so as to allow the reception of identical dependency sets from a fast quorum of processes
-(after an $\opn{announce}_p\left( c \right)$) to act as a fast-path decision of Fast Paxos.
+(after an $\opn{announce}_p\left( c \right)$) to act as a single round-trip decision of Fast Paxos.
 

### 2016-02-12 podc16/paper2.tex
minor changes

@@
 
-We show that the core idea underlying EPaxos can be captured in a generic algorithm that uses two abstractions: a dependency-set algorithm, which suggests dependencies for commands, and a map-agreement algorithm, which ensures that, for each submitted command, processes agree on a dependency set.
+We show that the core idea underlying EPaxos can be captured in a generic leaderless generalized consensus algorithm that uses two abstractions: a \textit{dependency-set algorithm}, which suggests dependencies for commands, and a \textit{map-agreement algorithm}, which ensures that, for each submitted command, processes agree on a dependency set.
 Both abstractions lend themselves well to implementations that avoid the use of a unique leader process.
-Our generic algorithm gives rise to a family of algorithms whose members are obtained by using a concrete agreement algorithm and a concrete dependency-set algorithm. 
+Our generic algorithm gives rise to a family of algorithms whose members are obtained by using a concrete map-agreement algorithm and a concrete dependency-set algorithm. 
 %Note that any generalized consensus algorithm can be used to implement (2), but only a weaker problem than generalized consensus needs to be solved because no ordering among proposal is required.
@@
 On top of enabling modular correctness proofs of algorithms like EPaxos, we expect that the modular structure of our generic leaderless algorithm will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by different implementations of our two abstractions, as well as optimizations to their orchestration.
-We have formalized our generic leaderless algorithm in TLA+ and have thoroughly model-checked our safety claims with the TLC model-checker.
+We have formalized our generic leaderless algorithm in TLA+ and have thoroughly model-checked our safety claims with the TLC model-checker\footnote{•}.
 

### 2016-02-13 podc16/paper2.tex
link to tla

@@
 On top of enabling modular correctness proofs of algorithms like EPaxos, we expect that the modular structure of our generic leaderless algorithm will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by different implementations of our two abstractions, as well as optimizations to their orchestration.
-We have formalized our generic leaderless algorithm in TLA+ and have thoroughly model-checked our safety claims with the TLC model-checker\footnote{•}.
+We have formalized our generic leaderless algorithm in TLA+ and have thoroughly model-checked our safety claims with the TLC model-checker\footnote{\url{http://losa.fr/leaderless}}.
 

### 2016-02-13 podc16/paper2.tex
minor

--- 2016-02-12 podc16/paper2.tex
@@
 Both abstractions lend themselves well to implementations that avoid the use of a unique leader process.
-Our generic algorithm gives rise to a family of algorithms whose members are obtained by using a concrete agreement algorithm and a concrete dependency-set algorithm. 
+Our generic algorithm gives rise to a family of algorithms whose members are obtained by using a concrete map-agreement algorithm and a concrete dependency-set algorithm. 
 %Note that any generalized consensus algorithm can be used to implement (2), but only a weaker problem than generalized consensus needs to be solved because no ordering among proposal is required.
@@
 
-\noindent\textbf{Lemma 2.} \emph{If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible. }
+\noindent\textbf{Lemma 2.} \emph{If $g_p^1$ is the graph $g_p$ at some point in an execution, $g_q^2$ is the graph $g_q$ at another point, then $g_p^1$ and $g_q^2$ are compatible. }
 

### 2016-02-13 podc16/paper2.tex
minor

--- 2016-02-13 podc16/paper2.tex
@@
 
-\noindent\textbf{Lemma 2.} \emph{If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible. }
+\noindent\textbf{Lemma 2.} \emph{If $g_p^1$ is the graph $g_p$ at time $t_1$ and $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible. }
 

### 2016-02-13 podc16/paper2.tex
integrated Binoy's comments

@@
 
-On top of enabling modular correctness proofs of algorithms like EPaxos, we expect that the modular structure of our generic leaderless algorithm will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by different implementations of our two abstractions, as well as optimizations to their orchestration.
-We have formalized our generic leaderless algorithm in TLA+ and have thoroughly model-checked our safety claims with the TLC model-checker\footnote{\url{http://losa.fr/leaderless}}.
+On top of enabling modular correctness proofs of algorithms like EPaxos, we expect that the modular structure of our generic leaderless algorithm will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by different implementations of our two abstractions in varied environments such clusters or wide area networks, and on varied metrics such as agreement latency, impact of failures or conflicts, tolerance to slow processes, load
+balancing, size of quorums, etc.
+We have formalized\footnote{\url{http://losa.fr/leaderless}} and model-checked our generic leaderless algorithm with the TLA$^+$ toolbox.
 

### 2016-02-13 podc16/paper2.tex
minor
email

@@
 ECE, Virginia Tech\\
-giuli4no@vt.edu
+glosa@vt.edu
 \and
@@
 Our leaderless generalized consensus algorithm orchestrates a dependency-set algorithm and map-agreement algorithm as shown in \cref{fig:deps-set-algo3}.
-A process $p$ maintains two local variables $m_p$ (a map from commands to sets of commands) and $\sigma_p$ (a c-struct).
+A process $p$ maintains two local variables $m_p$, a map from commands to sets of commands, and $\sigma_p$, a c-struct .
 Upon  receiving a command $c$, $p$ proceeds in three phases.

### 2016-02-13 podc16/paper2.tex
various minor fixes

@@
 
-We show that the core idea underlying EPaxos can be captured in a generic leaderless generalized consensus algorithm that uses two abstractions: a \textit{dependency-set algorithm}, which suggests dependencies for commands, and a \textit{map-agreement algorithm}, which ensures that, for each submitted command, processes agree on a dependency set.
+We show that the core idea underlying EPaxos can be captured in a generic leaderless generalized consensus algorithm that uses two new abstractions: a \textit{dependency-set algorithm}, which suggests dependencies for commands, and a \textit{map-agreement algorithm}, which ensures that, for each submitted command, processes agree on a dependency set.
 Both abstractions lend themselves well to implementations that avoid the use of a unique leader process.
@@
 We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network.
-Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}, in which each process receives proposals for commands of the form $\opn{Propose}(c)$, and must produce responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct, and the sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
-We use the same notation and refer the reader to~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of c-structs and generalized consensus.
+Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}, in which each process receives proposals for commands of the form $\opn{GC-Propose}(c)$, and must produce responses of the form $\opn{GC-Decide}(\sigma)$, where $\sigma$ is a c-struct, and the sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
 Briefly, the c-structs are a set, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct.
 Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands.
-We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
-C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $s$ such that $\sigma_2 = \sigma_1\bullet s$ and $\bot\sqsubseteq \sigma$ for any
+We say that two commands when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
+C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $s$ such that $\sigma_2 = \sigma_1\bullet s$. Moreover, $\bot\sqsubseteq \sigma$ for any
 $\sigma$. Two c-structs are compatible when they have a
@@
 The consistency property of generalized consensus requires that any two decided c-structs be compatible.
+We use the same notation as and refer the reader to~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of c-structs and generalized consensus.
 
@@
 Second, $p$ proposes the mapping  $c\mapsto \mathcal{D}$ to a \textit{map-agreement} algorithm (\Cref{map-algo}); when the map-agreement algorithm
-commits a mapping $c\mapsto \mathcal{D}'$ for $c$, $p$ inserts the mapping $\left[c \mapsto \mathcal{D}'\right]$ in $m_p$, and we say that $\mathcal{D}'$ is the dependency set of $c$.
+commits a mapping $c\mapsto \mathcal{D}'$ for $c$, $p$ inserts the mapping $c \mapsto \mathcal{D}'$ in $m_p$, and we say that $\mathcal{D}'$ is the dependency set of $c$.
 We say that $c$ is executable in $m_p$ when all its transitive dependencies are known.
@@
 
-A dependency-set algorithm exposes at each process $p$ the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{commit}\left( c,\mathcal{D} \right)$,  to observe which command is committed by the other abstraction.
-Moreover, a dependency-set algorithm must ensure that:
+A dependency-set algorithm exposes at each process $p$ the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{commit}\left( c,\mathcal{D} \right)$,  to observe which command is committed by the map-agreement abstraction.
+Moreover, a dependency-set algorithm must ensure the following properties:
 \begin{itemize}[noitemsep,nolistsep]
@@
 %    \item If $c$ was announced by $p$, then eventually a dependency set is suggested for $c$ to $p$.
-    \item[-] Liveness: (\setword{L1}{eventual-suggestion}) If a suggestion $\opn{suggest}_p\left( c \right)$ was made, then eventually a suggestion $\opn{suggest}\left( c,\mathcal{D} \right)_p$ is made, and (\setword{L2}{re-suggest}), if a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
+    \item[-] Liveness: (\setword{L1}{eventual-suggestion}) If an announcement $\opn{announce}_p\left( c \right)$ was made, then eventually a suggestion $\opn{suggest}\left( c,\mathcal{D} \right)_p$ is made, and (\setword{L2}{re-suggest}), if a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
 \end{itemize}
@@
 }	
-\hspace*{0.1in}\subfigure[Example of dependency-set algorithm implementation - part 2.]{
+\hspace*{0.1in}\subfigure[Dependency-set algorithm implementation - part 2.]{
 \noindent\fbox{%
@@
 			\line(1,0){125}
-			\START[]{{\bf upon}}{$\exists c$ $:$ $ | heard_p[c] | \geq \mathcal{Q}$ $\land$} \label{alg:receive-deps}
+            \START[]{{\bf upon}}{$\exists c$ $:$ $ heard_p[c] \in \opn{quorums}$ $\land$} \label{alg:receive-deps}
 			\Statex \hspace{1cm}$\land$ $suggested_p[c]=false$
@@
 \begin{enumerate}[noitemsep,nolistsep]
-    \item[-] Safety: (\setword{S3}{commit-validity}), for any commit $\opn{commit}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time; (\setword{S4}{commit-agreement}), for any two decisions $\opn{commit}_p\left( c, \mathcal{D}_1 \right)$, $\opn{commit}_q\left( c, \mathcal{D}_2 \right)$, $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
+    \item[-] Safety: (\setword{S3}{commit-validity}), for any commit $\opn{commit}_p\left( c, \mathcal{D} \right)$, then $\mathcal{D}$ has been proposed for $c$ at an earlier time; (\setword{S4}{commit-agreement}), for any two decisions $\opn{commit}_p\left( c, \mathcal{D}_1 \right)$ and $\opn{commit}_q\left( c, \mathcal{D}_2 \right)$, $\mathcal{D}_1$ is equal to $\mathcal{D}_2$.
     \item[-] Liveness: (\setword{L3}{eventual-commit}) if a proposal $propose_p\left( c,\mathcal{D} \right)$ is made, then eventually a decision $\opn{commit}_p\left( c, \mathcal{D} \right)$ is made.
@@
 Define graph intersection such that $V\left( g_1 \cap g_2 \right)=V\left( g_1 \right)\cap V\left( g_2 \right)$ and $E\left( g_1 \cap g_2 \right)= E\left( g_1 \right)\cap E\left( g_2 \right)$, and graph union such that $V\left( g_1 \cup g_2 \right)=V\left( g_1 \right)\cup V\left( g_2 \right)$ and $E\left( g_1 \cup g_2 \right)= E\left( g_1 \right)\cup E\left( g_2 \right)$.
-We say that $g$ is a vertex-induced subgraph of $g'$ iff $V\left( g \right)\subseteq V\left( g' \right)$, and for every $e\in E\left( g' \right)$, if both endpoints of $e$ are in $V\left( g \right)$, then $e\in E\left( g \right)$).
+We say that $g$ is a vertex-induced subgraph of $g'$ iff $V\left( g \right)\subseteq V\left( g' \right)$, and for every $e\in E\left( g' \right)$, if both endpoints of $e$ are in $V\left( g \right)$, then $e\in E\left( g \right)$.
 Define two graphs $g_1$ and $g_2$ as \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$.
@@
 We define $\opn{c-struct}\left( m_p \right) = \bot\bullet l$ where $l$ is a linearization of $g_p$.
+\textnormal{By lemma 1(b), $\opn{c-struct}\left( m_p \right)$ is well-defined.}
 \end{mydef}
-By lemma 1(b), $\opn{c-struct}\left( m_p \right)$ is well-defined.
 
@@
 
-\noindent\textbf{Lemma 2.} \emph{If $g_p^1$ is the graph $g_p$ at time $t_1$ and $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible. }
+\noindent\textbf{Lemma 2.} \emph{$g_p^1$, the graph $g_p$ at any time $t_1$, and $g_q^2$, the graph $g_q$ at any time $t_2$, are compatible. }
 
@@
 
-Theorem 1 follows from lemma 2 and property \ref{suggestion-agreement}, $g_p^1$ and $g_q^2$ satisfy the premises of \cref{lemma:graph}. Therefore we get that $g_p^1$ and $g_q^2$ are compatible.
+From lemma 2 and property \ref{suggestion-agreement}, $g_p^1$ and $g_q^2$ satisfy the premises of \cref{lemma:graph}.
+Therefore we get that $\opn{c-struct\left( m_p \right)}$ at time $t_1$ and $\opn{c-struct}\left( m_q \right)$ at time $t_2$ are compatible.
 %\begin{mythm}
@@
 \subsection{Implementing the Abstractions}
-Both the dependency-set and map-agreement abstractions are well suited to leaderless implementations.
+Both the dependency-set and map-agreement abstractions are well suited for leaderless implementations.
 
@@
 Due to space constraints, our example focuses on satisfying the safety property only.
-A process $p$ announcing a command $c$ asks all the processes for the set of commands they have seen so far and that do not commute with $c$.
+A process $p$ announcing a command $c$ asks all the processes for the set of commands $c'$ that they have seen so far and that do not commute with $c$ (noted $c' \asymp c$).
 Then, it suggests the union of the dependency sets received from a \emph{quorum} of processes.
@@
 
-Any state machine replication algorithm, like MultiPaxos~\cite{lamport2001paxos}, can implement the map-agreement algorithm by uniquely associating commands with positions in the sequence of consensus instances.
+A state-machine replication algorithm like MultiPaxos~\cite{lamport2001paxos} can implement the map-agreement algorithm by uniquely associating commands with positions in its sequence of consensus instances.
 %used to implementing map-agreement by uniquely associating commands to MultiPaxos instance numbers.
@@
 Upon a MultiPaxos decision $ c\mapsto\mathcal{D'} $, $p$ calls $\opn{commit}_p\left( c,\mathcal{D'} \right)$.
-Note that in a failure-free case, a command is committed in one round-trip by its originator.
+In a failure-free case, a command is committed in one round-trip by its originator.
 
 EPaxos improves upon this MultiPaxos scheme in two ways.
-First, EPaxos uses Fast Paxos\cite{Lamport05GeneralizeConsensus} incarnations instead of MultiPaxos incarnations.
-Second, EPaxos combines the dependency-set algorithm outlined above, but using fast quorums, with the first phase of Fast Paxos: the state of the dependency-set algorithm and of the Fast Paxos incarnations are merged so as to allow the reception of identical dependency sets from a fast quorum of processes
-(after an $\opn{announce}_p\left( c \right)$) to act as a single round-trip decision of Fast Paxos.
+First, EPaxos uses an algorithm similar to Fast Paxos~\cite{Lamport05GeneralizeConsensus} instead of MultiPaxos.
+Second, EPaxos combines the dependency-set algorithm outlined above (except for using larger quorums), with the fast rounds of Fast Paxos: the state of the dependency-set algorithm and of the Fast Paxos incarnations are merged to allow the reception of identical dependency sets from a fast quorum of processes
+(after an $\opn{announce}_p\left( c \right)$) to act as a single-round-trip decision of Fast Paxos.
 

### 2016-02-13 podc16/paper2.tex
minor fix

@@
 
-On top of enabling modular correctness proofs of algorithms like EPaxos, we expect that the modular structure of our generic leaderless algorithm will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by different implementations of our two abstractions in varied environments such clusters or wide area networks, and on varied metrics such as agreement latency, impact of failures or conflicts, tolerance to slow processes, load
+On top of enabling modular correctness proofs of algorithms like EPaxos, we expect that the modular structure of our generic leaderless algorithm will allow a principled theoretical and empirical evaluation of the trade-offs that can be achieved by different implementations of our two abstractions in varied environments such as clusters or wide area networks, and on varied metrics such as agreement latency, impact of failures or conflicts, tolerance to slow processes, load
 balancing, size of quorums, etc.

### 2016-02-13 podc16/paper2.tex
minor fix

@@
 
-We show that the core idea underlying EPaxos can be captured in a generic leaderless generalized consensus algorithm that uses two new abstractions: a \textit{dependency-set algorithm}, which suggests dependencies for commands, and a \textit{map-agreement algorithm}, which ensures that, for each submitted command, processes agree on a dependency set.
+We show that the core idea underlying EPaxos can be captured in a \emph{generic leaderless generalized consensus algorithm} that uses two new abstractions: a \textit{dependency-set algorithm}, which suggests dependencies for commands, and a \textit{map-agreement algorithm}, which ensures that, for each submitted command, processes agree on a dependency set.
 Both abstractions lend themselves well to implementations that avoid the use of a unique leader process.

### 2016-02-13 podc16/paper2.tex
minor fixes

@@
 We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network.
-Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}, in which each process receives proposals for commands of the form $\opn{GC-Propose}(c)$, and must produce responses of the form $\opn{GC-Decide}(\sigma)$, where $\sigma$ is a c-struct, and the sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
+Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}, in which each process receives proposals for commands of the form $\opn{GC-Propose}(c)$, must produce responses of the form $\opn{GC-Decide}(\sigma)$, where $\sigma$ is a c-struct, and the sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
 Briefly, the c-structs are a set, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct.
 Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands.
-We say that two commands when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
+We say that two commands \emph{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
 C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $s$ such that $\sigma_2 = \sigma_1\bullet s$. Moreover, $\bot\sqsubseteq \sigma$ for any
@@
 The consistency property of generalized consensus requires that any two decided c-structs be compatible.
-We use the same notation as and refer the reader to~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of c-structs and generalized consensus.
+We refer the reader to~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of c-structs and generalized consensus, and we use the same notation.
 

### 2016-02-15 podc16/paper2.tex
Work on GraphProcessing2.tla and a small fix in the paper.
Minor fixes to paper2.tex

@@
 
-We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network.
+We consider a set of processes $P$ which are subject to crash-stop faults and which communicate by message-passing in an asynchronous network.
 Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}, in which each process receives proposals for commands of the form $\opn{GC-Propose}(c)$, must produce responses of the form $\opn{GC-Decide}(\sigma)$, where $\sigma$ is a c-struct, and the sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
@@
 We say that two commands \emph{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
-C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $s$ such that $\sigma_2 = \sigma_1\bullet s$. Moreover, $\bot\sqsubseteq \sigma$ for any
+C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ if and only if there exists a sequence of commands $s$ such that $\sigma_2 = \sigma_1\bullet s$.
+Moreover, $\bot\sqsubseteq \sigma$ for any
 $\sigma$. Two c-structs are compatible when they have a
-common upper bound.
+common upper bound \todo{that is constructible from \ldots}.
 A c-struct contains a command $c$ when it is of the form $\bot\bullet cs$ where $c$ appears in the sequence of commands $cs$.
@@
 Upon  receiving a command $c$, $p$ proceeds in three phases.
-First, it calls a \textit{dependency-set} algorithm (\Cref{dep-algo}) to determine a set of commands $\mathcal{D}$ that the command $c$ may depend on.
+First, it calls a \textit{dependency-set} algorithm (\Cref{dep-algo}) to determine a set of commands $\mathcal{D}$ on which the command $c$ may depend.
 Second, $p$ proposes the mapping  $c\mapsto \mathcal{D}$ to a \textit{map-agreement} algorithm (\Cref{map-algo}); when the map-agreement algorithm
@@
 We say that a command $c$ in the domain of $m_p$ is \emph{executable} when no command $c'$ which is not in the domain of $m_p$ can be reached from $c$ by recursively following $c$'s dependencies.
-The local variable $m_p$ denotes a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the executable commands of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ iff $c_2 \in m_p\left[ c_1 \right]$ (i.e., $c_1$ depends on $c_2$).
+The local variable $m_p$ denotes a directed graph $g_p$ whose set of vertices $V\left( g_p \right)$ is the executable commands of $m_p$ and whose edges $E\left( g_p \right)$ are such that there is an edge from $c_1$ to $c_2$ if and only if $c_2 \in m_p\left[ c_1 \right]$ (i.e., $c_1$ depends on $c_2$).
 For example, if $m_p = \left[ c_1 \mapsto \aset{c_2,c_4}, c_2 \mapsto \aset{c_3}, c_3 \mapsto \aset{c_2}\right]$ then $V\left( g_p \right) = \aset{c_2,c_3}$ and $E\left( g_p \right) = \aset{\left( c_2,c_3 \right),\left( c_3,c_2 \right)}$.
 
-A directed graph $g$ \emph{induces a partial order} $\preceq_g$ on its vertices defined such that $c_1 \preceq_g c_2$ iff there is a path from $c_2$ to $c_1$ and none from $c_1$ to $c_2$.
+A directed graph $g$ \emph{induces a partial order} $\preceq_g$ on its vertices defined such that $c_1 \preceq_g c_2$ if and only if there is a path from $c_2$ to $c_1$ and none from $c_1$ to $c_2$.
 For example, consider $h$ where $V\left( h \right)=\aset{c_1,c_2,c_3,c_4}$ and $E\left( h \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right),\left( c_1,c_3 \right)}$. 
@@
 We now define a partial order on graphs.
-If $g_1$ and $g_2$ are two graphs, then we write $g_1 \leq g_2$ iff $V\left( g_1 \right)\subseteq V\left( g_2 \right)$, and if $v\in V\left( g_1 \right)$, $e\in E\left( g_2 \right)$, and $v$ is an endpoint of $e$, then $e\in E\left( g_1 \right)$.
+If $g_1$ and $g_2$ are two graphs, then we write $g_1 \leq g_2$ if and only if $V\left( g_1 \right)\subseteq V\left( g_2 \right)$, and if $v\in V\left( g_1 \right)$, $e\in E\left( g_2 \right)$, and $v$ is an endpoint of $e$, then $e\in E\left( g_1 \right)$.
 For example, consider the graphs $h'$ and $h''$ such that $V\left( h' \right)=\aset{c_1,c_2,c_4}$, $E\left( h' \right)=\aset{\left( c_1,c_2 \right),\left( c_2,c_1 \right)}$, $V\left( h'' \right)=\aset{c_1,c_3,c_4}$, and $E\left( h'' \right)=\aset{\left( c_1,c_3 \right)}$.
@@
 Define graph intersection such that $V\left( g_1 \cap g_2 \right)=V\left( g_1 \right)\cap V\left( g_2 \right)$ and $E\left( g_1 \cap g_2 \right)= E\left( g_1 \right)\cap E\left( g_2 \right)$, and graph union such that $V\left( g_1 \cup g_2 \right)=V\left( g_1 \right)\cup V\left( g_2 \right)$ and $E\left( g_1 \cup g_2 \right)= E\left( g_1 \right)\cup E\left( g_2 \right)$.
-We say that $g$ is a vertex-induced subgraph of $g'$ iff $V\left( g \right)\subseteq V\left( g' \right)$, and for every $e\in E\left( g' \right)$, if both endpoints of $e$ are in $V\left( g \right)$, then $e\in E\left( g \right)$.
-Define two graphs $g_1$ and $g_2$ as \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$.
+We say that $g$ is a vertex-induced subgraph of $g'$ if and only if $V\left( g \right)\subseteq V\left( g' \right)$, and for every $e\in E\left( g' \right)$, if both endpoints of $e$ are in $V\left( g \right)$, then $e\in E\left( g \right)$.
+Define two graphs $g_1$ and $g_2$ as \emph{compatible} if and only if $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$.
 
@@
 EPaxos improves upon this MultiPaxos scheme in two ways.
-First, EPaxos uses an algorithm similar to Fast Paxos~\cite{Lamport05GeneralizeConsensus} instead of MultiPaxos.
+First, EPaxos uses an algorithm similar to Fast Paxos~\cite{Lamport06FastPaxos} instead of MultiPaxos.
 Second, EPaxos combines the dependency-set algorithm outlined above (except for using larger quorums), with the fast rounds of Fast Paxos: the state of the dependency-set algorithm and of the Fast Paxos incarnations are merged to allow the reception of identical dependency sets from a fast quorum of processes
@@
 We also write the $\sigma\bullet\aseq{c_1,\cdots,c_n}$ for the c-struct $\sigma\bullet c_1\bullet\cdots\bullet c_n$.
-C-structs are partially ordered by the relation $\sqsubseteq$ defined such that $\sigma_1\sqsubseteq\sigma_2$ iff there exists a sequence of commands $\mathcal{D}$ such that $\sigma_2=\sigma_1\bullet \mathcal{D}$.
+C-structs are partially ordered by the relation $\sqsubseteq$ defined such that $\sigma_1\sqsubseteq\sigma_2$ if and only if there exists a sequence of commands $\mathcal{D}$ such that $\sigma_2=\sigma_1\bullet \mathcal{D}$.
 Two c-structs are compatible when they have a common upper bound; when this is the case, two c-structs $\sigma_1$ and $\sigma_2$ have a least upper bound, noted $\bigsqcup\left( \sigma_1,\sigma_2 \right)$.
@@
 The provided implementation satisfies the specification of a dependency-set algorithm for the following reasons.
-Since any command in a dependency-set that is suggested by a process (lines~\ref{alg:update-deps-p} and~\ref{alg:call-suggest}) must be in the set $cmds_q$ of the commands received by some process $q$ (line~\ref{alg:update-d}), and a process receives a command iff there exists some process that previously announced the command (line~\ref{alg:send-new-cmd}), then it follows that any command in any suggested dependency-set has been previously announced.
+Since any command in a dependency-set that is suggested by a process (lines~\ref{alg:update-deps-p} and~\ref{alg:call-suggest}) must be in the set $cmds_q$ of the commands received by some process $q$ (line~\ref{alg:update-d}), and a process receives a command if and only if there exists some process that previously announced the command (line~\ref{alg:send-new-cmd}), then it follows that any command in any suggested dependency-set has been previously announced.
 Further, since a process can suggest a set of dependencies $deps_p[c]$ for a command $c$ after it has received a reply for $c$ from at least a quorum of processes (lines~\ref{alg:receive-deps}--\ref{alg:call-suggest}), we have that $\opn{suggest}_p \left( c,deps_p[c]\right)$ is called only if a quorum of processes $q$ have $c$ in their $cmds_q$ (line~\ref{alg:update-cmds-p}).

### 2016-05-17 podc16/paper2.tex
Working on the PODC camera ready.

@@
 
+\CopyrightYear{2016}
+\setcopyright{rightsretained}
+\conferenceinfo{PODC'16}{July 25-28, 2016, Chicago, IL, USA}
+\isbn{978-1-4503-3964-3/16/07}
+\doi{http://dx.doi.org/10.1145/2933057.2933072}
+
 \maketitle
@@
 In practice, the unique leader can easily become a performance bottleneck for the system.
-Therefore, recently proposed agreement algorithms like Mencius~\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine}, EPaxos~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments}, or Alvin~\cite{TurcuETAL14BeGeneralDontGiveUpConsistency} strive to increase performance by avoiding the use of a unique leader.
-Mencius rotates the leader among the processes but its structure remains close to MultiPaxos. 
+Therefore, recently proposed agreement algorithms like S-Paxos~\cite{BielyETAL12SpaxosOffloadingLeaderHighThroughputState}, Mencius~\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine}, EPaxos~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments}, or Alvin~\cite{TurcuETAL14BeGeneralDontGiveUpConsistency} strive to increase performance by avoiding the use of a unique leader.
+S-Paxos offloads the computation that the leader has to perform to other replicas, while Mencius rotates the leader among the processes; the structure of both algorithms remains close to MultiPaxos. 
 However, EPaxos and Alvin are based on a novel algorithmic idea first introduced in EPaxos.
@@
 balancing, size of quorums, etc.
-We have formalized\footnote{\url{http://losa.fr/leaderless}} and model-checked our generic leaderless algorithm with the TLA$^+$ toolbox.
+We have formalized\footnote{\url{http://losa.fr/leaderless}} and model-checked our generic leaderless algorithm using the TLA$^+$ toolbox.
 
@@
 Moreover, $\bot\sqsubseteq \sigma$ for any
-$\sigma$. Two c-structs are compatible when they have a
-common upper bound \todo{that is constructible from \ldots}.
+$\sigma$. 
 A c-struct contains a command $c$ when it is of the form $\bot\bullet cs$ where $c$ appears in the sequence of commands $cs$. 
-The consistency property of generalized consensus requires that any two decided c-structs be compatible.
+Two c-structs are compatible when they have a common upper bound that is constructible from the commands contained in the two c-structs. 
+The non-triviality property of generalized consensus requires that any decided c-struct be of the form $\bot\bullet \mathcal{D}$, where $\mathcal{D}$ is a sequence of proposed commands; consistency requires than any two decided c-structs $\sigma_1$ and $\sigma_2$ be compatible; stability requires that when a process $p$ decides a c-struct $\sigma_1$ and time $t_1$ and $\sigma_2$ at time $t_2$, then $t_1 \leq t_2$ implies that $\sigma_1\sqsubseteq \sigma_2$; finally, liveness requires that
+if a command keeps beeing proposed, then a c-struct $\sigma$ containing the command is eventually decided.
 We refer the reader to~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of c-structs and generalized consensus, and we use the same notation.
